---
title: "Untitled"
author: "CLEANINCLE API"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# CLEANinCLE API

```{r}
library(httr2)
library(jsonlite)
library(dplyr)
library(tidyr)
library(readr)
library(lubridate)


# 1. API settings

api_url <- "https://data.cleanincle.issacs.dev"
api_key <- "2KYc26u6sn6bPAaZHJx5R3DSzQL8Y6sV38hy0c3t"

params <- list(
  start_date = "2024-11-01T00:00:00Z",
  end_date   = "2026-01-01T00:00:00Z"
)

# 2. Download all pages
# -----------------------------
all_records <- list()
current_params <- params

message("Starting CLEANinCLE API download...")

repeat {
  resp <- request(api_url) |>
    req_headers(`x-api-key` = api_key) |>
    req_url_query(!!!current_params) |>
    req_retry(max_tries = 3) |>
    req_perform()

  body <- resp_body_json(resp, simplifyVector = FALSE)

  all_records <- c(all_records, body$records)

  if (is.null(body$next_token)) break

  current_params$next_token <- body$next_token
}

# -----------------------------
# 3. Convert to dataframe
# -----------------------------
df_raw <- bind_rows(all_records)

df_raw <- df_raw %>%
  unnest_wider(where(is.list), names_sep = "_")

# -----------------------------
# 4. Save raw downloaded file
# -----------------------------
write_csv(df_raw, "CLEANinCLE_raw_2024-11_to_2025-12.csv")
saveRDS(df_raw, "CLEANinCLE_raw_2024-11_to_2025-12.rds")

message("Download complete.")
dim(df_raw)
names(df_raw)

```


# PM2.5-only CLEANinCLE dataset
```{r}
library(dplyr)
library(lubridate)
library(readr)

cinc_pm25 <- df_raw %>%
  mutate(
    time = ymd_hms(time, tz = "UTC"),
    date = as.Date(time),
    pm_2.5 = as.numeric(pm_2.5)
  ) %>%
  select(
    device_eui,
    location_name,
    latitude,
    longitude,
    device_number,
    device_status,
    time,
    date,
    pm_2.5
  ) %>%
  filter(!is.na(pm_2.5)) %>%
  arrange(device_number, time)

# Save PM2.5-only dataset
write_csv(cinc_pm25, "CLEANinCLE_PM25_2024-11_to_2025-12.csv")
saveRDS(cinc_pm25, "CLEANinCLE_PM25_2024-11_to_2025-12.rds")

# Checks
dim(cinc_pm25)
names(cinc_pm25)
head(cinc_pm25)
range(cinc_pm25$time, na.rm = TRUE)
```

# Exploratory Data Analysis - Sensors in the dataset
```{r}
# CinC devices present
cinc_pm25 %>%
  distinct(device_number, location_name) %>%
  arrange(device_number)

# Number of observations per device
cinc_pm25 %>%
  count(device_number, location_name, sort = TRUE)

# missing PM2.5 values (filtered NAs)
sum(is.na(cinc_pm25$pm_2.5))

# PM2.5 statistics (Detect outliers or unrealistic values)
summary(cinc_pm25$pm_2.5)

# PM2.5 distribution (Histogram) - [if values are skewed/there are extreme spikes]
library(ggplot2)

ggplot(cinc_pm25, aes(x = pm_2.5)) +
  geom_histogram(bins = 50) +
  labs(
    title = "Distribution of PM2.5 Measurements (CLEANinCLE)",
    x = "PM2.5 (µg/m³)",
    y = "Frequency"
  )

# Time coverage check
range(cinc_pm25$time)

# Time series preview (to confirm the time data behaves correctly)
library(ggplot2)

ggplot(
  cinc_pm25 %>% filter(device_number == unique(device_number)[1]),
  aes(time, pm_2.5)
) +
  geom_point(alpha = 0.4, size = 0.7) +
  scale_x_datetime(
    date_breaks = "1 month",
    date_labels = "%b %Y"
  ) +
  labs(
    title = "PM2.5 Time Series (One CinC Sensor)",
    x = "Time",
    y = "PM2.5 (µg/m³)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(
  cinc_pm25 %>% filter(device_number == unique(device_number)[1]),
  aes(time, pm_2.5)
) +
  geom_point(alpha = 0.4, size = 0.7) +
  scale_x_datetime(
    limits = as.POSIXct(c("2024-11-01", "2025-12-31")),
    date_breaks = "1 month",
    date_labels = "%b %Y"
  ) +
  labs(
    title = "PM2.5 Time Series (One CinC Sensor)",
    x = "Time",
    y = "PM2.5 (µg/m³)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

cinc_pm25 %>%
  filter(device_number == unique(device_number)[1]) %>%
  count(month = lubridate::floor_date(time, "month"))

```
```{r}
library(ggplot2)
library(dplyr)

# get all device numbers
device_ids <- sort(unique(cinc_pm25$device_number))

# create folder to save plots
dir.create("CinC_PM25_timeseries_plots", showWarnings = FALSE)

for (dev in device_ids) {
  
  sensor_data <- cinc_pm25 %>% filter(device_number == dev)
  loc_name <- unique(sensor_data$location_name)[1]
  
  p <- ggplot(sensor_data, aes(time, pm_2.5)) +
    geom_point(alpha = 0.4, size = 0.7) +
    scale_x_datetime(
      limits = as.POSIXct(c("2024-11-01", "2025-12-31")),
      date_breaks = "1 month",
      date_labels = "%b %Y"
    ) +
    labs(
      title = paste("PM2.5 Time Series - Device", dev, "-", loc_name),
      x = "Time",
      y = "PM2.5 (µg/m³)"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  print(p)
  
  ggsave(
    filename = paste0("CinC_PM25_timeseries_plots/CinC_sensor_", dev, ".png"),
    plot = p,
    width = 10,
    height = 5,
    dpi = 300
  )
}

```

```{r}
# Sensor coverage summary (start, end, number of observations)
cinc_pm25 %>%
  mutate(month = lubridate::floor_date(time, "month")) %>%
  count(device_number, location_name, month) %>%
  arrange(device_number, month)

library(gt)

cinc_summary <- cinc_pm25 %>%
  group_by(device_number, location_name) %>%
  summarise(
    start_time = min(time, na.rm = TRUE),
    end_time = max(time, na.rm = TRUE),
    n_obs = n(),
    .groups = "drop"
  ) %>%
  arrange(device_number)

gt_table <- gt(cinc_summary)

gtsave(gt_table, "CLEANinCLE_sensor_summary.pdf")

# Monthly coverage for all sensors (shows how many measurements each sensor has per month)
cinc_pm25 %>%
  mutate(month = lubridate::floor_date(time, "month")) %>%
  count(device_number, location_name, month) %>%
  arrange(device_number, month)

cinc_pm25 %>%
  mutate(
    month = lubridate::floor_date(time, "month"),
    device_number = factor(device_number)
  ) %>%
  distinct(device_number, location_name, month) %>%
  ggplot(aes(x = month, y = device_number)) +
  geom_point(size = 2) +
  scale_x_datetime(
    limits = as.POSIXct(c("2024-11-01", "2025-12-31")),
    date_breaks = "1 month",
    date_labels = "%b %Y"
  ) +
  labs(
    title = "CLEANinCLE Sensor Data Coverage",
    x = "Month",
    y = "Sensor"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
# Daily PM2.5 datasets
```{r}
cinc_daily <- cinc_pm25 %>%
  group_by(device_number, date) %>%
  summarise(
    pm25_daily = mean(pm_2.5, na.rm = TRUE),
    n_obs = n(),
    .groups = "drop"
  ) %>%
  arrange(device_number, date)

# check
dim(cinc_daily)
head(cinc_daily)

# Convert to WIDE format (To compare sensors, each sensor needs its own column)
library(tidyr)

cinc_daily_wide <- cinc_daily %>%
  select(device_number, date, pm25_daily) %>%
  pivot_wider(
    names_from = device_number,
    values_from = pm25_daily,
    names_prefix = "device_"
  )

head(cinc_daily_wide)

#Pairwise correlations between sensors (This tells which sensors behave similarly, which sensors disagree)
# High correlation → sensors measure similar PM patterns.
cor_matrix <- cor(
  cinc_daily_wide %>% select(-date),
  use = "pairwise.complete.obs"
)

cor_matrix
write.csv(cor_matrix, "CinC_sensor_correlation_matrix.csv")

#Visualize sensor agreement (This produces a sensor correlation heatmap)
library(ggplot2)
library(tidyr)

cor_df <- as.data.frame(cor_matrix)

cor_long <- cor_df %>%
  tibble::rownames_to_column("Sensor1") %>%
  pivot_longer(-Sensor1, names_to = "Sensor2", values_to = "correlation")

ggplot(cor_long, aes(Sensor1, Sensor2, fill = correlation)) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(
    title = "Correlation Between CinC Sensors (Daily PM2.5)",
    x = "Sensor",
    y = "Sensor"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


```{r}
library(pheatmap)

pheatmap(
  cor_matrix,
  main = "Correlation Between CinC Sensors (Daily PM2.5)"
)

pdf("CinC_sensor_correlation_heatmap.pdf", width = 10, height = 8)

dev.off()

# Extract sensor coordinates

library(dplyr)

sensor_locations <- cinc_pm25 %>%
  select(device_number, latitude, longitude) %>%
  distinct() %>%
  mutate(
    latitude = as.numeric(latitude),
    longitude = as.numeric(longitude),
    sensor = paste0("device_", device_number)
  ) %>%
  arrange(device_number)

sensor_locations
str(sensor_locations)










sensor_locations <- cinc_pm25 %>%
  select(device_number, latitude, longitude) %>%
  distinct() %>%
  arrange(device_number)

sensor_locations

# distance between sensor
library(geosphere)

coords <- sensor_locations %>%
  select(longitude, latitude)

dist_matrix <- distm(coords, fun = distHaversine) / 1000

library(tidyr)

cor_long <- as.data.frame(as.table(cor_matrix))

names(cor_long) <- c("sensor1","sensor2","correlation")

sensor_compare <- cor_long %>%
  left_join(dist_long, by = c("sensor1","sensor2"))

head(cor_long)

library(ggplot2)

ggplot(sensor_compare, aes(distance_km, correlation)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", color = "red") +
  labs(
    title = "Sensor Similarity vs Distance",
    x = "Distance Between Sensors (km)",
    y = "Correlation of Daily PM2.5"
  ) +
  theme_minimal()
```



```{r}

library(dplyr)

sensor_locations <- cinc_pm25 %>%
  select(device_number, latitude, longitude) %>%
  distinct() %>%
  mutate(
    latitude = as.numeric(latitude),
    longitude = as.numeric(longitude),
    sensor = paste0("device_", device_number)
  ) %>%
  arrange(device_number)

sensor_locations
str(sensor_locations)

library(geosphere)

coords <- sensor_locations %>%
  select(longitude, latitude) %>%
  as.matrix()

dist_matrix <- distm(coords, fun = distHaversine) / 1000

rownames(dist_matrix) <- sensor_locations$sensor
colnames(dist_matrix) <- sensor_locations$sensor

dist_long <- as.data.frame(as.table(dist_matrix))
colnames(dist_long) <- c("sensor1", "sensor2", "distance_km")

head(dist_long)

cor_long <- as.data.frame(as.table(cor_matrix))
colnames(cor_long) <- c("sensor1", "sensor2", "correlation")

head(cor_long)

sensor_compare <- cor_long %>%
  left_join(dist_long, by = c("sensor1", "sensor2")) %>%
  filter(sensor1 != sensor2)

head(sensor_compare)

library(ggplot2)

ggplot(sensor_compare, aes(x = distance_km, y = correlation)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm") +
  labs(
    title = "Correlation vs Distance Between CinC Sensors",
    x = "Distance Between Sensors (km)",
    y = "PM2.5 Correlation"
  ) +
  theme_minimal()
```



```{r}

sensor_locations %>%
  count(device_number)

range(sensor_locations$latitude)
range(sensor_locations$longitude)
```

```{r}
sensor_locations_clean <- cinc_pm25 %>%
  mutate(
    latitude = as.numeric(latitude),
    longitude = as.numeric(longitude)
  ) %>%
  filter(
    !is.na(latitude),
    !is.na(longitude),
    latitude != 0,
    longitude != 0
  ) %>%
  group_by(device_number) %>%
  summarise(
    latitude = first(latitude),
    longitude = first(longitude),
    .groups = "drop"
  ) %>%
  mutate(
    sensor = paste0("device_", device_number)
  ) %>%
  arrange(as.numeric(device_number))

sensor_locations_clean %>%
  count(device_number)

range(sensor_locations_clean$latitude)
range(sensor_locations_clean$longitude)

sensor_locations_clean

library(geosphere)

coords <- sensor_locations_clean %>%
  select(longitude, latitude) %>%
  as.matrix()

dist_matrix <- distm(coords, fun = distHaversine) / 1000

rownames(dist_matrix) <- sensor_locations_clean$sensor
colnames(dist_matrix) <- sensor_locations_clean$sensor

dist_long <- as.data.frame(as.table(dist_matrix))
colnames(dist_long) <- c("sensor1", "sensor2", "distance_km")

head(dist_long)

cor_long <- as.data.frame(as.table(cor_matrix))
colnames(cor_long) <- c("sensor1", "sensor2", "correlation")

sensor_compare <- cor_long %>%
  left_join(dist_long, by = c("sensor1", "sensor2")) %>%
  filter(sensor1 != sensor2)

head(sensor_compare)

library(ggplot2)

ggplot(sensor_compare, aes(x = distance_km, y = correlation)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm") +
  labs(
    title = "Correlation vs Distance Between CinC Sensors",
    x = "Distance Between Sensors (km)",
    y = "PM2.5 Correlation"
  ) +
  theme_minimal()

sensor_locations_clean %>% count(device_number)
range(sensor_locations_clean$latitude)
range(sensor_locations_clean$longitude)
```



```{r}
library(ggplot2)
library(maps)
library(dplyr)

county_map <- map_data("county")
cuyahoga_map <- county_map %>%
  filter(region == "ohio", subregion == "cuyahoga")

ggplot() +
  geom_polygon(
    data = cuyahoga_map,
    aes(x = long, y = lat, group = group),
    fill = "gray95",
    color = "black"
  ) +
  geom_point(
    data = sensor_locations_clean,
    aes(x = longitude, y = latitude),
    color = "firebrick",
    size = 3
  ) +
  coord_fixed(
    xlim = c(-81.85, -81.45),
    ylim = c(41.35, 41.65)
  ) +
  labs(
    title = "CLEANinCLE Sensor Locations (Cuyahoga County)",
    x = "Longitude",
    y = "Latitude"
  ) +
  theme_minimal()
```

 # make one clean location per sensor
```{r}
sensor_locations_clean <- cinc_pm25 %>%
  mutate(
    latitude = as.numeric(latitude),
    longitude = as.numeric(longitude)
  ) %>%
  filter(
    !is.na(latitude),
    !is.na(longitude),
    latitude != 0,
    longitude != 0
  ) %>%
  group_by(device_number) %>%
  summarise(
    latitude = first(latitude),
    longitude = first(longitude),
    .groups = "drop"
  ) %>%
  mutate(sensor = paste0("device_", device_number))

# make the distance matrix
library(geosphere)
library(dplyr)

coords <- sensor_locations_clean %>%
  select(longitude, latitude) %>%
  as.matrix()

dist_matrix <- distm(coords, fun = distHaversine) / 1000

rownames(dist_matrix) <- sensor_locations_clean$sensor
colnames(dist_matrix) <- sensor_locations_clean$sensor

# turn both matrices into long tables
library(tidyr)

cor_long <- as.data.frame(as.table(cor_matrix))
colnames(cor_long) <- c("sensor1", "sensor2", "correlation")

dist_long <- as.data.frame(as.table(dist_matrix))
colnames(dist_long) <- c("sensor1", "sensor2", "distance_km")

# join them
sensor_compare <- cor_long %>%
  left_join(dist_long, by = c("sensor1", "sensor2")) %>%
  filter(sensor1 != sensor2)

sensor_compare <- sensor_compare %>%
  mutate(
    sensor1 = as.character(sensor1),
    sensor2 = as.character(sensor2)
  )

# remove duplicate pairs
sensor_compare_unique <- sensor_compare %>%
  rowwise() %>%
  mutate(
    pair1 = min(sensor1, sensor2),
    pair2 = max(sensor1, sensor2)
  ) %>%
  ungroup() %>%
  distinct(pair1, pair2, .keep_all = TRUE)

# plot correlation vs distance
library(ggplot2)

ggplot(sensor_compare_unique, aes(distance_km, correlation)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = "Correlation vs Distance Between CLEANinCLE Sensors",
    x = "Distance between sensors (km)",
    y = "Correlation of daily PM2.5"
  ) +
  theme_minimal()

```
# sensors with unusually low correlation
```{r}
sensor_compare_unique %>%
  filter(correlation < 0.4) %>%
  arrange(correlation)

# average correlation by distance group
sensor_compare_unique %>%
  mutate(distance_group = cut(
    distance_km,
    breaks = c(0,2,5,10,20,30),
    labels = c("0–2 km","2–5 km","5–10 km","10–20 km","20–30 km")
  )) %>%
  group_by(distance_group) %>%
  summarise(mean_correlation = mean(correlation, na.rm = TRUE)) 
```

# pairwise comparison for CLEANinCLE sites using daily averages.
```{r}
# daily averages for each CinC sensor
library(dplyr)
library(tidyr)
library(ggplot2)
library(reshape2)
library(geosphere)

# Daily averages for each CinC sensor
cinc_daily <- cinc_pm25 %>%
  mutate(date = as.Date(time)) %>%
  group_by(device_number, date) %>%
  summarise(
    pm25_daily = mean(pm_2.5, na.rm = TRUE),
    n_obs = n(),
    .groups = "drop"
  ) %>%
  arrange(device_number, date)

head(cinc_daily)

# Convert to wide format
cinc_daily_wide <- cinc_daily %>%
  pivot_wider(
    names_from = device_number,
    values_from = pm25_daily,
    names_prefix = "device_"
  )

head(cinc_daily_wide)

# pairwise correlation matrix
cinc_cor_matrix <- cor(
  cinc_daily_wide %>% select(-date),
  use = "pairwise.complete.obs"
)

round(cinc_cor_matrix, 2)

# Plot the CinC correlation heatmap
cinc_cor_long <- melt(cinc_cor_matrix)

ggplot(cinc_cor_long, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_viridis_c(name = "Correlation") +
  labs(
    title = "Correlation Between CinC Sensors (Daily PM2.5)",
    x = "Sensor",
    y = "Sensor"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  )

# clean sensor location table
sensor_locations_clean <- cinc_pm25 %>%
  mutate(
    latitude = as.numeric(latitude),
    longitude = as.numeric(longitude)
  ) %>%
  filter(
    !is.na(latitude),
    !is.na(longitude),
    latitude != 0,
    longitude != 0
  ) %>%
  group_by(device_number) %>%
  summarise(
    latitude = first(latitude),
    longitude = first(longitude),
    .groups = "drop"
  ) %>%
  mutate(sensor = paste0("device_", device_number)) %>%
  arrange(sensor)

sensor_locations_clean

# distance matrix between CinC sensors
coords <- sensor_locations_clean %>%
  select(longitude, latitude) %>%
  as.matrix()

dist_matrix <- distm(coords, fun = distHaversine) / 1000

rownames(dist_matrix) <- sensor_locations_clean$sensor
colnames(dist_matrix) <- sensor_locations_clean$sensor

# correlation and distance to long format
cor_long <- as.data.frame(as.table(cinc_cor_matrix))
colnames(cor_long) <- c("sensor1", "sensor2", "correlation")

dist_long <- as.data.frame(as.table(dist_matrix))
colnames(dist_long) <- c("sensor1", "sensor2", "distance_km")

# Join them and remove duplicate pairs
sensor_compare <- cor_long %>%
  mutate(
    sensor1 = as.character(sensor1),
    sensor2 = as.character(sensor2)
  ) %>%
  left_join(dist_long, by = c("sensor1", "sensor2")) %>%
  filter(sensor1 != sensor2) %>%
  rowwise() %>%
  mutate(
    pair1 = min(sensor1, sensor2),
    pair2 = max(sensor1, sensor2)
  ) %>%
  ungroup() %>%
  distinct(pair1, pair2, .keep_all = TRUE) %>%
  filter(!is.na(correlation), !is.na(distance_km))

head(sensor_compare)

# correlation vs distance for CinC daily averages
ggplot(sensor_compare, aes(x = distance_km, y = correlation)) +
  geom_point(alpha = 0.35, size = 2) +
  geom_smooth(method = "lm", color = "blue", se = TRUE, linewidth = 1) +
  coord_cartesian(ylim = c(0.5, 1.02)) +
  labs(
    title = "Correlation vs Distance Between CLEANinCLE Sensors 
    (Daily PM2.5)",
    x = "Distance between sensors (km)",
    y = "Correlation"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    panel.grid.minor = element_blank()
  )

```
# Pairwise comparison for CLEANinCLE sites (Hourly averages)
```{r}
# CinC half-hour data to hourly averages

library(dplyr)
library(lubridate)

cinc_hourly <- cinc_pm25 %>%
  mutate(
    datetime = as.POSIXct(time),
    hour = floor_date(datetime, "hour")
  ) %>%
  group_by(device_number, hour) %>%
  summarise(
    pm25_hourly = mean(pm_2.5, na.rm = TRUE),
    .groups = "drop"
  )

# wide format (one column per sensor)
library(tidyr)

cinc_hourly_wide <- cinc_hourly %>%
  pivot_wider(
    names_from = device_number,
    values_from = pm25_hourly,
    names_prefix = "device_"
  )

# correlation matrix
cinc_hourly_cor <- cor(
  cinc_hourly_wide %>% select(-hour),
  use = "pairwise.complete.obs"
)

# Convert correlations to long format
library(reshape2)

cor_long <- as.data.frame(as.table(cinc_hourly_cor))
colnames(cor_long) <- c("sensor1", "sensor2", "correlation")

# Combine with distance table
sensor_compare_hourly <- cor_long %>%
  left_join(dist_long, by = c("sensor1", "sensor2")) %>%
  filter(sensor1 != sensor2)

# Plot correlation vs distance (hourly)
library(ggplot2)

ggplot(sensor_compare_hourly, aes(distance_km, correlation)) +
  geom_point(alpha = 0.35, size = 2) +
  geom_smooth(method = "lm", color = "blue", se = TRUE) +
  coord_cartesian(ylim = c(0.5, 1.02)) +
  labs(
    title = "Correlation vs Distance Between CLEANinCLE Sensors (Hourly PM2.5)",
    x = "Distance between sensors (km)",
    y = "Correlation"
  ) +
  theme_minimal(base_size = 13)


```



# EPA Analysis
```{r}
library(RAQSAPI)
library(purrr)
library(dplyr)

aqs_credentials(username = "rnl34@case.edu", key = "copperhawk19")

site_numbers <- c("0034", "0038", "0045", "0060", "0065", "0073")

epa_data_list <- map(site_numbers, function(site) {
  RAQSAPI::aqs_sampledata_by_site(
    parameter = "88101",
    bdate = as.Date("20241101", format = "%Y%m%d"),
    edate = as.Date("20251231", format = "%Y%m%d"),
    stateFIPS = "39",
    countycode = "035",
    sitenum = site,
    return_header = FALSE
  ) %>%
    mutate(site_number = site)
})

epa_data <- bind_rows(epa_data_list)

head(epa_data)
dim(epa_data)

# daily averages:
epa_daily <- epa_data %>%
  mutate(date = as.Date(date_local)) %>%
  group_by(site_number, date) %>%
  summarise(
    pm25_daily = mean(sample_measurement, na.rm = TRUE),
    n_obs = n(),
    .groups = "drop"
  ) %>%
  arrange(site_number, date)

head(epa_daily)
```
# Convert EPA data to wide format
```{r}
library(tidyr)

epa_daily_wide <- epa_daily %>%
  pivot_wider(
    names_from = site_number,
    values_from = pm25_daily,
    names_prefix = "site_"
  )

head(epa_daily_wide)

# EPA correlation matrix
epa_cor_matrix <- cor(
  epa_daily_wide %>% select(-date),
  use = "pairwise.complete.obs"
)

round(epa_cor_matrix, 2)

# EPA correlation heatmap
library(reshape2)
library(ggplot2)

epa_cor_long <- melt(epa_cor_matrix)

ggplot(epa_cor_long, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(
    title = "EPA PM2.5 Correlation Between Sites",
    x = "Site",
    y = "Site",
    fill = "Correlation"
  ) +
  theme_minimal()

round(epa_cor_matrix, 2)

#
library(reshape2)

epa_cor_df <- melt(epa_cor_matrix) %>%
  filter(Var1 != Var2) %>%
  rename(
    site1 = Var1,
    site2 = Var2,
    correlation = value
  )
```

```{r}
# epa_locations 
epa_locations <- epa_data %>%
  mutate(site = paste0("site_", site_number)) %>%
  select(site, latitude, longitude) %>%
  distinct() %>%
  rename(
    lat = latitude,
    lon = longitude
  ) %>%
  arrange(site)

epa_locations

epa_locations %>%
  count(site)

epa_locations

```
# EPA FRM CORRELATION vs. DISTANCE 
```{r}
# distances between EPA sites
library(geosphere)

epa_pairs <- expand.grid(
  site1 = epa_locations$site,
  site2 = epa_locations$site
)

epa_pairs <- epa_pairs %>%
  left_join(epa_locations, by = c("site1" = "site")) %>%
  rename(lat1 = lat, lon1 = lon) %>%
  left_join(epa_locations, by = c("site2" = "site")) %>%
  rename(lat2 = lat, lon2 = lon) %>%
  mutate(
    distance_km = distHaversine(
      cbind(lon1, lat1),
      cbind(lon2, lat2)
    ) / 1000
  )

head(epa_pairs)

# merge with correlations
epa_compare <- epa_cor_df %>%
  left_join(
    epa_pairs,
    by = c("site1", "site2")
  )

# EPA correlation vs distance
epa_compare <- epa_compare %>%
  filter(!is.na(correlation), !is.na(distance_km))

ggplot(epa_compare, aes(distance_km, correlation)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", color = "blue") +
  labs(
    title = "EPA FRM PM2.5 Correlation vs Distance",
    x = "Distance between EPA sites (km)",
    y = "Correlation"
  ) +
  coord_cartesian(ylim = c(0.80, 1.00)) +
  theme_minimal()
```

# FRM filter data (24-hour reference measurements)
```{r}
epa_frm_24hr <- epa_data %>%
  filter(
    method_type == "FRM",
    sample_duration == "24 HOUR"
  )

head(epa_frm_24hr)
```


# FEM optical daily averages
```{r}
epa_fem_24hr <- epa_data %>%
  filter(
    method_type == "FEM",
    sample_duration == "24 HOUR"
  )

epa_data %>%
  count(method_type, sample_duration)
```

```{r}

# EPA hourly dataset
epa_fem_hourly <- epa_data %>%
  filter(method_type == "FEM", sample_duration == "1 HOUR")

# daily averages from hourly FEM
epa_fem_daily <- epa_fem_hourly %>%
  mutate(date = as.Date(date_local)) %>%
  group_by(site_number, date) %>%
  summarise(
    fem_pm25_daily = mean(sample_measurement, na.rm = TRUE),
    n_obs = n(),
    .groups = "drop"
  )

# FRM filter dataset
epa_frm_daily <- epa_data %>%
  filter(method_type == "FRM", sample_duration == "24 HOUR") %>%
  mutate(date = as.Date(date_local)) %>%
  select(site_number, date, sample_measurement) %>%
  rename(pm25_filter = sample_measurement)

# EPA optical vs filter (Merge them by site and date)
epa_compare <- inner_join(
  epa_frm_daily,
  epa_fem_daily,
  by = c("site_number", "date")
)

head(epa_compare)

ggplot(epa_compare,
       aes(pm25_filter, fem_pm25_daily)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue") +
  labs(
    title = "EPA Optical vs Filter PM2.5",
    x = "Filter (FRM) PM2.5",
    y = "Optical (FEM) PM2.5"
  ) +
  theme_minimal()
```


```{r figA-plot, fig.width=12, fig.height=9}
library(dplyr)
library(ggplot2)
library(scales)

# Build plotting table
figA_table <- epa_compare %>%
  filter(!is.na(pm25_filter), !is.na(pm25_daily)) %>%
  mutate(Site = site_number) %>%
  select(
    Site,
    PM25_FRM = pm25_filter,
    PM25_FEM = pm25_daily
  )

# Fixed panel limits to match template style
panel_min <- 0
panel_max <- 50
lims <- c(panel_min, panel_max)

# Per-site regression labels
lab_tbl <- figA_table %>%
  group_by(Site) %>%
  summarise(
    n = n(),
    slope = if (n() > 2) coef(lm(PM25_FEM ~ PM25_FRM))[2] else NA_real_,
    intercept = if (n() > 2) coef(lm(PM25_FEM ~ PM25_FRM))[1] else NA_real_,
    r2 = if (n() > 2) summary(lm(PM25_FEM ~ PM25_FRM))$r.squared else NA_real_,
    .groups = "drop"
  ) %>%
  mutate(
    label = ifelse(
      is.finite(r2),
      sprintf("y = %.2fx %+ .2f\nR² = %.2f", slope, intercept, r2),
      "n < 3"
    ),
    x = panel_min + (panel_max - panel_min) * 0.04,
    y = panel_max - (panel_max - panel_min) * 0.07
  )

ggplot(figA_table, aes(PM25_FRM, PM25_FEM)) +
  geom_point(alpha = 0.9, size = 2) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.9, color = "steelblue") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
  coord_equal(xlim = lims, ylim = lims, expand = FALSE) +
  facet_wrap(~Site, ncol = 3) +
  geom_text(
    data = lab_tbl,
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust = 0,
    vjust = 1,
    size = 4
  ) +
  labs(
    title = "EPA Continuous (FEM daily avg) vs EPA FRM (24hr) — by site",
    x = "EPA FRM PM2.5 (µg/m³)",
    y = "EPA Continuous PM2.5 (µg/m³)"
  ) +
  scale_x_continuous(breaks = seq(0, 50, 5)) +
  scale_y_continuous(breaks = seq(0, 50, 5)) +
  theme_minimal(base_size = 14) +
  theme(
    panel.spacing = unit(1.5, "lines"),
    axis.title = element_text(face = "bold"),
    plot.title = element_text(face = "bold", hjust = 0.5),
    strip.text = element_text(face = "bold", size = 14),
    panel.grid.minor = element_blank()
  )

```

# Compare CinC vs EPA (Filter / FRM)
```{r cinc_epa_frm_comparison, fig.width=10, fig.height=7, echo=FALSE}
# Identify the nearest CinC sensor to each EPA site
library(dplyr)
library(geosphere)

# EPA coordinates
epa_coords <- epa_locations %>%
  select(site, lat, lon)

# CinC coordinates
cinc_coords <- sensor_locations_clean %>%
  select(device_number, latitude, longitude)

# Find nearest CinC sensor for each EPA site
epa_cinc_pairs <- epa_coords %>%
  rowwise() %>%
  mutate(
    nearest_device = cinc_coords$device_number[
      which.min(
        distHaversine(
          c(lon, lat),
          as.matrix(cinc_coords[, c("longitude", "latitude")])
        )
      )
    ]
  ) %>%
  ungroup()

epa_cinc_pairs

# EPA daily filter data
epa_filter_daily <- epa_compare %>%
  select(date, site_number, pm25_filter) %>%
  distinct()

# CinC daily averages
cinc_daily <- cinc_pm25 %>%
  mutate(date = as.Date(time)) %>%
  group_by(device_number, date) %>%
  summarise(
    pm25_cinc = mean(pm_2.5, na.rm = TRUE),
    .groups = "drop"
  )

# EPA–CinC pairing table 
# Rename since epa_cinc_pairs currently uses site, while epa_filter_daily uses site_number
library(dplyr)

epa_cinc_pairs2 <- epa_cinc_pairs2 %>%
  mutate(site_number = sub("^site_", "", site_number))

epa_filter_daily <- epa_filter_daily %>%
  mutate(site_number = as.character(site_number))

cinc_daily <- cinc_daily %>%
  mutate(device_number = as.character(device_number))

# EPA filter data with nearest CinC daily data
epa_cinc_compare <- epa_cinc_pairs2 %>%
  left_join(epa_filter_daily, by = "site_number") %>%
  left_join(
    cinc_daily,
    by = c("nearest_device" = "device_number", "date" = "date")
  )

epa_cinc_compare_plot <- epa_cinc_compare %>%
  filter(!is.na(pm25_filter), !is.na(pm25_cinc))

#Check
head(epa_cinc_compare)

# Plot CinC vs EPA filter by site
library(dplyr)
library(ggplot2)
library(patchwork)

# Build plotting table
figB_table <- epa_cinc_compare %>%
  filter(!is.na(pm25_filter), !is.na(pm25_cinc)) %>%
  mutate(Site = site_number) %>%
  select(
    Site,
    EPA_FRM = pm25_filter,
    CINC_PM25 = pm25_cinc
  ) %>%
  left_join(
    epa_cinc_pairs2 %>%
      mutate(site_number = sub("^site_", "", site_number)),
    by = c("Site" = "site_number")
  )

# Common axis limits
panel_min <- 0
panel_max <- 45
lims <- c(panel_min, panel_max)

# Function to make one panel exactly in template style
make_panel <- function(site_id) {
  
  df <- figB_table %>%
    filter(Site == site_id)
  
  dev_id <- unique(df$nearest_device)[1]
  
  fit <- lm(CINC_PM25 ~ EPA_FRM, data = df)
  slope <- coef(fit)[2]
  intercept <- coef(fit)[1]
  r2 <- summary(fit)$r.squared
  
  eq_label <- sprintf("y = %.2fx %+ .2f\nR² = %.2f", slope, intercept, r2)
  
  ggplot(df, aes(x = EPA_FRM, y = CINC_PM25)) +
    geom_point(alpha = 0.8, size = 1.9) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 1.0, color = "steelblue") +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
    annotate(
      "text",
      x = panel_min + (panel_max - panel_min) * 0.04,
      y = panel_max - (panel_max - panel_min) * 0.07,
      label = eq_label,
      hjust = 0,
      vjust = 1,
      size = 4
    ) +
    coord_equal(xlim = lims, ylim = lims, expand = FALSE) +
    labs(
      title = paste0(site_id),
      x = paste0("Site ", site_id, " (µg/m³)"),
      y = paste0("Device ", dev_id, " (µg/m³)")
    ) +
    scale_x_continuous(breaks = seq(0, 45, 10)) +
    scale_y_continuous(breaks = seq(0, 45, 10)) +
    theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5, size = 13),
      axis.title = element_text(size = 11),
      axis.text = element_text(size = 10, color = "black"),
      panel.grid.minor = element_blank()
    )
}

# Build all six panels
p1 <- make_panel("0034")
p2 <- make_panel("0038")
p3 <- make_panel("0045")
p4 <- make_panel("0060")
p5 <- make_panel("0065")
p6 <- make_panel("0073")

# Combine like the template
(p1 + p2 + p3) / (p4 + p5 + p6) +
  plot_annotation(
    title = "CLEANinCLE vs EPA FRM PM2.5 - by Site",
    theme = theme(
      plot.title = element_text(face = "bold", hjust = 0.5, size = 16)
    )
  )
```

# CLEANinCLE vs EPA optical (daily averages)
```{r fig_cinc_vs_epa_optical, fig.width=11, fig.height=7}
# EPA optical daily data
epa_optical_daily <- epa_fem_daily %>%
  select(date, site_number, pm25_daily) %>%
  rename(pm25_epa_optical = pm25_daily) %>%
  mutate(site_number = as.character(site_number))

epa_cinc_pairs2 <- epa_cinc_pairs2 %>%
  mutate(site_number = sub("^site_", "", site_number))

# EPA optical with CinC daily
epa_cinc_optical_compare <- epa_cinc_pairs2 %>%
  left_join(epa_optical_daily, by = "site_number") %>%
  left_join(
    cinc_daily %>% mutate(device_number = as.character(device_number)),
    by = c("nearest_device" = "device_number", "date" = "date")
  )

# Filter matched rows
epa_cinc_optical_plot <- epa_cinc_optical_compare %>%
  filter(!is.na(pm25_epa_optical), !is.na(pm25_cinc))

# Plot
figC_table <- epa_cinc_optical_compare %>%
  filter(!is.na(pm25_epa_optical), !is.na(pm25_cinc)) %>%
  mutate(Site = site_number) %>%
  select(
    Site,
    EPA_OPTICAL = pm25_epa_optical,
    CINC_PM25 = pm25_cinc
  ) %>%
  left_join(
    epa_cinc_pairs2 %>%
      mutate(site_number = sub("^site_", "", site_number)),
    by = c("Site" = "site_number")
  )

# same style as your FRM plot
make_panel_optical <- function(site_id) {

  df <- figC_table %>%
    filter(Site == site_id)

  dev_id <- unique(df$nearest_device)[1]

  fit <- lm(CINC_PM25 ~ EPA_OPTICAL, data = df)

  slope <- coef(fit)[2]
  intercept <- coef(fit)[1]
  r2 <- summary(fit)$r.squared

  eq_label <- sprintf("y = %.2fx %+ .2f\nR² = %.2f", slope, intercept, r2)

  ggplot(df, aes(EPA_OPTICAL, CINC_PM25)) +
    geom_point(alpha = 0.8, size = 1.9) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 1.0, color = "steelblue") +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
    annotate(
      "text",
      x = min(df$EPA_OPTICAL) + 0.5,
      y = max(df$CINC_PM25) * 0.95,
      label = eq_label,
      hjust = 0,
      vjust = 1,
      size = 4
    ) +
    labs(
      title = site_id,
      x = paste0("Site ", site_id, " (µg/m³)"),
      y = paste0("Device ", dev_id, " (µg/m³)")
    ) +
    scale_x_continuous(breaks = scales::pretty_breaks(5)) +
    scale_y_continuous(breaks = scales::pretty_breaks(5)) +
    theme_minimal(base_size = 14) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5),
      panel.grid.minor = element_blank()
    )
}
p1 <- make_panel_optical("0034")
p2 <- make_panel_optical("0038")
p3 <- make_panel_optical("0045")
p4 <- make_panel_optical("0060")
p5 <- make_panel_optical("0065")
p6 <- make_panel_optical("0073")

(p1 + p2 + p3) / (p4 + p5 + p6) +
  plot_annotation(
    title = "CLEANinCLE vs EPA Optical PM2.5 - by Site",
    theme = theme(
      plot.title = element_text(face = "bold", hjust = 0.5, size = 16)
    )
  )
```


```{r fig_epa_optical_vs_frm, fig.width=11, fig.height=7}
figD_table <- epa_compare %>%
  filter(!is.na(pm25_filter), !is.na(pm25_daily)) %>%
  mutate(Site = site_number) %>%
  select(
    Site,
    EPA_FRM = pm25_filter,
    EPA_OPTICAL = pm25_daily
  )

make_panel_epa <- function(site_id) {

  df <- figD_table %>%
    filter(Site == site_id)

  fit <- lm(EPA_OPTICAL ~ EPA_FRM, data = df)

  slope <- coef(fit)[2]
  intercept <- coef(fit)[1]
  r2 <- summary(fit)$r.squared

  eq_label <- sprintf("y = %.2fx %+ .2f\nR² = %.2f", slope, intercept, r2)

  ggplot(df, aes(EPA_FRM, EPA_OPTICAL)) +
    geom_point(alpha = 0.8, size = 1.9) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 1.0, color = "steelblue") +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
    annotate(
      "text",
      x = min(df$EPA_FRM) + 0.5,
      y = max(df$EPA_OPTICAL) * 0.95,
      label = eq_label,
      hjust = 0,
      vjust = 1,
      size = 4
    ) +
    labs(
      title = site_id,
      x = paste0("EPA FRM — Site ", site_id, " (µg/m³)"),
      y = paste0("EPA Optical — Site ", site_id, " (µg/m³)")
    ) +
    scale_x_continuous(breaks = scales::pretty_breaks(5)) +
    scale_y_continuous(breaks = scales::pretty_breaks(5)) +
    theme_minimal(base_size = 14) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5),
      panel.grid.minor = element_blank()
    )
}

(p1 + p2 + p3) / (p4 + p5 + p6) +
  plot_annotation(
    title = "EPA Optical vs EPA FRM PM2.5 — by Site",
    theme = theme(
      plot.title = element_text(face = "bold", hjust = 0.5, size = 16)
    )
  )

```

```{r}
# EPA optical daily table
epa_optical_daily_wide <- epa_fem_daily %>%
  mutate(site = paste0("site_", site_number)) %>%
  select(date, site, pm25_daily) %>%
  pivot_wider(
    names_from = site,
    values_from = pm25_daily
  )

#Correlation matrix
epa_optical_cor_matrix <- cor(
  epa_optical_daily_wide %>% select(-date),
  use = "pairwise.complete.obs"
)

library(reshape2)
library(ggplot2)

epa_optical_cor_long <- melt(epa_optical_cor_matrix)

ggplot(epa_optical_cor_long, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_viridis_c(name = "Correlation") +
  labs(
    title = "Correlation Between EPA Optical Sites (Daily PM2.5)",
    x = "EPA Optical Site",
    y = "EPA Optical Site"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  )

```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
