---
title: "Untitled"
author: "CLEANINCLE API"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# CLEANinCLE API

```{r}
# Purpose / Description:
# Download the full CLEANinCLE PM2.5 dataset from the project API for the study
# period spanning November 1, 2024 through December 31, 2025. This chunk handles
# paginated API requests, collects all returned records across pages, converts the
# nested JSON structure into a flat dataframe suitable for analysis, and saves the
# raw downloaded data in both CSV and RDS formats. The resulting raw table serves
# as the starting point for all later CLEANinCLE workflows, including site mapping,
# site-level table construction, data cleaning, daily averaging, time series plots,
# site-to-site comparisons, and CinC–EPA matched analyses.

# Upstream workflow:
# - Requires valid CLEANinCLE API URL and API key
# - Requires the following packages:
#   'httr2', 'jsonlite', 'dplyr', 'tidyr', 'readr', and 'lubridate'
# - Assumes the API returns paginated records with an optional 'next_token'
# - Run before any CLEANinCLE-specific data preparation, cleaning, aggregation,
#   plotting, or matched-site comparison workflows

library(httr2)
library(jsonlite)
library(dplyr)
library(tidyr)
library(readr)
library(lubridate)


# 1. API settings

api_url <- "https://data.cleanincle.issacs.dev"
api_key <- "2KYc26u6sn6bPAaZHJx5R3DSzQL8Y6sV38hy0c3t"

params <- list(
  start_date = "2024-11-01T00:00:00Z",
  end_date   = "2026-01-01T00:00:00Z"
)

# 2. Download all pages
# -----------------------------
all_records <- list()
current_params <- params

message("Starting CLEANinCLE API download...")

repeat {
  resp <- request(api_url) |>
    req_headers(`x-api-key` = api_key) |>
    req_url_query(!!!current_params) |>
    req_retry(max_tries = 3) |>
    req_perform()

  body <- resp_body_json(resp, simplifyVector = FALSE)

  all_records <- c(all_records, body$records)

  if (is.null(body$next_token)) break

  current_params$next_token <- body$next_token
}

# -----------------------------
# 3. Convert to dataframe
# -----------------------------
df_raw <- bind_rows(all_records)

df_raw <- df_raw %>%
  unnest_wider(where(is.list), names_sep = "_")

# -----------------------------
# 4. Save raw downloaded file
# -----------------------------
write_csv(df_raw, "CLEANinCLE_raw_2024-11_to_2025-12.csv")
saveRDS(df_raw, "CLEANinCLE_raw_2024-11_to_2025-12.rds")

message("Download complete.")
dim(df_raw)
names(df_raw)
```

# PM2.5-only CLEANinCLE dataset
```{r}

# Purpose / Description:
# Create a PM2.5-only CLEANinCLE dataset from the raw API download by extracting
# the variables needed for later analysis and plotting. This chunk converts the
# raw timestamp into a proper UTC datetime, creates a corresponding calendar date
# column, ensures PM2.5 values are numeric, keeps only the key columns related to
# sensor identity, location, time, and PM2.5 concentration, removes rows with
# missing PM2.5 values, and sorts the resulting table by device and time. The
# output serves as the base PM2.5 dataset for later site mapping, cleaned site-table
# construction, time series plots, daily averaging, site-to-site comparisons, and
# CinC–EPA matched analyses.

# Upstream workflow:
# - Requires 'df_raw' from the CLEANinCLE API download workflow
# - Assumes the following columns exist and are correctly formatted or convertible:
#   'time', 'pm_2.5', 'device_eui', 'location_name', 'latitude', 'longitude',
#   'device_number', and 'device_status'
# - Run after raw API data have been downloaded and flattened
# - Run before site mapping, site-table construction, data cleaning, aggregation,
#   and all downstream CLEANinCLE PM2.5 analyses

library(dplyr)
library(lubridate)
library(readr)

cinc_pm25 <- df_raw %>%
  mutate(
    time = ymd_hms(time, tz = "UTC"),
    date = as.Date(time),
    pm_2.5 = as.numeric(pm_2.5)
  ) %>%
  select(
    device_eui,
    location_name,
    latitude,
    longitude,
    device_number,
    device_status,
    time,
    date,
    pm_2.5
  ) %>%
  filter(!is.na(pm_2.5)) %>%
  arrange(device_number, time)

# Save PM2.5-only dataset
write_csv(cinc_pm25, "CLEANinCLE_PM25_2024-11_to_2025-12.csv")
saveRDS(cinc_pm25, "CLEANinCLE_PM25_2024-11_to_2025-12.rds")

# Checks
dim(cinc_pm25)
names(cinc_pm25)
head(cinc_pm25)
range(cinc_pm25$time, na.rm = TRUE)
```
# CinC Exploratory Data Analysis 
```{r}
# Purpose / Description:
# Perform exploratory data analysis (EDA) on the CLEANinCLE PM2.5 dataset to
# understand the structure, coverage, and quality of the sensor data. This chunk
# summarizes device availability, observation counts per device, missing values,
# and basic PM2.5 statistics to identify potential issues such as uneven data
# coverage, outliers, or unrealistic measurements. The results guide subsequent
# cleaning, filtering, and aggregation steps used in time series analysis,
# site-to-site comparisons, and CinC–EPA matched analyses.

# Upstream workflow:
# - Requires 'cinc_pm25' from the PM2.5 dataset preparation workflow
# - Assumes the following columns exist and are correctly formatted:
#   'device_number', 'location_name', 'time', 'date', and 'pm_2.5'
# - Run after PM2.5 dataset creation and before data cleaning, trimming,
#   and aggregation into daily averages

library(dplyr)
library(ggplot2)


# 1. List all CinC devices
cinc_devices <- cinc_pm25 %>%
  distinct(device_number, location_name) %>%
  arrange(device_number)

cinc_devices


# 2. Number of observations per device
cinc_counts <- cinc_pm25 %>%
  count(device_number, location_name, sort = TRUE)

cinc_counts


# 3. Missing PM2.5 values
missing_pm25 <- sum(is.na(cinc_pm25$pm_2.5))
missing_pm25


# 4. Summary statistics of PM2.5
pm25_summary <- summary(cinc_pm25$pm_2.5)
pm25_summary


# 5. Distribution of PM2.5 (Histogram)
cinc_hist_plot <- ggplot(cinc_pm25, aes(x = pm_2.5)) +
  geom_histogram(bins = 50) +
  labs(
    title = "Distribution of CLEANinCLE PM2.5",
    x = "PM2.5 (µg/m³)",
    y = "Frequency"
  ) +
  theme_minimal()

print(cinc_hist_plot)


# 6. Boxplot (Outlier detection)
cinc_box_plot <- ggplot(cinc_pm25, aes(y = pm_2.5)) +
  geom_boxplot() +
  labs(
    title = "Boxplot of CLEANinCLE PM2.5",
    y = "PM2.5 (µg/m³)"
  ) +
  theme_minimal()

print(cinc_box_plot)


# 7. Time coverage (overall)
time_range <- range(cinc_pm25$time, na.rm = TRUE)
time_range


# 8. Observations per day
daily_counts <- cinc_pm25 %>%
  count(date)

head(daily_counts)


# 9. Quick time series (all data combined)
cinc_time_plot <- ggplot(cinc_pm25, aes(x = time, y = pm_2.5)) +
  geom_point(alpha = 0.2, size = 0.5) +
  labs(
    title = "CLEANinCLE PM2.5 Time Series (All Devices)",
    x = "Time",
    y = "PM2.5 (µg/m³)"
  ) +
  theme_minimal()

print(cinc_time_plot)

```

# CinC provide a unique device name for each site
```{r}
# Purpose / Description:
# Assign a unique site identifier to each CLEANinCLE device deployment based on
# changes in geographic location over time. Because a single CLEANinCLE device
# can be moved and redeployed at different coordinates, this chunk detects changes
# in rounded latitude and longitude within each device’s time series and uses those
# changes to define distinct site deployments. A unique Site_ID is then created in
# the form "device_number.site_number" (for example, 23.1, 23.2, etc.). The chunk
# also summarizes these unique deployments into a site-level lookup table containing
# coordinates, deployment start and end times, and the number of observations. This
# table is used later for site mapping, matching multiple sensors to combined sites,
# building the final site-level PM2.5 table, and supporting CLEANinCLE–EPA comparison
# workflows.

# Upstream workflow:
# - Requires 'cinc_pm25' from the PM2.5-only CLEANinCLE dataset workflow
# - Assumes the following columns exist and are correctly formatted or convertible:
#   'device_number', 'location_name', 'latitude', 'longitude', and 'time'
# - Requires timestamps that can be converted to POSIXct and coordinates that can
#   be converted to numeric
# - Run after creation of the PM2.5-only dataset and before combined-site mapping,
#   final site-table construction, data cleaning, and site-level analyses

library(dplyr)
library(lubridate)
library(readr)


# 1. Create site-level deployment IDs for each device
cinc_sites <- cinc_pm25 %>%
  mutate(
    latitude = as.numeric(latitude),
    longitude = as.numeric(longitude),
    datetime = as.POSIXct(time),
    lat_round = round(latitude, 5),
    lon_round = round(longitude, 5)
  ) %>%
  filter(
    !is.na(lat_round),
    !is.na(lon_round),
    lat_round != 0,
    lon_round != 0
  ) %>%
  arrange(device_number, datetime) %>%
  group_by(device_number) %>%
  mutate(
    loc_change = (lat_round != lag(lat_round, default = first(lat_round))) |
                 (lon_round != lag(lon_round, default = first(lon_round))),
    site_number = cumsum(loc_change) + 1,
    Site_ID = paste0(device_number, ".", site_number)
  ) %>%
  ungroup() %>%
  select(-loc_change, -site_number, -lat_round, -lon_round)


# 2. Summarize unique device deployments into a lookup table
cinc_unique_site_table <- cinc_sites %>%
  group_by(Site_ID, device_number, location_name) %>%
  summarise(
    latitude = first(latitude),
    longitude = first(longitude),
    start_time = min(datetime, na.rm = TRUE),
    end_time = max(datetime, na.rm = TRUE),
    n_obs = n(),
    .groups = "drop"
  ) %>%
  arrange(device_number, start_time)


# 3. Inspect unique site table
cinc_unique_site_table


# 4. Save unique site table
write_csv(cinc_unique_site_table, "CLEANinCLE_unique_site_table.csv")
saveRDS(cinc_unique_site_table, "CLEANinCLE_unique_site_table.rds")
```

```{r}
# Sensor coverage summary (start, end, number of observations)
cinc_pm25 %>%
  mutate(month = lubridate::floor_date(time, "month")) %>%
  count(device_number, location_name, month) %>%
  arrange(device_number, month)

library(gt)

cinc_summary <- cinc_pm25 %>%
  group_by(device_number, location_name) %>%
  summarise(
    start_time = min(time, na.rm = TRUE),
    end_time = max(time, na.rm = TRUE),
    n_obs = n(),
    .groups = "drop"
  ) %>%
  arrange(device_number)

gt_table <- gt(cinc_summary)

gtsave(gt_table, "CLEANinCLE_sensor_summary.pdf")

# Monthly coverage for all sensors (shows how many measurements each sensor has per month)
cinc_pm25 %>%
  mutate(month = lubridate::floor_date(time, "month")) %>%
  count(device_number, location_name, month) %>%
  arrange(device_number, month)

cinc_pm25 %>%
  mutate(
    month = lubridate::floor_date(time, "month"),
    device_number = factor(device_number)
  ) %>%
  distinct(device_number, location_name, month) %>%
  ggplot(aes(x = month, y = device_number)) +
  geom_point(size = 2) +
  scale_x_datetime(
    limits = as.POSIXct(c("2024-11-01", "2025-12-31")),
    date_breaks = "1 month",
    date_labels = "%b %Y"
  ) +
  labs(
    title = "CLEANinCLE Sensor Data Coverage",
    x = "Month",
    y = "Sensor"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# CinC Time Series Plot 0f PM2.5 for all devices
```{r}
# Purpose / Description:
# Generate time series plots of PM2.5 for each CLEANinCLE device to visualize
# temporal patterns, data coverage, and variability across the full study period.
# This chunk loops through all device numbers in the PM2.5 dataset, extracts the
# corresponding time series for each device, and creates individual scatter plots
# of PM2.5 versus time. The plots display the full study window from November 1,
# 2024 through December 31, 2025 with monthly date labels. Each plot is saved as a
# high-resolution PDF (and optional PNG) for inspection, quality control, and
# inclusion in reports or appendices.

# Upstream workflow:
# - Requires 'cinc_pm25' from the PM2.5-only CLEANinCLE dataset workflow
# - Assumes the following columns exist and are correctly formatted:
#   'device_number', 'location_name', 'time', and 'pm_2.5'
# - Requires 'time' to be in POSIXct format
# - Run after PM2.5 dataset creation and before data cleaning, trimming,
#   and site-level aggregation

library(ggplot2)
library(dplyr)


# 1. Get all device IDs
device_ids <- sort(unique(cinc_pm25$device_number))


# 2. Create output folders
dir.create("CinC_PM25_timeseries_plots_pdf", showWarnings = FALSE)
dir.create("CinC_PM25_timeseries_plots_png", showWarnings = FALSE)


# 3. Define full monthly sequence
full_dates <- seq(
  as.POSIXct("2024-11-01"),
  as.POSIXct("2025-12-31"),
  by = "1 month"
)


# 4. Loop through each device
for (dev in device_ids) {
  
  sensor_data <- cinc_pm25 %>% 
    filter(device_number == dev)
  
  loc_name <- unique(sensor_data$location_name)[1]
  
  cinc_ts_plot <- ggplot(sensor_data, aes(x = time, y = pm_2.5)) +
    
    geom_point(alpha = 0.4, size = 0.7) +
    
    scale_x_datetime(
      limits = as.POSIXct(c("2024-11-01", "2025-12-31")),
      breaks = full_dates,
      date_labels = "%b %Y"
    ) +
    
    labs(
      title = paste("PM2.5 Time Series - Device", dev, "-", loc_name),
      x = "Time",
      y = "PM2.5 (µg/m³)"
    ) +
    
    theme_minimal(base_size = 11) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5),
      axis.text.x = element_text(angle = 60, hjust = 1, size = 8),
      panel.grid.minor = element_blank()
    )
  
  
  # 5. Print plot
  print(cinc_ts_plot)
  
  
  # 6. Save HIGH-RES PDF (main output)

  ggsave(
    filename = paste0("CinC_PM25_timeseries_plots_pdf/CinC_sensor_", dev, ".pdf"),
    plot = cinc_ts_plot,
    width = 10,
    height = 5,
    units = "in"
  )
  
}
```

## CLEANinCLE PM2.5 Time Series by Site Deployment
```{r}
# Purpose / Description:
# Generate PM2.5 time series plots for each unique CLEANinCLE site deployment
# identified in the 'cinc_sites' table. Unlike the earlier device-level time series
# plots, this chunk uses the deployment-specific Site_ID (for example, 23.1, 23.2,
# etc.) so that periods when a device was moved to a new location are treated as
# separate site records. These plots are used to visually inspect temporal behavior,
# data coverage, and suspicious patterns within each deployment, including periods
# later flagged for screening or trimming (for example, extreme values above 250
# µg/m³, flat-line artifacts, or suspicious low-variability intervals). Each site
# deployment is plotted individually across the full study period from November 1,
# 2024 through December 31, 2025 and saved as a high-resolution PDF (and optional
# PNG) for review, cleaning decisions, and appendix/thesis use.

# Upstream workflow:
# - Requires 'cinc_sites' from the unique CLEANinCLE site deployment workflow
# - Assumes the following columns exist and are correctly formatted:
#   'Site_ID', 'device_number', 'location_name', 'datetime', and 'pm_2.5'
# - Requires 'datetime' to be in POSIXct format
# - Run after creation of deployment-based Site_ID values and before final manual
#   trimming, cleaned site-table generation, and downstream CinC–EPA comparisons

library(ggplot2)
library(dplyr)


# 1. Get all unique site deployment IDs
site_ids <- sort(unique(cinc_sites$Site_ID))


# 2. Create output folders
dir.create("CinC_site_timeseries_plots_pdf", showWarnings = FALSE)
dir.create("CinC_site_timeseries_plots_png", showWarnings = FALSE)


# 3. Define full monthly date sequence
full_dates <- seq(
  as.POSIXct("2024-11-01"),
  as.POSIXct("2025-12-31"),
  by = "1 month"
)


# 4. Loop through each unique site deployment
for (site in site_ids) {
  
  sensor_data <- cinc_sites %>%
    filter(Site_ID == site)
  
  dev <- unique(sensor_data$device_number)[1]
  loc_name <- unique(sensor_data$location_name)[1]
  
  cinc_site_plot <- ggplot(sensor_data, aes(x = datetime, y = pm_2.5)) +
    geom_point(alpha = 0.4, size = 0.7) +
    scale_x_datetime(
      limits = as.POSIXct(c("2024-11-01", "2025-12-31")),
      breaks = full_dates,
      date_labels = "%b %Y"
    ) +
    labs(
      title = paste("PM2.5 Time Series - Site", site, "(Device", dev, "-", loc_name, ")"),
      x = "Time",
      y = "PM2.5 (µg/m³)"
    ) +
    theme_minimal(base_size = 11) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5),
      axis.text.x = element_text(angle = 60, hjust = 1, size = 8),
      panel.grid.minor = element_blank()
    )
  
  
  # 5. Print plot
  print(cinc_site_plot)
  

  # 6. Save HIGH-RES PDF
  ggsave(
    filename = paste0("CinC_site_timeseries_plots_pdf/CinC_site_", site, ".pdf"),
    plot = cinc_site_plot,
    width = 10,
    height = 5,
    units = "in"
  )
}


```


# Create Combined CLEANinCLE Site Table
```{r}
# Purpose / Description:
# Construct the final CLEANinCLE site-level PM2.5 table by mapping individual
# deployment-specific sensor IDs (for example, 5.1, 19.3, 24.2) to combined site
# identifiers that represent a single physical monitoring location through time.
# This chunk defines a manual lookup table linking each sensor deployment to its
# corresponding combined site, preserves the original sensor identity, and then
# joins that lookup to the deployment-level CLEANinCLE table. The resulting final
# table contains one row per PM2.5 observation with both the original Sensor ID
# and the combined Site ID, along with coordinates, date, time, and PM2.5 values.
# This is the main site-based dataset used for later cleaning, trimming, daily
# averaging, time series plots by site, site-to-site comparisons, and CinC–EPA
# matched analyses.

# Upstream workflow:
# - Requires 'cinc_sites' from the unique CLEANinCLE site deployment workflow
# - Assumes the following columns exist and are correctly formatted:
#   'Site_ID', 'latitude', 'longitude', 'datetime', and 'pm_2.5'
# - Uses a manually defined sensor-to-site lookup table based on deployment review
#   and decisions about which sensor deployments belong to the same physical site
# - Run after creation of deployment-specific Site_ID values and before final
#   data cleaning, site-level time series plots, and CinC–EPA comparison workflows

library(readr)
library(dplyr)
library(stringr)


# 1. Define lookup table:
#    deployment-specific Sensor ID -> combined Site ID
site_lookup_clean <- tibble(
  sensor_id = c(
    "1.1","1.2","2.1","3.1","4.1","5.1","5.2","6.1","7.1","7.2",
    "8.1","8.2","9.1","9.2","11.1","11.2","11.3","12.1","13.1","14.1",
    "15.1","16.1","17.1","17.2","18.1","18.2","19.1","19.2","19.3","19.4",
    "20.1","20.2","21.1","22.1","22.2","23.1","24.1","24.2","25.1","26.1",
    "27.1","27.2","27.3","28.1","29.1","30.1","30.2","31.1","31.2","32.1",
    "32.2","33.1","33.2","34.1","34.2","35.1","35.2"
  ),
  site_id = c(
    "1.1","24.1_1.2","2.1","3.1","7.1_4.1","5.1_19.3_24.2","12.1_5.2","6.1","7.1_4.1","7.2",
    "8.1_34.2","8.2","9.1_17.2","9.2","11.1_32.1_18.2","11.2","11.3","12.1_5.2","13.1_33.2","14.1",
    "15.1","34.1_16.1","30.1_17.1","9.1_17.2","18.1_26.1","11.1_32.1_18.2","19.1","19.2","5.1_19.3_24.2","19.4",
    "20.1","20.2","21.1","22.1","22.1","23.1","24.1_1.2","5.1_19.3_24.2","25.1","18.1_26.1",
    "27.1","27.2","27.3","28.1","29.1","30.1_17.1","30.2","31.1","31.2","11.1_32.1_18.2",
    "32.2","33.1","13.1_33.2","34.1_16.1","8.1_34.2","35.1","35.2"
  )
)

site_lookup_clean


# 2. Define sensor ordering for the final table

sensor_order <- c(
  "1.1","1.2","2.1","3.1","4.1","5.1","5.2","6.1","7.1","7.2",
  "8.1","8.2","9.1","9.2","11.1","11.2","11.3","12.1","13.1","14.1",
  "15.1","16.1","17.1","17.2","18.1","18.2","19.1","19.2","19.3","19.4",
  "20.1","20.2","21.1","22.1","22.2","23.1","24.1","24.2","25.1","26.1",
  "27.1","27.2","27.3","28.1","29.1","30.1","30.2","31.1","31.2","32.1",
  "32.2","33.1","33.2","34.1","34.2","35.1","35.2"
)


# 3. Join deployment table to combined-site lookup
cinc_site_table_final <- cinc_sites %>%
  mutate(sensor_id = as.character(Site_ID)) %>%
  left_join(site_lookup_clean, by = "sensor_id") %>%
  filter(!is.na(site_id)) %>%
  transmute(
    `Sensor ID` = sensor_id,
    `Site ID`   = site_id,
    Lat         = as.numeric(latitude),
    Long        = as.numeric(longitude),
    Date        = as.Date(datetime),
    Time        = format(datetime, "%H:%M:%S"),
    `PM2.5`     = pm_2.5
  ) %>%
  mutate(`Sensor ID` = factor(`Sensor ID`, levels = sensor_order)) %>%
  arrange(`Sensor ID`, Date, Time)

# -------------------------
# 4. Inspect final site-level table
# -------------------------
dim(cinc_site_table_final)
names(cinc_site_table_final)
head(cinc_site_table_final)

# -------------------------
# 5. Save final site-level table
# -------------------------
#write_csv(cinc_site_table_final, "CLEANinCLE_final_site_table.csv")
#saveRDS(cinc_site_table_final, "CLEANinCLE_final_site_table.rds")

```

# CLEANinCLE Daily Sensor Coverage
```{r}

# Purpose / Description:
# Build a site-level CLEANinCLE coverage table and visualize daily sensor coverage
# across the combined Site_ID locations used in the final analysis. This chunk
# reconstructs the full site-level PM2.5 table by joining deployment-specific
# sensor IDs to the manually defined combined-site lookup table, verifies key
# site-mapping corrections (including the separation of 27.1 and 9.2 and the
# inclusion of 22.1), and then summarizes the presence of measurements by site
# and day. The resulting dot plot shows whether each combined site had at least
# one observation on a given date, making it useful for checking temporal data
# continuity, identifying deployment gaps, and confirming the site structure
# before cleaning, trimming, and site-to-site comparison analyses.

# Upstream workflow:
# - Requires 'cinc_sites' from the unique CLEANinCLE site deployment workflow
# - Requires 'site_lookup_clean' from the manual sensor-to-combined-site mapping step
# - Assumes the following columns exist and are correctly formatted:
#   'Site_ID', 'latitude', 'longitude', 'datetime', and 'pm_2.5'
# - Uses the manually defined combined Site_ID structure for later CLEANinCLE
#   time series plots, cleaned site-table construction, and CinC–EPA comparison work
# - Run after the final sensor-to-site lookup table is defined and before final
#   site-level cleaning and downstream analyses

library(dplyr)
library(ggplot2)
library(tibble)
library(readr)


# 1. Rebuild full site-level table from corrected lookup
cinc_site_table_full <- cinc_sites %>%
  mutate(sensor_id = as.character(Site_ID)) %>%
  left_join(site_lookup_clean, by = "sensor_id") %>%
  filter(!is.na(site_id)) %>%
  mutate(
    latitude = as.numeric(latitude),
    longitude = as.numeric(longitude),
    date = as.Date(datetime),
    time_only = format(datetime, "%H:%M:%S"),
    pm25 = pm_2.5
  ) %>%
  select(
    sensor_id,
    site_id,
    latitude,
    longitude,
    date,
    time = time_only,
    pm25
  ) %>%
  arrange(latitude, longitude, date, time)


# 2. Confirm key site mapping corrections
distinct(cinc_site_table_full, sensor_id, site_id) %>%
  filter(sensor_id %in% c("27.1", "9.2")) %>%
  print(n = Inf)

distinct(cinc_site_table_full, sensor_id, site_id) %>%
  filter(site_id == "22.1") %>%
  print(n = Inf)


# 3. Create daily coverage table
#    (one row per site, sensor, and date with at least one observation)
coverage_site_sensor <- cinc_site_table_full %>%
  group_by(site_id, sensor_id, date) %>%
  summarise(.groups = "drop")


# 4. Define ordered combined site list
site_order <- c(
  "1.1","2.1","3.1","5.1_19.3_24.2","7.1_4.1","7.2","8.1_34.2",
  "9.1_17.2","9.2",
  "11.1_32.1_18.2","12.1_5.2","13.1_33.2","14.1","15.1",
  "18.1_26.1","19.1","19.2","20.1","20.2","21.1","22.1","23.1",
  "24.1_1.2","25.1","27.1","28.1","29.1","30.1_17.1","31.2",
  "32.2","33.1","34.1_16.1","35.1","35.2"
)

# Keep only site IDs that actually exist in the data
valid_sites <- intersect(site_order, unique(coverage_site_sensor$site_id))

coverage_site_sensor <- coverage_site_sensor %>%
  filter(site_id %in% valid_sites) %>%
  mutate(site_id = factor(site_id, levels = rev(valid_sites)))


# 5. Plot daily coverage by combined site
cinc_coverage_plot <- ggplot(coverage_site_sensor, aes(x = date, y = site_id)) +
  geom_point(size = 0.45, color = "black") +
  labs(
    title = "CLEANinCLE Daily Sensor Coverage by Site ID",
    x = "Date",
    y = "Site ID"
  ) +
  scale_x_date(
    limits = c(as.Date("2024-11-01"), as.Date("2025-12-31")),
    date_breaks = "1 month",
    date_labels = "%b %Y"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 7),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank()
  )

print(cinc_coverage_plot)


# 6. Save high-resolution coverage plot
ggsave(
  filename = "cinc_daily_sensor_coverage_by_site.pdf",
  plot = cinc_coverage_plot,
  width = 12,
  height = 8,
  units = "in"
)

ggsave(
  filename = "cinc_daily_sensor_coverage_by_site.png",
  plot = cinc_coverage_plot,
  width = 12,
  height = 8,
  units = "in",
  dpi = 300
)


# 7. Save site-level table for reuse
cinc_site_table <- cinc_site_table_full

write_csv(cinc_site_table, "cinc_site_table.csv")
saveRDS(cinc_site_table, "cinc_site_table.rds")
```
# CLEANING WORKFLOW FOR CinC FINAL SITE TABLE
```{r}
# Purpose:
# Clean and standardize the CinC site dataset by removing unreliable or invalid PM2.5 data.
# This includes filtering extreme values, removing problematic sites, trimming known bad time periods,
# and ensuring physically realistic PM2.5 values. Also creates a log (cinc_removed) of all excluded data.

# Upstream workflow:
# - Requires 'cinc_site_table_final' to be fully processed and loaded
# - Assumes columns: 'Site ID', 'Date', and 'PM2.5' exist and are correctly formatted
# - Should be run AFTER initial data compilation/merging and BEFORE any plotting or analysis

# CLEANING WORKFLOW FOR CinC FINAL SITE TABLE
library(dplyr)
library(tibble)

# 1. Keep original dataset unchanged
cinc_original <- cinc_site_table_final

# 2. Create empty table to store removed data
cinc_removed <- cinc_original %>%
  slice(0) %>%
  mutate(removal_reason = character())

# 3. Remove obvious extreme values (>250)
bad_high <- cinc_original %>%
  filter(`PM2.5` > 250) %>%
  mutate(removal_reason = "PM2.5 > 250")

cinc_removed <- bind_rows(cinc_removed, bad_high)

cinc_no_extreme <- cinc_original %>%
  filter(`PM2.5` <= 250 | is.na(`PM2.5`))

# 4. Remove site 6.1 completely
bad_6 <- cinc_no_extreme %>%
  filter(`Site ID` == "6.1") %>%
  mutate(removal_reason = "Remove site 6.1 completely")

cinc_removed <- bind_rows(cinc_removed, bad_6)

cinc_clean <- cinc_no_extreme %>%
  filter(`Site ID` != "6.1")

# 5. Trim site 18.2 (late May onward)
bad_18 <- cinc_clean %>%
  filter(`Site ID` == "18.2" & Date >= as.Date("2025-05-20")) %>%
  mutate(removal_reason = "Trim 18.2 from 2025-05-20 onward")

cinc_removed <- bind_rows(cinc_removed, bad_18)

cinc_clean <- cinc_clean %>%
  filter(!(`Site ID` == "18.2" & Date >= as.Date("2025-05-20")))

# 6. Trim site 27.1 (late period)
bad_27 <- cinc_clean %>%
  filter(`Site ID` == "27.1" & Date >= as.Date("2025-07-01")) %>%
  mutate(removal_reason = "Trim 27.1 from 2025-07-01 onward")

cinc_removed <- bind_rows(cinc_removed, bad_27)

cinc_clean <- cinc_clean %>%
  filter(!(`Site ID` == "27.1" & Date >= as.Date("2025-07-01")))

# 7. Remove site 27.3 completely
bad_273 <- cinc_clean %>%
  filter(`Site ID` == "27.3") %>%
  mutate(removal_reason = "Remove site 27.3 completely")

cinc_removed <- bind_rows(cinc_removed, bad_273)

cinc_clean <- cinc_clean %>%
  filter(`Site ID` != "27.3")

# 8. Trim site 13.1 (late September onward)
bad_13 <- cinc_clean %>%
  filter(`Site ID` == "13.1" & Date >= as.Date("2025-09-20")) %>%
  mutate(removal_reason = "Trim 13.1 from 2025-09-20 onward")

cinc_removed <- bind_rows(cinc_removed, bad_13)

cinc_clean <- cinc_clean %>%
  filter(!(`Site ID` == "13.1" & Date >= as.Date("2025-09-20")))

# 9. Remove impossible negative values, if any
bad_negative <- cinc_clean %>%
  filter(`PM2.5` < 0) %>%
  mutate(removal_reason = "Negative PM2.5 value")

cinc_removed <- bind_rows(cinc_removed, bad_negative)

cinc_clean <- cinc_clean %>%
  filter(`PM2.5` >= 0 | is.na(`PM2.5`))

# 10. Check results
list(
  original_rows   = nrow(cinc_original),
  removed_rows    = nrow(cinc_removed),
  no_extreme_rows = nrow(cinc_no_extreme),
  clean_rows      = nrow(cinc_clean)
)

cinc_removed %>%
  count(removal_reason, sort = TRUE)

cinc_removed %>%
  count(`Site ID`, sort = TRUE)

write_csv(cinc_clean, "CinC_final_table.csv")
write_csv(cinc_removed, "CinC_removed_data.csv")

saveRDS(cinc_clean, "CinC_final_table.rds")
saveRDS(cinc_removed, "CinC_removed_data.rds")

```

# Cleaning up Sites Data 

```{r}
# Purpose / Description:
# Identify and remove periods of unreliable PM2.5 measurements (e.g., sensor drift or flat low values)
# for selected CinC sites using visual inspection and targeted filtering, and confirm improvements
# through updated time series plots.

# Upstream workflow:
# - Requires 'cinc_clean' dataset from the main cleaning workflow
# - Requires 'cinc_removed' initialized to log excluded observations
# - Assumes 'Site ID', 'Date', 'Time', and 'PM2.5' are properly formatted
# - Run after general data cleaning and before final analysis or comparisons

# Cleaning up Site 23.1 Data 
cinc_clean %>%
  filter(`Site ID` == "23.1") %>%
  summarise(
    min_date = min(Date),
    max_date = max(Date)
  )

ggplot(
  cinc_clean %>% filter(`Site ID` == "23.1"),
  aes(x = Date, y = `PM2.5`)
) +
  geom_point(alpha = 0.6, size = 0.8) +
  coord_cartesian(xlim = as.Date(c("2025-07-01", "2025-07-31"))) +
  labs(title = "Zoomed View - Site 23.1",
       x = "Date",
       y = "PM2.5") +
  theme_minimal()

cinc_clean %>%
  filter(`Site ID` == "23.1",
         Date >= as.Date("2025-07-14"),
         Date <= as.Date("2025-07-18")) %>%
  arrange(Date, Time) %>%
  select(Date, Time, `PM2.5`) %>%
  print(n = 100)

#
bad_51 <- cinc_clean %>%
  filter(`Site ID` == "23.1",
         Date >= as.Date("2025-07-14"),
         Date <= as.Date("2025-07-18")) %>%
  mutate(removal_reason = "Sensor drift/flat low values (July 2025)")

cinc_removed <- bind_rows(cinc_removed, bad_51)

cinc_clean <- cinc_clean %>%
  filter(!(`Site ID` == "23.1" &
           Date >= as.Date("2025-07-14") &
           Date <= as.Date("2025-07-18")))

ggplot(
  cinc_clean %>% filter(`Site ID` == "23.1"),
  aes(x = Date, y = `PM2.5`)
) +
  geom_point(alpha = 0.6, size = 0.8) +
  coord_cartesian(xlim = as.Date(c("2025-07-01", "2025-07-31"))) +
  labs(title = "Zoomed View2 - Site 23.1",
       x = "Date",
       y = "PM2.5") +
  theme_minimal()

ggplot(
  cinc_clean %>% filter(`Site ID` == "23.1"),
  aes(x = Date, y = `PM2.5`)
) +
  geom_point(alpha = 0.6, size = 0.8) +
  labs(title = "After Final Trim - Site 23.1",
       x = "Date",
       y = "PM2.5") +
  theme_minimal()

# Cleaning up Site 5.1_19.3_24.2
cinc_clean %>%
  filter(`Site ID` == "5.1_19.3_24.2") %>%
  summarise(
    min_date = min(Date),
    max_date = max(Date)
  )

ggplot(
  cinc_clean %>% filter(`Site ID` == "5.1_19.3_24.2"),
  aes(x = Date, y = `PM2.5`)
) +
  geom_point(alpha = 0.6, size = 0.8) +
  coord_cartesian(xlim = as.Date(c("2025-09-15", "2025-12-15"))) +
  labs(title = "Zoomed View - Site 5.1_19.3_24.2",
       x = "Date",
       y = "PM2.5") +
  theme_minimal()

cinc_clean %>%
  filter(`Site ID` == "5.1_19.3_24.2",
         Date >= as.Date("2025-09-28"),
         Date <= as.Date("2025-11-19")) %>%
  arrange(Date, Time) %>%
  select(Date, Time, `PM2.5`) %>%
  print(n = 100)

#
bad_51 <- cinc_clean %>%
  filter(`Site ID` == "5.1_19.3_24.2",
         Date >= as.Date("2025-09-28"),
         Date <= as.Date("2025-11-19")) %>%
  mutate(removal_reason = "Sensor drift/flat low values (Sept–Nov 2025)")

cinc_removed <- bind_rows(cinc_removed, bad_51)

cinc_clean <- cinc_clean %>%
  filter(!(`Site ID` == "5.1_19.3_24.2" &
           Date >= as.Date("2025-09-28") &
           Date <= as.Date("2025-11-19")))

ggplot(
  cinc_clean %>% filter(`Site ID` == "5.1_19.3_24.2"),
  aes(x = Date, y = `PM2.5`)
) +
  geom_point(alpha = 0.6, size = 0.8) +
  coord_cartesian(xlim = as.Date(c("2025-09-15", "2025-12-15"))) +
  labs(title = "Zoomed View2 - Site 5.1_19.3_24.2",
       x = "Date",
       y = "PM2.5") +
  theme_minimal()

ggplot(
  cinc_clean %>% filter(`Site ID` == "5.1_19.3_24.2"),
  aes(x = Date, y = `PM2.5`)
) +
  geom_point(alpha = 0.6, size = 0.8) +
  labs(title = "After Final Trim - Site 5.1_19.3_24.2",
       x = "Date",
       y = "PM2.5") +
  theme_minimal()

# Cleaning up Site 13.1_33.2
cinc_clean %>%
  filter(`Site ID` == "13.1_33.2") %>%
  summarise(
    min_date = min(Date),
    max_date = max(Date)
  )

ggplot(
  cinc_clean %>% filter(`Site ID` == "13.1_33.2"),
  aes(x = Date, y = `PM2.5`)
) +
  geom_point(alpha = 0.6, size = 0.8) +
  coord_cartesian(xlim = as.Date(c("2025-08-01", "2025-12-05"))) +
  labs(title = "Zoomed View - Site 13.1_33.2",
       x = "Date",
       y = "PM2.5") +
  theme_minimal()

cinc_clean %>%
  filter(`Site ID` == "13.1_33.2",
         Date >= as.Date("2025-08-27"),
         Date <= as.Date("2025-11-18")) %>%
  arrange(Date, Time) %>%
  select(Date, Time, `PM2.5`) %>%
  print(n = 100)

bad_13_combined <- cinc_clean %>%
  filter(`Site ID` == "13.1_33.2",
         Date >= as.Date("2025-08-27"),
         Date <= as.Date("2025-11-18")) %>%
  mutate(removal_reason = "Sensor drift/flat low values (Aug-Nov 2025)")

cinc_removed <- bind_rows(cinc_removed, bad_13_combined)

cinc_clean <- cinc_clean %>%
  filter(!(`Site ID` == "13.1_33.2" &
           Date >= as.Date("2025-08-27") &
           Date <= as.Date("2025-11-18")))

ggplot(
  cinc_clean %>% filter(`Site ID` == "13.1_33.2"),
  aes(x = Date, y = `PM2.5`)
) +
  geom_point(alpha = 0.6, size = 0.8) +
  coord_cartesian(xlim = as.Date(c("2025-08-01", "2025-12-05"))) +
  labs(title = "Zoomed View2 - Site 13.1_33.2",
       x = "Date",
       y = "PM2.5") +
  theme_minimal()

ggplot(
  cinc_clean %>% filter(`Site ID` == "13.1_33.2"),
  aes(x = Date, y = `PM2.5`)
) +
  geom_point(alpha = 0.6, size = 0.8) +
  labs(title = "After Final Trim - Site 13.1_33.2",
       x = "Date",
       y = "PM2.5") +
  theme_minimal()

write_csv(cinc_clean, "CinC_final_table_v2.csv")
write_csv(cinc_removed, "CinC_removed_data_v2.csv")

saveRDS(cinc_clean, "CinC_final_table_v2.rds")
saveRDS(cinc_removed, "CinC_removed_data_v2.rds")
```

# Time Series Plots of cleaned sites
```{r}
# Purpose / Description:
# Generate standardized PM2.5 time series plots for all cleaned CinC sites using the final cleaned dataset.
# This chunk prepares the date-time fields, applies a consistent site order and plotting format,
# and saves high-resolution plots for visual review and later use in analysis or reporting.

# Upstream workflow:
# - Requires the final cleaned CinC dataset file: "CinC_final_table_v2.csv"
# - Assumes the cleaned table includes 'Site ID', 'Date', 'Time', and 'PM2.5'
# - Run after all site-level cleaning and trimming decisions have been completed
# - Run before pairwise comparisons, closest-site matching, or figure selection for reports/presentations

library(readr)
library(dplyr)
library(ggplot2)

# Read the updated cleaned table
cinc_data <- read_csv("CinC_final_table_v2.csv", show_col_types = FALSE)

# Prepare datetime
cinc_data <- cinc_data %>%
  mutate(
    Date = as.Date(Date),
    datetime = as.POSIXct(paste(Date, Time), format = "%Y-%m-%d %H:%M:%S")
  )

# Site order
site_order <- c(
  "1.1","2.1","3.1","5.1_19.3_24.2","7.1_4.1","7.2","8.1_34.2",
  "9.1_17.2","9.2",
  "11.1_32.1_18.2","12.1_5.2","13.1_33.2","14.1","15.1",
  "18.1_26.1","19.1","19.2","20.1","20.2","21.1","22.1","23.1",
  "24.1_1.2","25.1","27.1","28.1","29.1","30.1_17.1","31.2",
  "32.2","33.1","34.1_16.1","35.1","35.2"
)

site_ids <- site_order[site_order %in% unique(cinc_data$`Site ID`)]

# Create output folder
dir.create("CinC_PM25_timeseries_plots_cleaned_v2", showWarnings = FALSE)

for (site in site_ids) {
  
  site_data <- cinc_data %>%
    filter(`Site ID` == site)
  
  p <- ggplot(site_data, aes(x = datetime, y = `PM2.5`)) +
    geom_point(alpha = 0.4, size = 0.7) +
    scale_x_datetime(
      limits = as.POSIXct(c("2024-11-01 00:00:00", "2025-12-31 23:59:59")),
      date_breaks = "1 month",
      date_labels = "%b %Y"
    ) +
    labs(
      title = paste("PM2.5 Time Series - Site", site),
      x = "Time",
      y = "PM2.5 (µg/m³)"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5),
      axis.text.x = element_text(angle = 45, hjust = 1)
    )
  
  print(p)
  
   # 7. Save HIGH-RES PDF (main)
  ggsave(
    filename = paste0("CinC_PM25_timeseries_plots_cleaned_v2_pdf/CinC_site_", site, ".pdf"),
    plot = p,
    width = 10,
    height = 5,
    units = "in"
  )
}
```

# CinC X–Y pairwise plots

```{r cinc_xy_matrix_all_sites_full, warning=FALSE, message=FALSE}
# Purpose / Description:
# Generate a full site-to-site X–Y comparison matrix of daily average PM2.5 for all cleaned CinC sites.
# This chunk reshapes the cleaned dataset to wide format, compares each site against every other site
# using overlapping daily averages, and produces a standardized matrix with identity lines, regression
# lines, and R² values to evaluate agreement and spatial variability across the CinC network.

# Upstream workflow:
# - Requires the final cleaned CinC dataset file: "CinC_final_table_v2.csv"
# - Assumes 'Site ID', 'Date', and 'PM2.5' are available and properly formatted
# - Run after all cleaning, trimming, and final time series checks have been completed
# - Run before selecting final figures or interpreting site-to-site consistency in the report

library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)

# Read cleaned data
cinc_data <- read_csv("CinC_final_table_v2.csv", show_col_types = FALSE)

# Prepare data
cinc_data <- cinc_data %>%
  mutate(
    Date = as.Date(Date),
    `PM2.5` = as.numeric(`PM2.5`)
  )

# Daily averages by Site ID
cinc_daily <- cinc_data %>%
  group_by(`Site ID`, Date) %>%
  summarise(
    pm25 = mean(`PM2.5`, na.rm = TRUE),
    .groups = "drop"
  )

# Wide format
cinc_wide <- cinc_daily %>%
  pivot_wider(names_from = `Site ID`, values_from = pm25)

# Full site order
site_order <- c(
  "1.1","2.1","3.1","5.1_19.3_24.2","7.1_4.1","7.2","8.1_34.2",
  "9.1_17.2","9.2","11.1_32.1_18.2","12.1_5.2","13.1_33.2","14.1","15.1",
  "18.1_26.1","19.1","19.2","20.1","20.2","21.1","22.1","23.1",
  "24.1_1.2","25.1","27.1","28.1","29.1","30.1_17.1","31.2",
  "32.2","33.1","34.1_16.1","35.1","35.2"
)

sites_use <- intersect(site_order, names(cinc_wide))

df_matrix <- cinc_wide %>%
  select(Date, all_of(sites_use))

# Common axis limits
axis_min <- 0
axis_max <- ceiling(max(unlist(df_matrix[sites_use]), na.rm = TRUE) / 10) * 10

plot_list <- list()

for (i in seq_along(sites_use)) {
  for (j in seq_along(sites_use)) {

    y_site <- sites_use[i]
    x_site <- sites_use[j]

    df_pair <- df_matrix %>%
      select(x = all_of(x_site), y = all_of(y_site)) %>%
      filter(!is.na(x) & !is.na(y))

    if (nrow(df_pair) < 10) {

      p <- ggplot() +
        annotate("text", x = 0.5, y = 0.5, label = "No overlap", size = 1.8) +
        theme_void()

    } else {

      fit <- lm(y ~ x, data = df_pair)
      r2 <- summary(fit)$r.squared

      p <- ggplot(df_pair, aes(x = x, y = y)) +
        geom_point(size = 0.35, alpha = 0.5) +
        geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
        geom_smooth(method = "lm", se = FALSE, linewidth = 0.25) +
        coord_fixed(xlim = c(axis_min, axis_max), ylim = c(axis_min, axis_max)) +
        annotate(
          "text",
          x = axis_min + 0.05 * (axis_max - axis_min),
          y = axis_max - 0.08 * (axis_max - axis_min),
          label = paste0("R²=", round(r2, 2)),
          hjust = 0,
          size = 1.6
        ) +
        labs(x = x_site, y = y_site) +
        theme_minimal(base_size = 5) +
        theme(
          axis.text = element_text(size = 4),
          axis.title = element_text(size = 5),
          panel.grid.minor = element_blank(),
          plot.margin = margin(1, 1, 1, 1)
        )
    }

    plot_list[[length(plot_list) + 1]] <- p
  }
}

full_grid <- arrangeGrob(
  grobs = plot_list,
  ncol = length(sites_use),
  nrow = length(sites_use),
  top = "CinC Site-to-Site PM2.5 Comparison"
)

# Save high-resolution PNG
ggsave(
  filename = "CinC_XY_matrix_all_sites.png",
  plot = full_grid,
  width = 40,
  height = 40,
  units = "in",
  dpi = 300,
  limitsize = FALSE
)

# Save PDF too
ggsave(
  filename = "CinC_XY_matrix_all_sites.pdf",
  plot = full_grid,
  width = 40,
  height = 40,
  units = "in",
  limitsize = FALSE
)

# Print to viewer
full_grid
```



# EPA Analysis

```{r}
# Purpose / Description:
# Download and prepare EPA PM2.5 data for selected sites, including both filter-based (FRM, 24-hour)
# and optical (FEM, hourly) measurements. This chunk structures the data into daily averages,
# separates FRM and FEM datasets, and creates a wide-format table for EPA site-to-site comparisons.

# Upstream workflow:
# - Requires valid AQS API credentials via RAQSAPI
# - Assumes selected EPA site numbers and date range are defined
# - Run before any EPA-based analyses (e.g., EPA vs EPA matrix plots, optical vs filter comparisons,
#   and CinC vs EPA matching)

library(RAQSAPI)
library(dplyr)
library(tidyr)

# Set AQS credentials
aqs_credentials(
  username = "rnl34@case.edu",
  key = "copperhawk19"
)

# EPA site numbers
site_numbers <- c("0034", "0038", "0045", "0060", "0065", "0073")

# Download EPA data for all sites
epa_data_list <- list()

for (site in site_numbers) {
  cat("Downloading site", site, "...\n")
  
  tmp <- RAQSAPI::aqs_sampledata_by_site(
    parameter = "88101",
    bdate = as.Date("2024-11-01"),
    edate = as.Date("2025-12-31"),
    stateFIPS = "39",
    countycode = "035",
    sitenum = site,
    return_header = FALSE
  )
  
  tmp <- tmp %>%
    mutate(site_number = site)
  
  epa_data_list[[site]] <- tmp
}

# Combine all sites
epa_data <- bind_rows(epa_data_list)

# Prepare date column
epa_data <- epa_data %>%
  mutate(date = as.Date(date_local))

# Check data types
epa_data %>%
  count(method_type, sample_duration)

# =========================
# FRM (Filter, 24-hour)
# =========================
epa_frm_daily <- epa_data %>%
  filter(method_type == "FRM", sample_duration == "24 HOUR") %>%
  select(site_number, date, sample_measurement) %>%
  rename(frm_pm25 = sample_measurement) %>%
  arrange(site_number, date)

head(epa_frm_daily)

# Wide format (for matrix plots)
epa_frm_wide <- epa_frm_daily %>%
  pivot_wider(
    names_from = site_number,
    values_from = frm_pm25,
    values_fn = mean
  )

head(epa_frm_wide)

# =========================
# FEM (Optical, hourly → daily)
# =========================
epa_fem_hourly <- epa_data %>%
  filter(method_type == "FEM", sample_duration == "1 HOUR")

epa_fem_daily <- epa_fem_hourly %>%
  group_by(site_number, date) %>%
  summarise(
    fem_pm25_daily = mean(sample_measurement, na.rm = TRUE),
    n_obs = n(),
    .groups = "drop"
  ) %>%
  arrange(site_number, date)

head(epa_fem_daily)

# Final checks
dim(epa_data)

# Save full hourly FEM dataset (before daily averaging)
write_csv(epa_fem_hourly, "EPA_FEM_hourly_PM25_2024-11_to_2025-12.csv")
saveRDS(epa_fem_hourly, "EPA_FEM_hourly_PM25_2024-11_to_2025-12.rds")

```

 # EPA FRM Site-to-Site Matrix

```{r epa_frm_matrix, fig.width=18, fig.height=18, echo=FALSE, message=FALSE, warning=FALSE}
# Purpose / Description:
# Generate a site-to-site X–Y comparison matrix of daily average PM2.5 using EPA
# filter-based (FRM, 24-hour) measurements. This chunk evaluates agreement and
# consistency between EPA monitoring sites by computing pairwise regressions,
# plotting identity and regression lines, and displaying R² values across all site pairs.

# Upstream workflow:
# - Requires 'epa_frm_daily' created from the EPA data processing workflow
# - Assumes 'site_number', 'date', and 'frm_pm25' are available and correctly formatted
# - Run after EPA data download and daily aggregation
# - Run before interpreting EPA site consistency or comparing with CinC data

library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)

# 1. Use FRM daily data (already created)
epa_filter <- epa_frm_daily %>%
  rename(pm25 = frm_pm25)


# 2. Wide format
epa_filter_wide <- epa_filter %>%
  pivot_wider(
    names_from = site_number,
    values_from = pm25,
    values_fn = mean
  )


# 3. Site names + axis range
site_names <- sort(setdiff(names(epa_filter_wide), "date"))
n_sites <- length(site_names)

axis_min <- 0
axis_max <- ceiling(max(unlist(epa_filter_wide[site_names]), na.rm = TRUE) / 5) * 5


# 4. Build matrix plots
plot_matrix <- vector("list", n_sites * n_sites)

for (i in seq_len(n_sites)) {
  for (j in seq_len(n_sites)) {

    x_site <- site_names[j]
    y_site <- site_names[i]
    index <- (i - 1) * n_sites + j

    df_pair <- epa_filter_wide %>%
      select(x = all_of(x_site), y = all_of(y_site)) %>%
      filter(!is.na(x) & !is.na(y))

    if (nrow(df_pair) > 2) {
      model <- lm(y ~ x, data = df_pair)
      r2 <- round(summary(model)$r.squared, 2)
      slope <- round(coef(model)[2], 2)
      intercept <- round(coef(model)[1], 2)
      label_text <- paste0("y = ", slope, "x + ", intercept, "\nR² = ", r2)
    } else {
      label_text <- "No overlap"
    }

    x_pos <- axis_min + 1
    y_pos <- axis_max - 1

    p <- ggplot(df_pair, aes(x = x, y = y)) +
      geom_point(alpha = 0.7, size = 0.9) +
      geom_smooth(method = "lm", se = FALSE, linewidth = 0.5) +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed", linewidth = 0.4) +
      coord_fixed(xlim = c(axis_min, axis_max), ylim = c(axis_min, axis_max)) +
      annotate(
        "text",
        x = x_pos, y = y_pos,
        label = label_text,
        hjust = 0, vjust = 1,
        size = 2.2
      ) +
      labs(
        x = paste("Site", x_site),
        y = paste("Site", y_site)
      ) +
      theme_minimal(base_size = 8) +
      theme(
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 6),
        panel.grid.minor = element_blank(),
        plot.margin = margin(2, 2, 2, 2)
      )

    plot_matrix[[index]] <- p
  }
}


# 5. Draw + save
grid_plot <- grid.arrange(
  grobs = plot_matrix,
  ncol = n_sites,
  nrow = n_sites,
  top = "EPA FRM Site-to-Site Matrix (Daily PM2.5)"
)

print(grid_plot)

# SAVE EPA FRM SITE-TO-SITE MATRIX AS HIGH-RES PDF

ggsave(
  "epa_frm_matrix.pdf",
  plot = grid_plot,
  width = 18,
  height = 18,
  units = "in"
)
```

# EPA Optical (Daily, ALL data) - Continuous monitors (FEM) averaged to daily.
```{r epa-optical-daily-matrix, fig.width=18, fig.height=18, echo=FALSE, message=FALSE, warning=FALSE}
# Purpose / Description:
# Generate a site-to-site X–Y comparison matrix of daily average PM2.5 using EPA
# optical (FEM) measurements. This evaluates agreement across EPA continuous
# monitoring sites using paired daily averages, regression lines, identity lines,
# and R² values.

# Upstream workflow:
# - Requires 'epa_fem_daily' from EPA processing
# - Uses only PM2.5 daily averages (excludes n_obs)
# - Run after EPA data preparation and before matched FRM–FEM comparisons

library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)

# =========================
# 1. Clean FEM daily data (REMOVE n_obs)
# =========================
epa_optical_daily <- epa_fem_daily %>%
  select(site_number, date, fem_pm25_daily) %>%   # remove n_obs
  rename(pm25 = fem_pm25_daily)

# =========================
# 2. Wide format
# =========================
epa_optical_wide <- epa_optical_daily %>%
  pivot_wider(
    names_from = site_number,
    values_from = pm25
  )

# =========================
# 3. Site names and axis limits
# =========================
site_names <- sort(setdiff(names(epa_optical_wide), "date"))
n_sites <- length(site_names)

axis_min <- 0
axis_max <- ceiling(max(unlist(epa_optical_wide[site_names]), na.rm = TRUE) / 5) * 5

# =========================
# 4. Build matrix plots
# =========================
plot_matrix <- vector("list", n_sites * n_sites)

for (i in seq_len(n_sites)) {
  for (j in seq_len(n_sites)) {

    x_site <- site_names[j]
    y_site <- site_names[i]
    index <- (i - 1) * n_sites + j

    df_pair <- epa_optical_wide %>%
      select(x = all_of(x_site), y = all_of(y_site)) %>%
      filter(!is.na(x) & !is.na(y))

    if (nrow(df_pair) > 2) {
      model <- lm(y ~ x, data = df_pair)
      r2 <- round(summary(model)$r.squared, 2)
      slope <- round(coef(model)[2], 2)
      intercept <- round(coef(model)[1], 2)
      label_text <- paste0("y = ", slope, "x + ", intercept, "\nR² = ", r2)
    } else {
      label_text <- "No overlap"
    }

    x_pos <- axis_min + 1
    y_pos <- axis_max - 1

    p <- ggplot(df_pair, aes(x = x, y = y)) +
      geom_point(alpha = 0.7, size = 0.9) +
      geom_smooth(method = "lm", se = FALSE, linewidth = 0.5) +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed", linewidth = 0.4) +
      coord_fixed(xlim = c(axis_min, axis_max), ylim = c(axis_min, axis_max)) +
      annotate(
        "text",
        x = x_pos, y = y_pos,
        label = label_text,
        hjust = 0, vjust = 1,
        size = 2.2
      ) +
      labs(
        x = paste("Site", x_site),
        y = paste("Site", y_site)
      ) +
      theme_minimal(base_size = 8) +
      theme(
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 6),
        panel.grid.minor = element_blank(),
        plot.margin = margin(2, 2, 2, 2)
      )

    plot_matrix[[index]] <- p
  }
}

# =========================
# 5. Draw and save
# =========================
grid_plot <- grid.arrange(
  grobs = plot_matrix,
  ncol = n_sites,
  nrow = n_sites,
  top = "EPA Optical Site-to-Site Matrix (Daily PM2.5)"
)

print(grid_plot)

# SAVE EPA OPTICAL (DAILY, ALL DATA) MATRIX AS HIGH-RES PDF

ggsave(
  "epa_optical_daily_matrix.pdf",
  plot = grid_plot,
  width = 18,
  height = 18,
  units = "in"
)
```

# EPA Optical (Daily, only filter days)
```{r epa-optical-matrix, fig.width=18, fig.height=18, echo=FALSE, message=FALSE, warning=FALSE}
# Purpose / Description:
# Generate a site-to-site X–Y comparison matrix of daily average PM2.5 using EPA
# optical (FEM) measurements restricted to days when EPA filter-based (FRM) data
# are available. This allows comparison of optical site-to-site agreement under
# the same sampling dates used for filter measurements.

# Upstream workflow:
# - Requires 'epa_fem_daily' and 'epa_frm_daily' from the EPA processing workflow
# - Assumes 'site_number', 'date', 'fem_pm25_daily', and 'frm_pm25' are available
# - Run after EPA FRM and FEM daily datasets are created
# - Run before EPA optical vs filter comparisons and CinC–EPA comparisons

library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)


# 1. Keep optical daily data only on dates with FRM measurements
epa_optical_matched <- epa_fem_daily %>%
  select(site_number, date, fem_pm25_daily) %>%
  semi_join(
    epa_frm_daily %>% select(site_number, date),
    by = c("site_number", "date")
  ) %>%
  rename(pm25 = fem_pm25_daily)


# 2. Wide format
epa_optical_matched_wide <- epa_optical_matched %>%
  pivot_wider(
    names_from = site_number,
    values_from = pm25
  )

# 3. Site names and common axis range
site_names <- sort(setdiff(names(epa_optical_matched_wide), "date"))
n_sites <- length(site_names)

axis_min <- 0
axis_max <- ceiling(max(unlist(epa_optical_matched_wide[site_names]), na.rm = TRUE) / 5) * 5

# 4. Build matrix plots
plot_matrix <- vector("list", n_sites * n_sites)

for (i in seq_len(n_sites)) {
  for (j in seq_len(n_sites)) {

    x_site <- site_names[j]
    y_site <- site_names[i]
    index <- (i - 1) * n_sites + j

    df_pair <- epa_optical_matched_wide %>%
      select(x = all_of(x_site), y = all_of(y_site)) %>%
      filter(!is.na(x) & !is.na(y))

    if (nrow(df_pair) > 2) {
      model <- lm(y ~ x, data = df_pair)
      r2 <- round(summary(model)$r.squared, 2)
      slope <- round(coef(model)[2], 2)
      intercept <- round(coef(model)[1], 2)
      label_text <- paste0("y = ", slope, "x + ", intercept, "\nR² = ", r2)
    } else {
      label_text <- "No overlap"
    }

    x_pos <- axis_min + 1
    y_pos <- axis_max - 1

    p <- ggplot(df_pair, aes(x = x, y = y)) +
      geom_point(alpha = 0.7, size = 0.9) +
      geom_smooth(method = "lm", se = FALSE, linewidth = 0.5) +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed", linewidth = 0.4) +
      coord_fixed(xlim = c(axis_min, axis_max), ylim = c(axis_min, axis_max)) +
      annotate(
        "text",
        x = x_pos, y = y_pos,
        label = label_text,
        hjust = 0, vjust = 1,
        size = 2.2
      ) +
      labs(
        x = paste("Site", x_site),
        y = paste("Site", y_site)
      ) +
      theme_minimal(base_size = 8) +
      theme(
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 6),
        panel.grid.minor = element_blank(),
        plot.margin = margin(2, 2, 2, 2)
      )

    plot_matrix[[index]] <- p
  }
}

# 5. Draw and save
grid_plot <- grid.arrange(
  grobs = plot_matrix,
  ncol = n_sites,
  nrow = n_sites,
  top = "EPA Optical Site-to-Site Matrix (Daily PM2.5, FRM Days Only)"
)

print(grid_plot)

# SAVE EPA OPTICAL (FRM DAYS ONLY) MATRIX AS HIGH-RES PDF

ggsave(
  "epa_optical_matrix.pdf",
  plot = grid_plot,
  width = 18,
  height = 18,
  units = "in"
)

```

# Correlation vs Distance
```{r}
epa_coords <- epa_data %>%
  select(site_number, latitude, longitude) %>%
  distinct() %>%
  mutate(site_number = as.character(site_number))

# Compute distance matrix
library(geosphere)

dist_matrix <- distm(
  epa_coords[, c("longitude", "latitude")],
  fun = distHaversine
)

rownames(dist_matrix) <- epa_coords$site_number
colnames(dist_matrix) <- epa_coords$site_number 

# Convert to long format
library(tidyr)

dist_long <- as.data.frame(dist_matrix) %>%
  mutate(site1 = rownames(dist_matrix)) %>%
  pivot_longer(-site1, names_to = "site2", values_to = "distance_m") %>%
  mutate(distance_km = distance_m / 1000)

epa_r2 <- list()

site_names <- sort(setdiff(names(epa_filter_wide), "date"))

for (i in seq_along(site_names)) {
  for (j in seq_along(site_names)) {

    site_i <- site_names[i]
    site_j <- site_names[j]

    df_pair <- epa_filter_wide %>%
      select(x = all_of(site_i), y = all_of(site_j)) %>%
      filter(!is.na(x) & !is.na(y))

    if (nrow(df_pair) > 5) {
      r2 <- summary(lm(y ~ x, data = df_pair))$r.squared
    } else {
      r2 <- NA
    }

    epa_r2[[length(epa_r2) + 1]] <- data.frame(
      site1 = site_i,
      site2 = site_j,
      r2 = r2
    )
  }
}

epa_r2_df <- bind_rows(epa_r2)

cor_dist <- epa_r2_df %>%
  left_join(dist_long, by = c("site1", "site2")) %>%
  filter(site1 != site2)

cor_dist_clean <- cor_dist %>%
  filter(site1 < site2)   # removes duplicates + diagonal

# Plot correlation vs distance
ggplot(cor_dist, aes(x = distance_km, y = r2)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Correlation vs Distance Between EPA Sites",
    x = "Distance (km)",
    y = "R² (Correlation Strength)"
  ) +
  theme_minimal()


```
# 
```{r epa-fem-vs-frm-by-site, fig.width=16, fig.height=10, echo=FALSE, message=FALSE, warning=FALSE}
# Purpose / Description:
# Compare EPA optical (FEM) and filter-based (FRM) PM2.5 measurements at each EPA
# site using matched daily averages. This chunk joins daily FEM and FRM data by site
# and date, then generates site-specific X–Y plots with regression lines, identity
# lines, and R² values to evaluate agreement between the two EPA measurement methods.

# Upstream workflow:
# - Requires 'epa_frm_daily' and 'epa_fem_daily' from the EPA processing workflow
# - Assumes 'site_number', 'date', 'frm_pm25', and 'fem_pm25_daily' are available
# - Run after EPA pairwise site-to-site comparisons
# - Run before CinC vs EPA comparisons

library(dplyr)
library(ggplot2)
library(gridExtra)


# 1. Match FRM and FEM by site and date
epa_method_compare <- epa_frm_daily %>%
  inner_join(
    epa_fem_daily %>% select(site_number, date, fem_pm25_daily),
    by = c("site_number", "date")
  )


# 2. EPA site list
site_names <- sort(unique(epa_method_compare$site_number))

axis_min <- 0
axis_max <- ceiling(
  max(c(epa_method_compare$frm_pm25, epa_method_compare$fem_pm25_daily), na.rm = TRUE) / 5
) * 5


# 3. Build one plot per site
plot_list <- vector("list", length(site_names))

for (i in seq_along(site_names)) {

  site_i <- site_names[i]

  df_site <- epa_method_compare %>%
    filter(site_number == site_i)

  if (nrow(df_site) > 2) {
    model <- lm(fem_pm25_daily ~ frm_pm25, data = df_site)
    r2 <- round(summary(model)$r.squared, 2)
    slope <- round(coef(model)[2], 2)
    intercept <- round(coef(model)[1], 2)
    label_text <- paste0("y = ", slope, "x + ", intercept, "\nR² = ", r2)
  } else {
    label_text <- "No overlap"
  }

  x_pos <- axis_min + 1
  y_pos <- axis_max - 1

  p <- ggplot(df_site, aes(x = frm_pm25, y = fem_pm25_daily)) +
    geom_point(alpha = 0.7, size = 1.2) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 0.6) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", linewidth = 0.5) +
    coord_fixed(xlim = c(axis_min, axis_max), ylim = c(axis_min, axis_max)) +
    annotate(
      "text",
      x = x_pos, y = y_pos,
      label = label_text,
      hjust = 0, vjust = 1,
      size = 3
    ) +
    labs(
      title = paste("EPA Site", site_i),
      x = "FRM daily PM2.5",
      y = "FEM daily PM2.5"
    ) +
    theme_minimal(base_size = 10) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5),
      panel.grid.minor = element_blank()
    )

  plot_list[[i]] <- p
}


# 4. Arrange and save
grid_plot <- grid.arrange(
  grobs = plot_list,
  ncol = 3,
  top = "EPA FEM vs FRM by Site (Daily PM2.5)"
)

print(grid_plot)

ggsave(
  "epa_fem_vs_frm_by_site.pdf",
  plot = grid_plot,
  width = 16,
  height = 10,
  units = "in"
)
```
# 

```{r}
# Purpose / Description:
# Match CinC sites to EPA sites using geographic distance. This chunk extracts unique
# coordinates for each CinC and EPA site, computes distances between all possible EPA–CinC
# pairs using the Haversine formula, identifies the closest CinC site for each EPA monitor,
# and classifies each matched pair as Collocated, Nearby, or Far for use in later CinC vs EPA analyses.

# Upstream workflow:
# - Requires 'cinc_site_table_final' with columns: 'Site ID', 'Lat', and 'Long'
# - Requires 'epa_data' with columns: 'site_number', 'latitude', and 'longitude'
# - Run after EPA data download and before any CinC vs EPA comparison plots or summaries

library(dplyr)
library(geosphere)

cinc_coords <- cinc_site_table_final %>%
  select(`Site ID`, Lat, Long) %>%
  distinct() %>%
  rename(
    cinc_site = `Site ID`,
    lat_cinc = Lat,
    lon_cinc = Long
  )

epa_coords <- epa_data %>%
  select(site_number, latitude, longitude) %>%
  distinct() %>%
  rename(
    epa_site = site_number,
    lat_epa = latitude,
    lon_epa = longitude
  )

distance_table <- expand.grid(
  epa_site = epa_coords$epa_site,
  cinc_site = cinc_coords$cinc_site
) %>%
  left_join(epa_coords, by = "epa_site") %>%
  left_join(cinc_coords, by = "cinc_site") %>%
  rowwise() %>%
  mutate(
    distance_km = distHaversine(
      c(lon_epa, lat_epa),
      c(lon_cinc, lat_cinc)
    ) / 1000
  ) %>%
  ungroup()

closest_sites <- distance_table %>%
  group_by(epa_site) %>%
  slice_min(distance_km, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(
    category = case_when(
      distance_km < 0.5 ~ "Collocated",
      distance_km <= 3 ~ "Nearby",
      TRUE ~ "Far"
    )
  )

closest_sites


```

# 
```{r cinc-vs-epa-frm-daily, fig.width=14, fig.height=10, echo=FALSE, message=FALSE, warning=FALSE}
# Purpose / Description:
# Compare daily average PM2.5 from cleaned CinC sites with EPA filter-based (FRM, 24-hour)
# measurements using the closest EPA–CinC site pairs. This chunk computes CinC daily averages,
# joins them with EPA FRM daily data by matched site and date, adds distance-based pairing
# information, calculates per-pair regression statistics, and generates cleaner faceted comparison
# plots with consistent axes for easier visual comparison across site pairs.

# Upstream workflow:
# - Requires 'closest_sites' from the EPA–CinC distance matching workflow
# - Requires 'epa_frm_daily' from the EPA processing workflow
# - Requires 'cinc_clean' from the completed CinC cleaning workflow
# - Run after closest-site matching and before final figure selection or interpretation

library(dplyr)
library(ggplot2)

cinc_daily <- cinc_clean %>%
  group_by(`Site ID`, Date) %>%
  summarise(
    cinc_pm25 = mean(`PM2.5`, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  rename(
    cinc_site = `Site ID`,
    date = Date
  )

epa_frm_daily2 <- epa_frm_daily %>%
  rename(
    epa_site = site_number
  )

cinc_epa_frm <- closest_sites %>%
  select(epa_site, cinc_site, category, distance_km) %>%
  left_join(epa_frm_daily2, by = "epa_site") %>%
  left_join(cinc_daily, by = c("cinc_site", "date")) %>%
  filter(!is.na(frm_pm25), !is.na(cinc_pm25))

stats_df <- cinc_epa_frm %>%
  group_by(epa_site, cinc_site, category, distance_km) %>%
  summarise(
    slope = coef(lm(cinc_pm25 ~ frm_pm25))[2],
    intercept = coef(lm(cinc_pm25 ~ frm_pm25))[1],
    r2 = summary(lm(cinc_pm25 ~ frm_pm25))$r.squared,
    .groups = "drop"
  ) %>%
  mutate(
    panel_lab = paste0(
      "EPA ", epa_site, " vs CinC ", cinc_site,
      "\n", category, " (", round(distance_km, 2), " km)"
    ),
    eq_lab = paste0(
      "y = ", round(slope, 2), "x + ", round(intercept, 2),
      "\nR² = ", round(r2, 2)
    )
  )

cinc_epa_frm <- cinc_epa_frm %>%
  mutate(
    panel_lab = paste0(
      "EPA ", epa_site, " vs CinC ", cinc_site,
      "\n", category, " (", round(distance_km, 2), " km)"
    )
  )

axis_max <- ceiling(max(c(cinc_epa_frm$frm_pm25, cinc_epa_frm$cinc_pm25), na.rm = TRUE) / 5) * 5

ggplot(cinc_epa_frm, aes(x = frm_pm25, y = cinc_pm25)) +
  geom_point(alpha = 0.65, size = 1.4) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", linewidth = 0.5) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.7) +
  geom_text(
    data = stats_df,
    aes(x = 2, y = axis_max - 2, label = eq_lab),
    inherit.aes = FALSE,
    hjust = 0,
    vjust = 1,
    size = 3
  ) +
  facet_wrap(~ panel_lab, ncol = 3) +
  coord_fixed(xlim = c(0, axis_max), ylim = c(0, axis_max)) +
  labs(
    title = "CinC vs EPA FRM (Daily PM2.5)",
    x = "EPA FRM PM2.5 (µg/m³)",
    y = "CinC PM2.5 (µg/m³)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold"),
    strip.text = element_text(size = 10),
    panel.grid.minor = element_blank()
  )
#  Print plot

print(cinc_epa_frm_plot)


# Save HIGH-RES PDF

ggsave(
  filename = "cinc_vs_epa_frm_daily.pdf",
  plot = cinc_epa_frm_plot,
  width = 14,
  height = 10,
  units = "in"
  )
```

# Comparison of Daily PM2.5: CinC vs EPA Optical Monitors

```{r cinc-vs-epa-optical-daily, fig.width=14, fig.height=10, echo=FALSE, message=FALSE, warning=FALSE}
# Purpose / Description:
# Compare daily average PM2.5 from cleaned CinC sites with EPA optical (FEM) daily
# average PM2.5 using the closest EPA–CinC site pairs. This chunk computes CinC daily
# averages, prepares EPA optical daily averages, joins both datasets by matched site
# and date, carries forward the distance-based pairing category, calculates per-pair
# regression statistics, and generates faceted comparison plots with common x- and y-axis
# limits for direct visual comparison across site pairs.

# Upstream workflow:
# - Requires 'closest_sites' from the EPA–CinC distance matching workflow
# - Requires 'epa_fem_daily' from the EPA processing workflow
# - Requires 'cinc_clean' from the completed CinC cleaning workflow
# - Assumes the following columns exist and are correctly formatted:
#   'Site ID', 'Date', 'PM2.5', 'epa_site', 'cinc_site', 'category', 'distance_km',
#   'site_number', 'date', and 'fem_pm25_daily'
# - Run after closest-site matching and before final interpretation of CinC vs EPA optical agreement

library(dplyr)
library(ggplot2)

cinc_daily <- cinc_clean %>%
  group_by(`Site ID`, Date) %>%
  summarise(
    cinc_pm25 = mean(`PM2.5`, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  rename(
    cinc_site = `Site ID`,
    date = Date
  )

epa_optical_daily <- epa_fem_daily %>%
  select(site_number, date, fem_pm25_daily) %>%
  rename(
    epa_site = site_number,
    optical_pm25 = fem_pm25_daily
  )

cinc_epa_optical <- closest_sites %>%
  select(epa_site, cinc_site, category, distance_km) %>%
  left_join(epa_optical_daily, by = "epa_site") %>%
  left_join(cinc_daily, by = c("cinc_site", "date")) %>%
  filter(!is.na(optical_pm25), !is.na(cinc_pm25))

stats_df <- cinc_epa_optical %>%
  group_by(epa_site, cinc_site, category, distance_km) %>%
  summarise(
    n = n(),
    slope = if (n() > 1) coef(lm(cinc_pm25 ~ optical_pm25))[2] else NA_real_,
    intercept = if (n() > 1) coef(lm(cinc_pm25 ~ optical_pm25))[1] else NA_real_,
    r2 = if (n() > 1) summary(lm(cinc_pm25 ~ optical_pm25))$r.squared else NA_real_,
    .groups = "drop"
  ) %>%
  mutate(
    panel_lab = paste0(
      "EPA ", epa_site, " vs CinC ", cinc_site,
      "\n", category, " (", round(distance_km, 2), " km)"
    ),
    eq_lab = ifelse(
      is.na(r2),
      "Too few points",
      paste0(
        "y = ", round(slope, 2), "x + ", round(intercept, 2),
        "\nR² = ", round(r2, 2)
      )
    )
  )

cinc_epa_optical <- cinc_epa_optical %>%
  mutate(
    panel_lab = paste0(
      "EPA ", epa_site, " vs CinC ", cinc_site,
      "\n", category, " (", round(distance_km, 2), " km)"
    )
  )

axis_max <- ceiling(
  max(c(cinc_epa_optical$optical_pm25, cinc_epa_optical$cinc_pm25), na.rm = TRUE) / 5
) * 5

ggplot(cinc_epa_optical, aes(x = optical_pm25, y = cinc_pm25)) +
  geom_point(alpha = 0.5, size = 1.3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", linewidth = 0.5) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.7) +
  geom_text(
    data = stats_df,
    aes(x = 2, y = axis_max - 5, label = eq_lab),
    inherit.aes = FALSE,
    hjust = 0,
    vjust = 1,
    size = 3
  ) +
  facet_wrap(~ panel_lab, ncol = 3) +
  coord_fixed(xlim = c(0, axis_max), ylim = c(0, axis_max)) +
  labs(
    title = "Comparison of Daily PM2.5: CinC vs EPA Optical Monitors",
    x = "EPA Optical PM2.5 (µg/m³)",
    y = "CinC PM2.5 (µg/m³)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold"),
    strip.text = element_text(size = 10),
    panel.grid.minor = element_blank()
  )

print(cinc_epa_optical_plot)


# 9. Save high-resolution PDF

ggsave(
  filename = "cinc_vs_epa_optical_daily.pdf",
  plot = cinc_epa_optical_plot,
  width = 14,
  height = 10,
  units = "in"
)

```

```{r epa_time_series_by_sites, echo=FALSE, message=FALSE, warning=FALSE}

# Purpose / Description:
# Generate individual EPA PM2.5 time series plots for each EPA monitoring site using
# both filter-based (FRM, 24-hour) measurements and optical/continuous (FEM) daily
# average measurements. This chunk prepares the two EPA datasets in a common long
# format, combines them into a single plotting table, and then loops through each

# Upstream workflow:
# - Requires 'epa_frm_daily' from the EPA data processing workflow
# - Requires 'epa_fem_daily' from the EPA data processing workflow
# - Assumes the following columns exist and are correctly formatted:
#   'site_number', 'date', 'frm_pm25', and 'fem_pm25_daily'
# - Run after EPA data download and daily aggregation
# - Run before final figure selection, EPA time series interpretation, and PDF export
#   of the completed figure sets


library(dplyr)
library(ggplot2)


# 1. Prepare FRM data
epa_frm_ts <- epa_frm_daily %>%
  transmute(
    site_number,
    date,
    pm25 = frm_pm25,
    method = "FRM"
  )


# 2. Prepare FEM daily data

epa_fem_ts <- epa_fem_daily %>%
  transmute(
    site_number,
    date,
    pm25 = fem_pm25_daily,
    method = "FEM"
  )


# 3. Combine datasets
epa_ts <- bind_rows(epa_frm_ts, epa_fem_ts)


# 4. Create folder for output
dir.create("EPA_time_series_plots", showWarnings = FALSE)


# 5. Define FULL date sequence (ensures all months appear)
full_dates <- seq(
  as.Date("2024-11-01"),
  as.Date("2025-12-31"),
  by = "1 month"
)


# 6. Get site list
site_ids <- sort(unique(epa_ts$site_number))


# 7. Loop through sites
for (site in site_ids) {

  site_data <- epa_ts %>%
    filter(site_number == site)

  p <- ggplot(site_data, aes(x = date, y = pm25)) +
    
    geom_point(aes(shape = method, alpha = method), size = 1) +
    
    scale_shape_manual(values = c("FEM" = 16, "FRM" = 17)) +
    scale_alpha_manual(values = c("FEM" = 0.45, "FRM" = 0.9)) +
    
    # X-AXIS (ALL months shown)
    scale_x_date(
      limits = c(as.Date("2024-11-01"), as.Date("2025-12-31")),
      breaks = full_dates,
      date_labels = "%b %Y",
      expand = expansion(mult = c(0.01, 0.01))
    ) +
    
    labs(
      title = paste("EPA PM2.5 Time Series - Site", site),
      subtitle = "FEM daily averages and FRM 24-hour measurements",
      x = "Date",
      y = "PM2.5 (µg/m³)",
      shape = "Method",
      alpha = "Method"
    ) +
    
    theme_minimal(base_size = 11) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5),
      
      # rotate for readability since now MANY months
      axis.text.x = element_text(angle = 60, hjust = 1, size = 8),
      
      panel.grid.minor = element_blank()
    )

  print(p)

  ggsave(
    filename = paste0("EPA_time_series_plots/EPA_site_", site, "_time_series.pdf"),
    plot = p,
    width = 10,
    height = 5,
    units = "in"
  )
}

```

```{r}


```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
