---
title: "Untitled"
author: "CLEANINCLE API"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# CLEANinCLE API

```{r}
library(httr2)
library(jsonlite)
library(dplyr)
library(tidyr)
library(readr)
library(lubridate)


# 1. API settings

api_url <- "https://data.cleanincle.issacs.dev"
api_key <- "2KYc26u6sn6bPAaZHJx5R3DSzQL8Y6sV38hy0c3t"

params <- list(
  start_date = "2024-11-01T00:00:00Z",
  end_date   = "2026-01-01T00:00:00Z"
)

# 2. Download all pages
# -----------------------------
all_records <- list()
current_params <- params

message("Starting CLEANinCLE API download...")

repeat {
  resp <- request(api_url) |>
    req_headers(`x-api-key` = api_key) |>
    req_url_query(!!!current_params) |>
    req_retry(max_tries = 3) |>
    req_perform()

  body <- resp_body_json(resp, simplifyVector = FALSE)

  all_records <- c(all_records, body$records)

  if (is.null(body$next_token)) break

  current_params$next_token <- body$next_token
}

# -----------------------------
# 3. Convert to dataframe
# -----------------------------
df_raw <- bind_rows(all_records)

df_raw <- df_raw %>%
  unnest_wider(where(is.list), names_sep = "_")

# -----------------------------
# 4. Save raw downloaded file
# -----------------------------
write_csv(df_raw, "CLEANinCLE_raw_2024-11_to_2025-12.csv")
saveRDS(df_raw, "CLEANinCLE_raw_2024-11_to_2025-12.rds")

message("Download complete.")
dim(df_raw)
names(df_raw)
```

# PM2.5-only CLEANinCLE dataset
```{r}
library(dplyr)
library(lubridate)
library(readr)

cinc_pm25 <- df_raw %>%
  mutate(
    time = ymd_hms(time, tz = "UTC"),
    date = as.Date(time),
    pm_2.5 = as.numeric(pm_2.5)
  ) %>%
  select(
    device_eui,
    location_name,
    latitude,
    longitude,
    device_number,
    device_status,
    time,
    date,
    pm_2.5
  ) %>%
  filter(!is.na(pm_2.5)) %>%
  arrange(device_number, time)

# Save PM2.5-only dataset
write_csv(cinc_pm25, "CLEANinCLE_PM25_2024-11_to_2025-12.csv")
saveRDS(cinc_pm25, "CLEANinCLE_PM25_2024-11_to_2025-12.rds")

# Checks
dim(cinc_pm25)
names(cinc_pm25)
head(cinc_pm25)
range(cinc_pm25$time, na.rm = TRUE)
```

# CinC provide a unique device name for each site
```{r}
library(dplyr)
library(lubridate)

cinc_sites <- cinc_pm25 %>%
  mutate(
    latitude = as.numeric(latitude),
    longitude = as.numeric(longitude),
    datetime = as.POSIXct(time),
    lat_round = round(latitude, 5),
    lon_round = round(longitude, 5)
  ) %>%
  filter(
    !is.na(lat_round),
    !is.na(lon_round),
    lat_round != 0,
    lon_round != 0
  ) %>%
  arrange(device_number, datetime) %>%
  group_by(device_number) %>%
  mutate(
    loc_change = (lat_round != lag(lat_round, default = first(lat_round))) |
                 (lon_round != lag(lon_round, default = first(lon_round))),
    site_number = cumsum(loc_change) + 1,
    Site_ID = paste0(device_number, ".", site_number)
  ) %>%
  ungroup() %>%
  select(-loc_change, -site_number, -lat_round, -lon_round)

cinc_unique_site_table <- cinc_sites %>%
  group_by(Site_ID, device_number, location_name) %>%
  summarise(
    latitude = first(latitude),
    longitude = first(longitude),
    start_time = min(datetime, na.rm = TRUE),
    end_time = max(datetime, na.rm = TRUE),
    n_obs = n(),
    .groups = "drop"
  ) %>%
  arrange(device_number, start_time)

cinc_unique_site_table

write_csv(cinc_unique_site_table, "CLEANinCLE_unique_site_table.csv")
saveRDS(cinc_unique_site_table, "CLEANinCLE_unique_site_table.rds")
```


## Updated CinC Time Series Plot 0f PM2.5 for all devices
```{r}
library(ggplot2)
library(dplyr)

# get all unique site IDs
site_ids <- sort(unique(cinc_sites$Site_ID))

# create folder to save plots
dir.create("CinC_PM25_timeseries_plots", showWarnings = FALSE)

for (site in site_ids) {
  
  sensor_data <- cinc_sites %>% filter(Site_ID == site)
  
  dev <- unique(sensor_data$device_number)[1]
  loc_name <- unique(sensor_data$location_name)[1]
  
  p <- ggplot(sensor_data, aes(datetime, pm_2.5)) +
    geom_point(alpha = 0.4, size = 0.7) +
    scale_x_datetime(
      limits = as.POSIXct(c("2024-11-01", "2025-12-31")),
      date_breaks = "1 month",
      date_labels = "%b %Y"
    ) +
    labs(
      title = paste("PM2.5 Time Series - Site", site, "(Device", dev, "-", loc_name, ")"),
      x = "Time",
      y = "PM2.5 (µg/m³)"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  print(p)
  
  ggsave(
    filename = paste0("CinC_PM25_timeseries_plots/CinC_site_", site, ".png"),
    plot = p,
    width = 10,
    height = 5,
    dpi = 300
  )
}


```

# CLEANinCLE Daily Sensor Coverage
```{r}
# Expand each site into daily dates
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)

coverage_daily <- cinc_unique_site_table %>%
  mutate(
    start_date = as.Date(start_time),
    end_date   = as.Date(end_time)
  ) %>%
  rowwise() %>%
  mutate(date = list(seq.Date(start_date, end_date, by = "day"))) %>%
  ungroup() %>%
  unnest(date)


coverage_daily <- coverage_daily %>%
  mutate(
    device = as.numeric(sub("\\..*", "", Site_ID)),
    deploy = as.numeric(sub(".*\\.", "", Site_ID))
  ) %>%
  arrange(device, deploy) %>%
  mutate(
    Site_ID = factor(Site_ID, levels = rev(unique(Site_ID)))
  )

# Plot the daily coverage
ggplot(coverage_daily, aes(x = date, y = Site_ID)) +
  geom_line(linewidth = 0.5) +
  labs(
    title = "CLEANinCLE Daily Sensor Coverage",
    x = "Date",
    y = "Site (Device.Location)"
  ) +
  scale_x_date(
    limits = c(as.Date("2024-11-01"), as.Date("2025-12-31")),
    date_breaks = "1 months",
    date_labels = "%b %Y"
  ) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.02))) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 6),
    panel.grid.major.y = element_blank(), 
    panel.grid.minor = element_blank()
  )

ggsave("CinC_daily_coverage.png",
       width = 10,
       height = 14,   
       dpi = 300)
```
```{r}
ggplot(coverage_daily, aes(x = date, y = Site_ID)) +
  geom_line(linewidth = 0.5) +
  labs(
    title = "CLEANinCLE Sensor Deployment Timeline",
    x = "Date",
    y = "Site (Device.Location)"
  ) +
  scale_x_date(
    limits = c(as.Date("2024-11-01"), as.Date("2025-12-31")),
    date_breaks = "2 months",
    date_labels = "%b %Y"
  ) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.02))) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 6),
    panel.grid.major.y = element_blank(),   # 👈 key
    panel.grid.minor = element_blank()
  )

ggsave("CinC_sensor_coverage.png",
       width = 10,
       height = 14,   
       dpi = 300)
```

```{r}
library(dplyr)

site_lookup_clean <- site_lookup %>%
  select(1, 5) %>%
  rename(
    sensor_id = 1,
    site_id = 2
  ) %>%
  mutate(
    sensor_id = as.character(sensor_id),
    site_id = as.character(site_id)
  ) %>%
  filter(
    grepl("^[0-9]+\\.[0-9]+$", sensor_id),
    !is.na(site_id),
    site_id != "",
    !grepl("exclude", site_id, ignore.case = TRUE),
    !grepl("combined", site_id, ignore.case = TRUE)
  )

head(site_lookup_clean)
distinct(site_lookup_clean)


library(dplyr)

cinc_site_table_full <- cinc_sites %>%
  mutate(sensor_id = as.character(Site_ID)) %>%
  left_join(site_lookup_clean, by = "sensor_id") %>%
  filter(!is.na(site_id)) %>%
  mutate(
    latitude = as.numeric(latitude),
    longitude = as.numeric(longitude),
    date = as.Date(datetime),
    time_only = format(datetime, "%H:%M:%S"),
    pm25 = pm_2.5
  ) %>%
  select(
    sensor_id,
    site_id,
    latitude,
    longitude,
    date,
    time = time_only,
    pm25
  ) %>%
  arrange(latitude, longitude, date, time)

head(cinc_site_table_full)
distinct(cinc_site_table_full, sensor_id, site_id)
nrow(cinc_site_table_full)

```

```{r}
cinc_daily_site <- cinc_site_table_full %>%
  group_by(site_id, date) %>%
  summarise(
    pm25_daily = mean(pm25, na.rm = TRUE),
    n_obs = n(),
    .groups = "drop"
  ) %>%
  arrange(site_id, date)

head(cinc_daily_site)

cinc_daily_site %>%
  count(site_id, name = "n_days") %>%
  arrange(desc(n_days))

```
```{r}
cinc_site_table_full <- cinc_site_table_full %>%
  arrange(site_id, sensor_id, date, time)

library(ggplot2)

ggplot(cinc_site_table_full, aes(x = date, y = sensor_id)) +
  geom_point(size = 0.5, alpha = 0.6) +
  labs(
    title = "CinC Sensor Data Coverage",
    x = "Date",
    y = "Sensor ID"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8)
  )

ggplot(cinc_site_table_full, aes(x = date, y = site_id)) +
  geom_point(size = 0.5, alpha = 0.6) +
  labs(
    title = "CinC Coverage by Combined Site",
    x = "Date",
    y = "Site ID"
  ) +
  theme_minimal()
```


# CinC Exploratory Data Analysis 
```{r}
# CinC devices present
cinc_pm25 %>%
  distinct(device_number, location_name) %>%
  arrange(device_number)

# Number of observations per device
cinc_pm25 %>%
  count(device_number, location_name, sort = TRUE)

# missing PM2.5 values (filtered NAs)
sum(is.na(cinc_pm25$pm_2.5))

# PM2.5 statistics (Detect outliers or unrealistic values)
summary(cinc_pm25$pm_2.5)

# PM2.5 distribution (Histogram) - [if values are skewed/there are extreme spikes]
library(ggplot2)

ggplot(cinc_pm25, aes(x = pm_2.5)) +
  geom_histogram(bins = 50) +
  labs(
    title = "Distribution of PM2.5 Measurements (CLEANinCLE)",
    x = "PM2.5 (µg/m³)",
    y = "Frequency"
  )

# Time coverage check
range(cinc_pm25$time)

# Time series preview (to confirm the time data behaves correctly)
library(ggplot2)

ggplot(
  cinc_pm25 %>% filter(device_number == unique(device_number)[1]),
  aes(time, pm_2.5)
) +
  geom_point(alpha = 0.4, size = 0.7) +
  scale_x_datetime(
    date_breaks = "1 month",
    date_labels = "%b %Y"
  ) +
  labs(
    title = "PM2.5 Time Series (One CinC Sensor)",
    x = "Time",
    y = "PM2.5 (µg/m³)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(
  cinc_pm25 %>% filter(device_number == unique(device_number)[1]),
  aes(time, pm_2.5)
) +
  geom_point(alpha = 0.4, size = 0.7) +
  scale_x_datetime(
    limits = as.POSIXct(c("2024-11-01", "2025-12-31")),
    date_breaks = "1 month",
    date_labels = "%b %Y"
  ) +
  labs(
    title = "PM2.5 Time Series (One CinC Sensor)",
    x = "Time",
    y = "PM2.5 (µg/m³)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

cinc_pm25 %>%
  filter(device_number == unique(device_number)[1]) %>%
  count(month = lubridate::floor_date(time, "month"))
```
# CinC Time Series Plot 0f PM2.5 for all devices
```{r}
library(ggplot2)
library(dplyr)

# get all device numbers
device_ids <- sort(unique(cinc_pm25$device_number))

# create folder to save plots
dir.create("CinC_PM25_timeseries_plots", showWarnings = FALSE)

for (dev in device_ids) {
  
  sensor_data <- cinc_pm25 %>% filter(device_number == dev)
  loc_name <- unique(sensor_data$location_name)[1]
  
  p <- ggplot(sensor_data, aes(time, pm_2.5)) +
    geom_point(alpha = 0.4, size = 0.7) +
    scale_x_datetime(
      limits = as.POSIXct(c("2024-11-01", "2025-12-31")),
      date_breaks = "1 month",
      date_labels = "%b %Y"
    ) +
    labs(
      title = paste("PM2.5 Time Series - Device", dev, "-", loc_name),
      x = "Time",
      y = "PM2.5 (µg/m³)"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  print(p)
  
  ggsave(
    filename = paste0("CinC_PM25_timeseries_plots/CinC_sensor_", dev, ".png"),
    plot = p,
    width = 10,
    height = 5,
    dpi = 300
  )
}
```

# Cleaning bad sites / bad periods
```{r}
#clean dataset
cinc_clean <- cinc_site_table_full
cinc_removed <- tibble()

# Remove obvious bad values (>250)
bad_high <- cinc_clean %>%
  filter(pm25 > 250)

cinc_removed <- bind_rows(cinc_removed, bad_high)

cinc_clean2 <- cinc_clean %>%
  filter(pm25 <= 250)

# Remove site 6.1 completely
bad_6 <- cinc_clean %>% filter(site_id == "6.1")

cinc_removed <- bind_rows(cinc_removed, bad_6)

cinc_clean3 <- cinc_clean2 %>% filter(site_id != "6.1")

# Trim 18.2 (late May onward)
bad_18 <- cinc_clean %>%
  filter(site_id == "18.2" & date >= as.Date("2025-05-20"))

cinc_removed <- bind_rows(cinc_removed, bad_18)

cinc_clean4 <- cinc_clean3 %>%
  filter(!(site_id == "18.2" & date >= as.Date("2025-05-20")))

# Trim 27.1 (late period)
bad_27 <- cinc_clean %>%
  filter(site_id == "27.1" & date >= as.Date("2025-07-01"))

cinc_removed <- bind_rows(cinc_removed, bad_27)

cinc_clean5 <- cinc_clean4 %>%
  filter(!(site_id == "27.1" & date >= as.Date("2025-07-01")))

# Remove 27.3
bad_273 <- cinc_clean %>% filter(site_id == "27.3")

cinc_removed <- bind_rows(cinc_removed, bad_273)

cinc_clean6 <- cinc_clean5 %>% filter(site_id != "27.3")

# Trim 13.1 (late Sept)
bad_13 <- cinc_clean %>%
  filter(site_id == "13.1" & date >= as.Date("2025-09-20"))

cinc_removed <- bind_rows(cinc_removed, bad_13)

cinc_clean <- cinc_clean %>%
  filter(!(site_id == "13.1" & date >= as.Date("2025-09-20")))

distinct(cinc_clean, site_id)
cinc_removed %>% count(site_id)
```

# CLEANinCLE site-to-site matrix plot 
```{r}
#
cinc_daily <- cinc_clean %>%
  group_by(site_id, date) %>%
  summarise(pm25 = mean(pm25, na.rm = TRUE), .groups = "drop")

# Convert to wide format
library(tidyr)

cinc_wide <- cinc_daily %>%
  pivot_wider(
    names_from = site_id,
    values_from = pm25
  )

# check
names(cinc_wide)

site_order <- c(
  "1.1","2.1","3.1","5.1_19.3_24.2","7.1_4.1","7.2","8.1_34.2",
  "9.1_17.2","9.2",
  "11.1_32.1_18.2","12.1_5.2","13.1_33.2","14.1","15.1",
  "18.1_26.1","19.1","19.2","20.1","20.2","21.1","22.1","23.1",
  "24.1_1.2","25.1","27.1","28.1","29.1","30.1_17.1","31.2",
  "32.2","33.1","34.1_16.1","35.1","35.2"
)



valid_sites <- intersect(site_order, names(cinc_wide))

cinc_wide <- cinc_wide %>%
  select(date, all_of(valid_sites))
names(cinc_wide)
```
```{r}
# choose the sites you want to compare
sites_use <- valid_sites

# optional: if you want to test with fewer first, use something like:
# sites_use <- c("1.1","2.1","3.1","9.2","14.1","27.1")

df_matrix <- cinc_wide %>%
  select(date, all_of(sites_use))

library(ggplot2)
library(gridExtra)
library(dplyr)

axis_min <- 0
axis_max <- ceiling(max(unlist(df_matrix[sites_use]), na.rm = TRUE) / 5) * 5

plot_matrix <- vector("list", length(sites_use) * length(sites_use))

for (i in seq_along(sites_use)) {
  for (j in seq_along(sites_use)) {
    
    x_site <- sites_use[j]
    y_site <- sites_use[i]
    index <- (i - 1) * length(sites_use) + j
    
    df_pair <- df_matrix %>%
      select(x = all_of(x_site), y = all_of(y_site)) %>%
      filter(!is.na(x) & !is.na(y))
    
    if (nrow(df_pair) < 10) {
      p <- ggplot() +
        annotate("text", x = 0.5, y = 0.5, label = "No overlap", size = 4) +
        theme_void()
    } else {
      
      fit <- lm(y ~ x, data = df_pair)
      r2 <- summary(fit)$r.squared
      slope <- coef(fit)[2]
      intercept <- coef(fit)[1]
      
      label_text <- paste0(
        "y = ", round(slope, 2), "x + ", round(intercept, 2),
        "\nR² = ", round(r2, 2)
      )
      
      p <- ggplot(df_pair, aes(x = x, y = y)) +
        geom_point(size = 0.8, alpha = 0.6, color = "darkgreen") +
        geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
        geom_smooth(method = "lm", se = FALSE, color = "steelblue", linewidth = 0.5) +
        coord_fixed(xlim = c(axis_min, axis_max), ylim = c(axis_min, axis_max)) +
        annotate(
          "text",
          x = axis_min + 0.05 * (axis_max - axis_min),
          y = axis_max - 0.05 * (axis_max - axis_min),
          label = label_text,
          hjust = 0, vjust = 1,
          size = 2.5
        ) +
        labs(
          x = paste("Site", x_site),
          y = paste("Site", y_site)
        ) +
        theme_minimal(base_size = 8) +
        theme(
          axis.title = element_text(size = 7),
          axis.text = element_text(size = 6),
          panel.grid.minor = element_blank(),
          plot.margin = margin(2, 2, 2, 2)
        )
    }
    
    plot_matrix[[index]] <- p
  }
}

grid_plot <- grid.arrange(
  grobs = plot_matrix,
  ncol = length(sites_use),
  nrow = length(sites_use),
  top = "CinC Site-to-Site Matrix Plot (Daily PM2.5)"
)

print(grid_plot)
```


```{r}

site_lookup_clean <- site_lookup_clean %>%
  filter(sensor_id != "27.1") %>%
  bind_rows(
    tibble(sensor_id = "27.1", site_id = "27.1"),
    tibble(sensor_id = "9.2",  site_id = "9.2")
  )

cinc_site_table_full <- cinc_sites %>%
  mutate(sensor_id = as.character(Site_ID)) %>%
  left_join(site_lookup_clean, by = "sensor_id") %>%
  filter(!is.na(site_id)) %>%
  mutate(
    latitude = as.numeric(latitude),
    longitude = as.numeric(longitude),
    date = as.Date(datetime),
    time_only = format(datetime, "%H:%M:%S"),
    pm25 = pm_2.5
  ) %>%
  select(
    sensor_id,
    site_id,
    latitude,
    longitude,
    date,
    time = time_only,
    pm25
  ) %>%
  arrange(latitude, longitude, date, time)

site_order <- c(
  "1.1","2.1","3.1","5.1_19.3_24.2","7.1_4.1","7.2","8.1_34.2",
  "9.1_17.2","9.2","11.1_32.1_18.2","12.1_5.2","13.1_33.2","14.1","15.1",
  "18.1_26.1","19.1","19.2","20.1","20.2","21.1","22.1","23.1",
  "24.1_1.2","25.1","27.1","28.1","29.1","30.1_17.1","31.2",
  "32.2","33.1","34.1_16.1","35.1","35.2"
)

coverage_site_sensor <- coverage_site_sensor %>%
  filter(site_id %in% site_order) %>%
  mutate(site_id = factor(site_id, levels = rev(site_order)))

distinct(cinc_site_table_full, site_id)

```

# CinC Sensor Data Coverage by Site ID
```{r}
library(dplyr)
library(ggplot2)
library(tibble)

# 1. Fix lookup so 27.1 and 9.2 are separate
site_lookup_clean <- site_lookup_clean %>%
  filter(!(sensor_id %in% c("27.1", "9.2") | site_id == "27.1_9.2")) %>%
  bind_rows(
    tibble(sensor_id = "27.1", site_id = "27.1"),
    tibble(sensor_id = "9.2",  site_id = "9.2")
  ) %>%
  filter(grepl("^[0-9]+\\.[0-9]+(_[0-9]+\\.[0-9]+)*$", site_id))

site_lookup_clean <- site_lookup_clean %>%
  filter(site_id != "22.1") %>%
  bind_rows(
    tibble(sensor_id = "22.1", site_id = "22.1")
  )

# 2. Rebuild full table from corrected lookup
cinc_site_table_full <- cinc_sites %>%
  mutate(sensor_id = as.character(Site_ID)) %>%
  left_join(site_lookup_clean, by = "sensor_id") %>%
  filter(!is.na(site_id)) %>%
  mutate(
    latitude = as.numeric(latitude),
    longitude = as.numeric(longitude),
    date = as.Date(datetime),
    time_only = format(datetime, "%H:%M:%S"),
    pm25 = pm_2.5
  ) %>%
  select(
    sensor_id,
    site_id,
    latitude,
    longitude,
    date,
    time = time_only,
    pm25
  ) %>%
  arrange(latitude, longitude, date, time)

# 3. Confirm 27.1 and 9.2 now exist separately
distinct(cinc_site_table_full, sensor_id, site_id) %>%
  filter(sensor_id %in% c("27.1", "9.2")) %>%
  print(n = Inf)
distinct(cinc_site_table_full, sensor_id, site_id) %>%
  filter(site_id == "22.1") %>%
  print(n = Inf)

# 4. Daily coverage table
coverage_site_sensor <- cinc_site_table_full %>%
  group_by(site_id, sensor_id, date) %>%
  summarise(.groups = "drop")

# 5. Ordered from lowest to highest
site_order <- c(
  "1.1","2.1","3.1","5.1_19.3_24.2","7.1_4.1","7.2","8.1_34.2",
  "9.1_17.2","9.2",
  "11.1_32.1_18.2","12.1_5.2","13.1_33.2","14.1","15.1",
  "18.1_26.1","19.1","19.2","20.1","20.2","21.1","22.1","23.1",
  "24.1_1.2","25.1","27.1","28.1","29.1","30.1_17.1","31.2",
  "32.2","33.1","34.1_16.1","35.1","35.2"
)

# keep only site IDs that actually exist
valid_sites <- intersect(site_order, unique(coverage_site_sensor$site_id))

coverage_site_sensor <- coverage_site_sensor %>%
  filter(site_id %in% valid_sites) %>%
  mutate(site_id = factor(site_id, levels = rev(valid_sites)))

# 6. Plot with dots so you can see real daily coverage
ggplot(coverage_site_sensor, aes(x = date, y = site_id)) +
  geom_point(size = 0.45, color = "black") +
  labs(
    title = "CinC Sensor Data Coverage by Site ID",
    x = "Date",
    y = "Site ID"
  ) +
  scale_x_date(
    limits = c(as.Date("2024-11-01"), as.Date("2025-12-31")),
    date_breaks = "1 month",
    date_labels = "%b %Y"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 7),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank()
  )

cinc_site_table <- cinc_site_table_full

write.csv(cinc_site_table, "cinc_site_table.csv", row.names = FALSE)
```



```{r}
# Sensor coverage summary (start, end, number of observations)
cinc_pm25 %>%
  mutate(month = lubridate::floor_date(time, "month")) %>%
  count(device_number, location_name, month) %>%
  arrange(device_number, month)

library(gt)

cinc_summary <- cinc_pm25 %>%
  group_by(device_number, location_name) %>%
  summarise(
    start_time = min(time, na.rm = TRUE),
    end_time = max(time, na.rm = TRUE),
    n_obs = n(),
    .groups = "drop"
  ) %>%
  arrange(device_number)

gt_table <- gt(cinc_summary)

gtsave(gt_table, "CLEANinCLE_sensor_summary.pdf")

# Monthly coverage for all sensors (shows how many measurements each sensor has per month)
cinc_pm25 %>%
  mutate(month = lubridate::floor_date(time, "month")) %>%
  count(device_number, location_name, month) %>%
  arrange(device_number, month)

cinc_pm25 %>%
  mutate(
    month = lubridate::floor_date(time, "month"),
    device_number = factor(device_number)
  ) %>%
  distinct(device_number, location_name, month) %>%
  ggplot(aes(x = month, y = device_number)) +
  geom_point(size = 2) +
  scale_x_datetime(
    limits = as.POSIXct(c("2024-11-01", "2025-12-31")),
    date_breaks = "1 month",
    date_labels = "%b %Y"
  ) +
  labs(
    title = "CLEANinCLE Sensor Data Coverage",
    x = "Month",
    y = "Sensor"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Daily PM2.5 datasets
```{r}
cinc_daily <- cinc_pm25 %>%
  group_by(device_number, date) %>%
  summarise(
    pm25_daily = mean(pm_2.5, na.rm = TRUE),
    n_obs = n(),
    .groups = "drop"
  ) %>%
  arrange(device_number, date)

# check
dim(cinc_daily)
head(cinc_daily)

# Convert to WIDE format (To compare sensors, each sensor needs its own column)
library(tidyr)

cinc_daily_wide <- cinc_daily %>%
  select(device_number, date, pm25_daily) %>%
  pivot_wider(
    names_from = device_number,
    values_from = pm25_daily,
    names_prefix = "device_"
  )

head(cinc_daily_wide)

#Pairwise correlations between sensors (This tells which sensors behave similarly, which sensors disagree)
# High correlation → sensors measure similar PM patterns.
cor_matrix <- cor(
  cinc_daily_wide %>% select(-date),
  use = "pairwise.complete.obs"
)

cor_matrix
write.csv(cor_matrix, "CinC_sensor_correlation_matrix.csv")

#Visualize sensor agreement (This produces a sensor correlation heatmap)
library(ggplot2)
library(tidyr)

cor_df <- as.data.frame(cor_matrix)

cor_long <- cor_df %>%
  tibble::rownames_to_column("Sensor1") %>%
  pivot_longer(-Sensor1, names_to = "Sensor2", values_to = "correlation")

ggplot(cor_long, aes(Sensor1, Sensor2, fill = correlation)) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(
    title = "Correlation Between CinC Sensors (Daily PM2.5)",
    x = "Sensor",
    y = "Sensor"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
library(pheatmap)

pheatmap(
  cor_matrix,
  main = "Correlation Between CinC Sensors (Daily PM2.5)"
)

pdf("CinC_sensor_correlation_heatmap.pdf", width = 10, height = 8)

dev.off()

# Extract sensor coordinates

library(dplyr)

sensor_locations <- cinc_pm25 %>%
  select(device_number, latitude, longitude) %>%
  distinct() %>%
  mutate(
    latitude = as.numeric(latitude),
    longitude = as.numeric(longitude),
    sensor = paste0("device_", device_number)
  ) %>%
  arrange(device_number)

sensor_locations
str(sensor_locations)










sensor_locations <- cinc_pm25 %>%
  select(device_number, latitude, longitude) %>%
  distinct() %>%
  arrange(device_number)

sensor_locations

# distance between sensor
library(geosphere)

coords <- sensor_locations %>%
  select(longitude, latitude)

dist_matrix <- distm(coords, fun = distHaversine) / 1000

library(tidyr)

cor_long <- as.data.frame(as.table(cor_matrix))

names(cor_long) <- c("sensor1","sensor2","correlation")

sensor_compare <- cor_long %>%
  left_join(dist_long, by = c("sensor1","sensor2"))

head(cor_long)

library(ggplot2)

ggplot(sensor_compare, aes(distance_km, correlation)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", color = "red") +
  labs(
    title = "Sensor Similarity vs Distance",
    x = "Distance Between Sensors (km)",
    y = "Correlation of Daily PM2.5"
  ) +
  theme_minimal()
```

```{r}
# Create daily averages per site
cinc_daily <- cinc_site_table_full %>%
  group_by(site_id, date) %>%
  summarise(pm25 = mean(pm25, na.rm = TRUE), .groups = "drop")

# Convert to wide format
library(tidyr)

cinc_wide <- cinc_daily %>%
  pivot_wider(names_from = site_id, values_from = pm25)

library(ggplot2)

ggplot(cinc_wide, aes(x = `1.1`, y = `2.1`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Site 1.1", y = "Site 2.1")

sites <- setdiff(names(cinc_wide), "date")

library(ggplot2)
library(gridExtra)

plot_list <- list()

for (i in 1:length(sites)) {
  for (j in 1:length(sites)) {
    
    x_site <- sites[i]
    y_site <- sites[j]
    
    df_pair <- cinc_wide %>%
      select(all_of(c("date", x_site, y_site))) %>%
      na.omit()
    
    if (nrow(df_pair) > 5) {
      
      r2 <- cor(df_pair[[x_site]], df_pair[[y_site]])^2
      
      p <- ggplot(df_pair, aes(x = .data[[x_site]], y = .data[[y_site]])) +
        geom_point(size = 0.5) +
        geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
        geom_smooth(method = "lm", se = FALSE) +
        annotate("text", x = Inf, y = -Inf,
                 label = paste0("R²=", round(r2, 2)),
                 hjust = 1.1, vjust = -0.5, size = 2.5) +
        labs(x = x_site, y = y_site) +
        theme_minimal(base_size = 8)
      
    } else {
      
      p <- ggplot() + theme_void()
    }
    
    plot_list[[length(plot_list) + 1]] <- p
  }
}

grid.arrange(grobs = plot_list, ncol = length(sites))

```


```{r cinc-site-matrix-all, fig.width=38, fig.height=38, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)

# daily averages by combined CinC site
cinc_daily_site <- cinc_site_table_full %>%
  filter(grepl("^[0-9]+\\.[0-9]+(_[0-9]+\\.[0-9]+)*$", site_id)) %>%
  group_by(site_id, date) %>%
  summarise(
    pm25 = mean(pm25, na.rm = TRUE),
    .groups = "drop"
  )

# site order
site_order <- c(
  "1.1","2.1","3.1","5.1_19.3_24.2","7.1_4.1","7.2","8.1_34.2",
  "9.1_17.2","11.1_32.1_18.2","12.1_5.2","13.1_33.2","14.1","15.1",
  "18.1_26.1","19.1","19.2","20.1","20.2","21.1","23.1",
  "24.1_1.2","25.1","27.1_9.2","28.1","29.1","30.1_17.1","31.2",
  "32.2","33.1","34.1_16.1","35.1","35.2"
)

cinc_daily_site <- cinc_daily_site %>%
  filter(site_id %in% site_order)

# wide table
cinc_wide <- cinc_daily_site %>%
  pivot_wider(
    names_from = site_id,
    values_from = pm25,
    values_fn = mean
  )

sites <- site_order[site_order %in% names(cinc_wide)]
n_sites <- length(sites)

# better common axis limit
all_vals <- unlist(cinc_wide[sites], use.names = FALSE)
all_vals <- all_vals[is.finite(all_vals)]

# use a high percentile instead of the absolute max so one outlier
# does not flatten all the other panels
axis_min <- 0
axis_max <- ceiling(as.numeric(quantile(all_vals, 0.99, na.rm = TRUE)) / 5) * 5

plot_matrix <- vector("list", n_sites * n_sites)

for (i in seq_len(n_sites)) {
  for (j in seq_len(n_sites)) {

    x_site <- sites[j]
    y_site <- sites[i]
    index <- (i - 1) * n_sites + j

    df_pair <- cinc_wide %>%
      select(x = all_of(x_site), y = all_of(y_site)) %>%
      filter(!is.na(x), !is.na(y))

    # default labels
    x_lab <- if (i == n_sites) x_site else NULL
    y_lab <- if (j == 1) y_site else NULL

    if (nrow(df_pair) < 3) {

      p <- ggplot() +
        annotate(
          "text",
          x = 0.5, y = 0.5,
          label = "No overlap",
          size = 2.2
        ) +
        xlim(0, 1) + ylim(0, 1) +
        labs(x = x_lab, y = y_lab) +
        theme_minimal(base_size = 6) +
        theme(
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          panel.grid = element_blank(),
          axis.title.x = element_text(size = 5),
          axis.title.y = element_text(size = 5),
          plot.margin = margin(1, 1, 1, 1)
        )

    } else {

      model <- lm(y ~ x, data = df_pair)
      r2 <- summary(model)$r.squared
      slope <- coef(model)[2]
      intercept <- coef(model)[1]

      label_text <- sprintf(
        "y = %.2fx %+ .2f\nR² = %.2f",
        slope, intercept, r2
      )

      x_pos <- axis_min + 0.05 * (axis_max - axis_min)
      y_pos <- axis_max - 0.05 * (axis_max - axis_min)

      p <- ggplot(df_pair, aes(x = x, y = y)) +
        geom_point(size = 0.35, alpha = 0.55, color = "darkgreen") +
        geom_abline(
          slope = 1, intercept = 0,
          linetype = "dashed", color = "gray50", linewidth = 0.25
        ) +
        geom_smooth(
          method = "lm", se = FALSE,
          color = "steelblue", linewidth = 0.35
        ) +
        coord_fixed(
          xlim = c(axis_min, axis_max),
          ylim = c(axis_min, axis_max),
          expand = FALSE
        ) +
        annotate(
          "text",
          x = x_pos, y = y_pos,
          label = label_text,
          hjust = 0, vjust = 1,
          size = 1.4
        ) +
        labs(x = x_lab, y = y_lab) +
        theme_minimal(base_size = 6) +
        theme(
          axis.title.x = element_text(size = 5),
          axis.title.y = element_text(size = 5),
          axis.text = element_text(size = 3.5),
          panel.grid.minor = element_blank(),
          plot.margin = margin(1, 1, 1, 1)
        )
    }

    plot_matrix[[index]] <- p
  }
}

grid_plot <- grid.arrange(
  grobs = plot_matrix,
  ncol = n_sites,
  nrow = n_sites,
  top = "CinC Site-to-Site Matrix Plot (Daily PM2.5)"
)

# save large
ggsave(
  "cinc_site_matrix_daily.pdf",
  plot = grid_plot,
  width = 34,
  height = 34
)

ggsave(
  "cinc_site_matrix_daily.png",
  plot = grid_plot,
  width = 34,
  height = 34,
  dpi = 300
)
```


```{r}

library(dplyr)

sensor_locations <- cinc_pm25 %>%
  select(device_number, latitude, longitude) %>%
  distinct() %>%
  mutate(
    latitude = as.numeric(latitude),
    longitude = as.numeric(longitude),
    sensor = paste0("device_", device_number)
  ) %>%
  arrange(device_number)

sensor_locations
str(sensor_locations)

library(geosphere)

coords <- sensor_locations %>%
  select(longitude, latitude) %>%
  as.matrix()

dist_matrix <- distm(coords, fun = distHaversine) / 1000

rownames(dist_matrix) <- sensor_locations$sensor
colnames(dist_matrix) <- sensor_locations$sensor

dist_long <- as.data.frame(as.table(dist_matrix))
colnames(dist_long) <- c("sensor1", "sensor2", "distance_km")

head(dist_long)

cor_long <- as.data.frame(as.table(cor_matrix))
colnames(cor_long) <- c("sensor1", "sensor2", "correlation")

head(cor_long)

sensor_compare <- cor_long %>%
  left_join(dist_long, by = c("sensor1", "sensor2")) %>%
  filter(sensor1 != sensor2)

head(sensor_compare)

library(ggplot2)

ggplot(sensor_compare, aes(x = distance_km, y = correlation)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm") +
  labs(
    title = "Correlation vs Distance Between CinC Sensors",
    x = "Distance Between Sensors (km)",
    y = "PM2.5 Correlation"
  ) +
  theme_minimal()
```

```{r}

sensor_locations %>%
  count(device_number)

range(sensor_locations$latitude)
range(sensor_locations$longitude)
```

```{r}
sensor_locations_clean <- cinc_pm25 %>%
  mutate(
    latitude = as.numeric(latitude),
    longitude = as.numeric(longitude)
  ) %>%
  filter(
    !is.na(latitude),
    !is.na(longitude),
    latitude != 0,
    longitude != 0
  ) %>%
  group_by(device_number) %>%
  summarise(
    latitude = first(latitude),
    longitude = first(longitude),
    .groups = "drop"
  ) %>%
  mutate(
    sensor = paste0("device_", device_number)
  ) %>%
  arrange(as.numeric(device_number))

sensor_locations_clean %>%
  count(device_number)

range(sensor_locations_clean$latitude)
range(sensor_locations_clean$longitude)

sensor_locations_clean

library(geosphere)

coords <- sensor_locations_clean %>%
  select(longitude, latitude) %>%
  as.matrix()

dist_matrix <- distm(coords, fun = distHaversine) / 1000

rownames(dist_matrix) <- sensor_locations_clean$sensor
colnames(dist_matrix) <- sensor_locations_clean$sensor

dist_long <- as.data.frame(as.table(dist_matrix))
colnames(dist_long) <- c("sensor1", "sensor2", "distance_km")

head(dist_long)

cor_long <- as.data.frame(as.table(cor_matrix))
colnames(cor_long) <- c("sensor1", "sensor2", "correlation")

sensor_compare <- cor_long %>%
  left_join(dist_long, by = c("sensor1", "sensor2")) %>%
  filter(sensor1 != sensor2)

head(sensor_compare)

library(ggplot2)

ggplot(sensor_compare, aes(x = distance_km, y = correlation)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm") +
  labs(
    title = "Correlation vs Distance Between CinC Sensors",
    x = "Distance Between Sensors (km)",
    y = "PM2.5 Correlation"
  ) +
  theme_minimal()

sensor_locations_clean %>% count(device_number)
range(sensor_locations_clean$latitude)
range(sensor_locations_clean$longitude)
```

```{r}
library(ggplot2)
library(maps)
library(dplyr)

county_map <- map_data("county")
cuyahoga_map <- county_map %>%
  filter(region == "ohio", subregion == "cuyahoga")

ggplot() +
  geom_polygon(
    data = cuyahoga_map,
    aes(x = long, y = lat, group = group),
    fill = "gray95",
    color = "black"
  ) +
  geom_point(
    data = sensor_locations_clean,
    aes(x = longitude, y = latitude),
    color = "firebrick",
    size = 3
  ) +
  coord_fixed(
    xlim = c(-81.85, -81.45),
    ylim = c(41.35, 41.65)
  ) +
  labs(
    title = "CLEANinCLE Sensor Locations (Cuyahoga County)",
    x = "Longitude",
    y = "Latitude"
  ) +
  theme_minimal()
```

# make one clean location per sensor

```{r}
sensor_locations_clean <- cinc_pm25 %>%
  mutate(
    latitude = as.numeric(latitude),
    longitude = as.numeric(longitude)
  ) %>%
  filter(
    !is.na(latitude),
    !is.na(longitude),
    latitude != 0,
    longitude != 0
  ) %>%
  group_by(device_number) %>%
  summarise(
    latitude = first(latitude),
    longitude = first(longitude),
    .groups = "drop"
  ) %>%
  mutate(sensor = paste0("device_", device_number))

# make the distance matrix
library(geosphere)
library(dplyr)

coords <- sensor_locations_clean %>%
  select(longitude, latitude) %>%
  as.matrix()

dist_matrix <- distm(coords, fun = distHaversine) / 1000

rownames(dist_matrix) <- sensor_locations_clean$sensor
colnames(dist_matrix) <- sensor_locations_clean$sensor

# turn both matrices into long tables
library(tidyr)

cor_long <- as.data.frame(as.table(cor_matrix))
colnames(cor_long) <- c("sensor1", "sensor2", "correlation")

dist_long <- as.data.frame(as.table(dist_matrix))
colnames(dist_long) <- c("sensor1", "sensor2", "distance_km")

# join them
sensor_compare <- cor_long %>%
  left_join(dist_long, by = c("sensor1", "sensor2")) %>%
  filter(sensor1 != sensor2)

sensor_compare <- sensor_compare %>%
  mutate(
    sensor1 = as.character(sensor1),
    sensor2 = as.character(sensor2)
  )

# remove duplicate pairs
sensor_compare_unique <- sensor_compare %>%
  rowwise() %>%
  mutate(
    pair1 = min(sensor1, sensor2),
    pair2 = max(sensor1, sensor2)
  ) %>%
  ungroup() %>%
  distinct(pair1, pair2, .keep_all = TRUE)

# plot correlation vs distance
library(ggplot2)

ggplot(sensor_compare_unique, aes(distance_km, correlation)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = "Correlation vs Distance Between CLEANinCLE Sensors",
    x = "Distance between sensors (km)",
    y = "Correlation of daily PM2.5"
  ) +
  theme_minimal()

```

# sensors with unusually low correlation

```{r}
sensor_compare_unique %>%
  filter(correlation < 0.4) %>%
  arrange(correlation)

# average correlation by distance group
sensor_compare_unique %>%
  mutate(distance_group = cut(
    distance_km,
    breaks = c(0,2,5,10,20,30),
    labels = c("0–2 km","2–5 km","5–10 km","10–20 km","20–30 km")
  )) %>%
  group_by(distance_group) %>%
  summarise(mean_correlation = mean(correlation, na.rm = TRUE)) 
```

# pairwise comparison for CLEANinCLE sites using daily averages.

```{r}
# daily averages for each CinC sensor
library(dplyr)
library(tidyr)
library(ggplot2)
library(reshape2)
library(geosphere)

# Daily averages for each CinC sensor
cinc_daily <- cinc_pm25 %>%
  mutate(date = as.Date(time)) %>%
  group_by(device_number, date) %>%
  summarise(
    pm25_daily = mean(pm_2.5, na.rm = TRUE),
    n_obs = n(),
    .groups = "drop"
  ) %>%
  arrange(device_number, date)

head(cinc_daily)

# Convert to wide format
cinc_daily_wide <- cinc_daily %>%
  pivot_wider(
    names_from = device_number,
    values_from = pm25_daily,
    names_prefix = "device_"
  )

head(cinc_daily_wide)

# pairwise correlation matrix
cinc_cor_matrix <- cor(
  cinc_daily_wide %>% select(-date),
  use = "pairwise.complete.obs"
)

round(cinc_cor_matrix, 2)

# Plot the CinC correlation heatmap
cinc_cor_long <- melt(cinc_cor_matrix)

ggplot(cinc_cor_long, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_viridis_c(name = "Correlation") +
  labs(
    title = "Correlation Between CinC Sensors (Daily PM2.5)",
    x = "Sensor",
    y = "Sensor"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  )

# clean sensor location table
sensor_locations_clean <- cinc_pm25 %>%
  mutate(
    latitude = as.numeric(latitude),
    longitude = as.numeric(longitude)
  ) %>%
  filter(
    !is.na(latitude),
    !is.na(longitude),
    latitude != 0,
    longitude != 0
  ) %>%
  group_by(device_number) %>%
  summarise(
    latitude = first(latitude),
    longitude = first(longitude),
    .groups = "drop"
  ) %>%
  mutate(sensor = paste0("device_", device_number)) %>%
  arrange(sensor)

sensor_locations_clean

# distance matrix between CinC sensors
coords <- sensor_locations_clean %>%
  select(longitude, latitude) %>%
  as.matrix()

dist_matrix <- distm(coords, fun = distHaversine) / 1000

rownames(dist_matrix) <- sensor_locations_clean$sensor
colnames(dist_matrix) <- sensor_locations_clean$sensor

# correlation and distance to long format
cor_long <- as.data.frame(as.table(cinc_cor_matrix))
colnames(cor_long) <- c("sensor1", "sensor2", "correlation")

dist_long <- as.data.frame(as.table(dist_matrix))
colnames(dist_long) <- c("sensor1", "sensor2", "distance_km")

# Join them and remove duplicate pairs
sensor_compare <- cor_long %>%
  mutate(
    sensor1 = as.character(sensor1),
    sensor2 = as.character(sensor2)
  ) %>%
  left_join(dist_long, by = c("sensor1", "sensor2")) %>%
  filter(sensor1 != sensor2) %>%
  rowwise() %>%
  mutate(
    pair1 = min(sensor1, sensor2),
    pair2 = max(sensor1, sensor2)
  ) %>%
  ungroup() %>%
  distinct(pair1, pair2, .keep_all = TRUE) %>%
  filter(!is.na(correlation), !is.na(distance_km))

head(sensor_compare)

# correlation vs distance for CinC daily averages
ggplot(sensor_compare, aes(x = distance_km, y = correlation)) +
  geom_point(alpha = 0.35, size = 2) +
  geom_smooth(method = "lm", color = "blue", se = TRUE, linewidth = 1) +
  coord_cartesian(ylim = c(0.5, 1.02)) +
  labs(
    title = "Correlation vs Distance Between CLEANinCLE Sensors 
    (Daily PM2.5)",
    x = "Distance between sensors (km)",
    y = "Correlation"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    panel.grid.minor = element_blank()
  )

```

# Pairwise comparison for CLEANinCLE sites (Hourly averages)

```{r}
# CinC half-hour data to hourly averages

library(dplyr)
library(lubridate)

cinc_hourly <- cinc_pm25 %>%
  mutate(
    datetime = as.POSIXct(time),
    hour = floor_date(datetime, "hour")
  ) %>%
  group_by(device_number, hour) %>%
  summarise(
    pm25_hourly = mean(pm_2.5, na.rm = TRUE),
    .groups = "drop"
  )

# wide format (one column per sensor)
library(tidyr)

cinc_hourly_wide <- cinc_hourly %>%
  pivot_wider(
    names_from = device_number,
    values_from = pm25_hourly,
    names_prefix = "device_"
  )

# correlation matrix
cinc_hourly_cor <- cor(
  cinc_hourly_wide %>% select(-hour),
  use = "pairwise.complete.obs"
)

# Convert correlations to long format
library(reshape2)

cor_long <- as.data.frame(as.table(cinc_hourly_cor))
colnames(cor_long) <- c("sensor1", "sensor2", "correlation")

# Combine with distance table
sensor_compare_hourly <- cor_long %>%
  left_join(dist_long, by = c("sensor1", "sensor2")) %>%
  filter(sensor1 != sensor2)

# Plot correlation vs distance (hourly)
library(ggplot2)

ggplot(sensor_compare_hourly, aes(distance_km, correlation)) +
  geom_point(alpha = 0.35, size = 2) +
  geom_smooth(method = "lm", color = "blue", se = TRUE) +
  coord_cartesian(ylim = c(0.5, 1.02)) +
  labs(
    title = "Correlation vs Distance Between CLEANinCLE Sensors (Hourly PM2.5)",
    x = "Distance between sensors (km)",
    y = "Correlation"
  ) +
  theme_minimal(base_size = 13)

# 
```

```{r}


```



# EPA Analysis
```{r}
library(RAQSAPI)
library(purrr)
library(dplyr)

aqs_credentials(username = "rnl34@case.edu", key = "copperhawk19")

site_numbers <- c("0034", "0038", "0045", "0060", "0065", "0073")

epa_data_list <- map(site_numbers, function(site) {
  RAQSAPI::aqs_sampledata_by_site(
    parameter = "88101",
    bdate = as.Date("20241101", format = "%Y%m%d"),
    edate = as.Date("20251231", format = "%Y%m%d"),
    stateFIPS = "39",
    countycode = "035",
    sitenum = site,
    return_header = FALSE
  ) %>%
    mutate(site_number = site)
})

epa_data <- bind_rows(epa_data_list)

head(epa_data)
dim(epa_data)

# EPA FRM and EPA FEM
library(dplyr)

epa_data <- epa_data %>%
  mutate(date = as.Date(date_local))

# Check
epa_data %>%
  count(method_type, sample_duration)

# FRM = filter, 24-hour
epa_frm_daily <- epa_data %>%
  filter(method_type == "FRM", sample_duration == "24 HOUR") %>%
  select(site_number, date, sample_measurement) %>%
  rename(frm_pm25 = sample_measurement) %>%
  arrange(site_number, date)

head(epa_frm_daily)

# FEM = continuous/optical, hourly data averaged to daily
epa_fem_hourly <- epa_data %>%
  filter(method_type == "FEM", sample_duration == "1 HOUR")

epa_fem_daily <- epa_fem_hourly %>%
  group_by(site_number, date) %>%
  summarise(
    fem_pm25_daily = mean(sample_measurement, na.rm = TRUE),
    n_obs = n(),
    .groups = "drop"
  ) %>%
  arrange(site_number, date)

head(epa_fem_daily)

# 
```

```{r}
# Filter daily 24-hour EPA data
epa_filter <- epa_data %>%
  filter(sample_duration == "24 HOUR") %>%
  group_by(site_number, date) %>%
  summarise(
    pm25 = mean(sample_measurement, na.rm = TRUE),
    .groups = "drop"
  )

# Check
head(epa_filter)
distinct(epa_filter, site_number)


#
library(tidyr)

epa_filter_wide <- epa_filter %>%
  pivot_wider(
    names_from = site_number,
    values_from = pm25,
    values_fn = mean
  )

head(epa_filter_wide)
```

```{r epa-6x6-filter-grid, fig.width=18, fig.height=18, echo=FALSE, message=FALSE, warning=FALSE}

# 
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)

# 1. EPA filter daily table
epa_filter <- epa_data %>%
  filter(sample_duration == "24 HOUR") %>%
  group_by(site_number, date) %>%
  summarise(
    pm25 = mean(sample_measurement, na.rm = TRUE),
    .groups = "drop"
  )

# 2. Wide format
epa_filter_wide <- epa_filter %>%
  mutate(site_number = as.character(site_number)) %>%
  pivot_wider(
    names_from = site_number,
    values_from = pm25,
    values_fn = mean
  )

# 3. Site names and common axis range
site_names <- sort(setdiff(names(epa_filter_wide), "date"))
n_sites <- length(site_names)

axis_min <- 0
axis_max <- ceiling(max(unlist(epa_filter_wide[site_names]), na.rm = TRUE) / 5) * 5

# 4. Build plot matrix
plot_matrix <- vector("list", n_sites * n_sites)

for (i in seq_len(n_sites)) {
  for (j in seq_len(n_sites)) {

    x_site <- site_names[j]
    y_site <- site_names[i]
    index <- (i - 1) * n_sites + j

    df_pair <- epa_filter_wide %>%
      select(x = all_of(x_site), y = all_of(y_site)) %>%
      filter(!is.na(x) & !is.na(y))

    if (nrow(df_pair) > 2) {
      model <- lm(y ~ x, data = df_pair)
      r2 <- round(summary(model)$r.squared, 2)
      slope <- round(coef(model)[2], 2)
      intercept <- round(coef(model)[1], 2)
      label_text <- paste0("y = ", slope, "x + ", intercept, "\nR² = ", r2)
    } else {
      label_text <- "No overlap"
    }

    x_pos <- axis_min + 1
    y_pos <- axis_max - 1

    p <- ggplot(df_pair, aes(x = x, y = y)) +
      geom_point(color = "darkgreen", alpha = 0.7, size = 0.9) +
      geom_smooth(method = "lm", se = FALSE, color = "steelblue", linewidth = 0.5) +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50", linewidth = 0.4) +
      coord_fixed(xlim = c(axis_min, axis_max), ylim = c(axis_min, axis_max)) +
      annotate(
        "text",
        x = x_pos, y = y_pos,
        label = label_text,
        hjust = 0, vjust = 1,
        size = 2.2
      ) +
      labs(
        x = paste("Site", x_site),
        y = paste("Site", y_site)
      ) +
      theme_minimal(base_size = 8) +
      theme(
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 6),
        panel.grid.minor = element_blank(),
        plot.margin = margin(2, 2, 2, 2)
      )

    plot_matrix[[index]] <- p
  }
}

# 5. Draw the 6x6 matrix
grid_plot <- grid.arrange(
  grobs = plot_matrix,
  ncol = n_sites,
  nrow = n_sites,
  top = "EPA Filter Site-to-Site Matrix Plot (Daily PM2.5)"
)

print(grid_plot)

ggsave("epa_filter_6x6_matrix.png", plot = grid_plot, width = 18, height = 18, dpi = 300)
```

# EPA Optical (Daily, ALL data) - Continuous monitors (FEM) averaged to daily.
```{r epa-optical-daily-6x6, fig.width=18, fig.height=18, echo=FALSE, message=FALSE, warning=FALSE}
# Create optical daily dataset
epa_optical_daily <- epa_data %>%
  filter(sample_duration != "24 HOUR") %>%   # keep continuous data
  group_by(site_number, date) %>%
  summarise(
    pm25 = mean(sample_measurement, na.rm = TRUE),
    .groups = "drop"
  )

#Wide format
epa_optical_wide <- epa_optical_daily %>%
  mutate(site_number = as.character(site_number)) %>%
  pivot_wider(
    names_from = site_number,
    values_from = pm25,
    values_fn = mean
  )

# Site names and common axis range
site_names <- sort(setdiff(names(epa_optical_wide), "date"))
n_sites <- length(site_names)

axis_min <- 0
axis_max <- ceiling(max(unlist(epa_optical_wide[site_names]), na.rm = TRUE) / 5) * 5

# 6×6 matrix plot
plot_matrix <- vector("list", n_sites * n_sites)

for (i in seq_len(n_sites)) {
  for (j in seq_len(n_sites)) {

    x_site <- site_names[j]
    y_site <- site_names[i]
    index <- (i - 1) * n_sites + j

    df_pair <- epa_optical_wide %>%
      select(x = all_of(x_site), y = all_of(y_site)) %>%
      filter(!is.na(x) & !is.na(y))

    if (nrow(df_pair) > 2) {
      model <- lm(y ~ x, data = df_pair)
      r2 <- round(summary(model)$r.squared, 2)
      slope <- round(coef(model)[2], 2)
      intercept <- round(coef(model)[1], 2)
      label_text <- paste0("y = ", slope, "x + ", intercept, "\nR² = ", r2)
    } else {
      label_text <- "No overlap"
    }

    x_pos <- axis_min + 1
    y_pos <- axis_max - 1

    p <- ggplot(df_pair, aes(x = x, y = y)) +
      geom_point(color = "darkgreen", alpha = 0.7, size = 0.9) +
      geom_smooth(method = "lm", se = FALSE, color = "steelblue", linewidth = 0.5) +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50", linewidth = 0.4) +
      coord_fixed(xlim = c(axis_min, axis_max), ylim = c(axis_min, axis_max)) +
      annotate(
        "text",
        x = x_pos, y = y_pos,
        label = label_text,
        hjust = 0, vjust = 1,
        size = 2.2
      ) +
      labs(
        x = paste("Site", x_site),
        y = paste("Site", y_site)
      ) +
      theme_minimal(base_size = 8) +
      theme(
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 6),
        panel.grid.minor = element_blank(),
        plot.margin = margin(2, 2, 2, 2)
      )

    plot_matrix[[index]] <- p
  }
}

#
grid_plot <- grid.arrange(
  grobs = plot_matrix,
  ncol = n_sites,
  nrow = n_sites,
  top = "EPA Optical Site-to-Site Matrix Plot (Daily PM2.5)"
)

print(grid_plot)

ggsave("epa_optical_daily_6x6_matrix.png", plot = grid_plot, width = 18, height = 18, dpi = 300)
```

# EPA Optical (Daily, only filter days)
```{r epa-optical-matched-6x6, fig.width=18, fig.height=18, echo=FALSE, message=FALSE, warning=FALSE}
epa_optical_matched <- epa_optical_daily %>%
  semi_join(epa_filter, by = c("site_number", "date"))

# Convert to wide format
library(tidyr)

epa_optical_matched_wide <- epa_optical_matched %>%
  mutate(site_number = as.character(site_number)) %>%
  pivot_wider(
    names_from = site_number,
    values_from = pm25,
    values_fn = mean
  )

# Site names and common axis range
site_names <- sort(setdiff(names(epa_optical_matched_wide), "date"))
n_sites <- length(site_names)

axis_min <- 0
axis_max <- ceiling(max(unlist(epa_optical_matched_wide[site_names]), na.rm = TRUE) / 5) * 5

# 6×6 matrix plot
library(ggplot2)
library(gridExtra)
library(dplyr)

plot_matrix <- vector("list", n_sites * n_sites)

for (i in seq_len(n_sites)) {
  for (j in seq_len(n_sites)) {

    x_site <- site_names[j]
    y_site <- site_names[i]
    index <- (i - 1) * n_sites + j

    df_pair <- epa_optical_matched_wide %>%
      select(x = all_of(x_site), y = all_of(y_site)) %>%
      filter(!is.na(x) & !is.na(y))

    if (nrow(df_pair) > 2) {
      model <- lm(y ~ x, data = df_pair)
      r2 <- round(summary(model)$r.squared, 2)
      slope <- round(coef(model)[2], 2)
      intercept <- round(coef(model)[1], 2)
      label_text <- paste0("y = ", slope, "x + ", intercept, "\nR² = ", r2)
    } else {
      label_text <- "No overlap"
    }

    x_pos <- axis_min + 1
    y_pos <- axis_max - 1

    p <- ggplot(df_pair, aes(x = x, y = y)) +
      geom_point(color = "darkgreen", alpha = 0.7, size = 0.9) +
      geom_smooth(method = "lm", se = FALSE, color = "steelblue", linewidth = 0.5) +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50", linewidth = 0.4) +
      coord_fixed(xlim = c(axis_min, axis_max), ylim = c(axis_min, axis_max)) +
      annotate(
        "text",
        x = x_pos, y = y_pos,
        label = label_text,
        hjust = 0, vjust = 1,
        size = 2.2
      ) +
      labs(
        x = paste("Site", x_site),
        y = paste("Site", y_site)
      ) +
      theme_minimal(base_size = 8) +
      theme(
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 6),
        panel.grid.minor = element_blank(),
        plot.margin = margin(2, 2, 2, 2)
      )

    plot_matrix[[index]] <- p
  }
}

grid_plot <- grid.arrange(
  grobs = plot_matrix,
  ncol = n_sites,
  nrow = n_sites,
  top = "EPA Optical Site-to-Site Matrix Plot (Daily PM2.5, Filter Days Only)"
)

print(grid_plot)

ggsave("epa_optical_matched_6x6_matrix.png", plot = grid_plot, width = 18, height = 18, dpi = 300)

```


```{r epa-optical-matched-6x6, fig.width=18, fig.height=18, echo=FALSE, message=FALSE, warning=FALSE}
# hourly dataset
epa_optical_hourly <- epa_data %>%
  filter(sample_duration != "24 HOUR") %>%
  mutate(datetime = as.POSIXct(paste(date, time_local))) %>%
  group_by(site_number, datetime) %>%
  summarise(
    pm25 = mean(sample_measurement, na.rm = TRUE),
    .groups = "drop"
  )

# Wide format
epa_optical_hourly_wide <- epa_optical_hourly %>%
  mutate(site_number = as.character(site_number)) %>%
  pivot_wider(
    names_from = site_number,
    values_from = pm25,
    values_fn = mean
  )

# Site names + axis range
site_names <- sort(setdiff(names(epa_optical_hourly_wide), "datetime"))
n_sites <- length(site_names)

axis_min <- 0
axis_max <- ceiling(max(unlist(epa_optical_hourly_wide[site_names]), na.rm = TRUE) / 5) * 5

# 6×6 matrix
library(ggplot2)
library(gridExtra)
library(dplyr)

plot_matrix <- vector("list", n_sites * n_sites)

for (i in seq_len(n_sites)) {
  for (j in seq_len(n_sites)) {

    x_site <- site_names[j]
    y_site <- site_names[i]
    index <- (i - 1) * n_sites + j

    df_pair <- epa_optical_hourly_wide %>%
      select(x = all_of(x_site), y = all_of(y_site)) %>%
      filter(!is.na(x) & !is.na(y))

    if (nrow(df_pair) > 2) {
      model <- lm(y ~ x, data = df_pair)
      r2 <- round(summary(model)$r.squared, 2)
      slope <- round(coef(model)[2], 2)
      intercept <- round(coef(model)[1], 2)
      label_text <- paste0("y = ", slope, "x + ", intercept, "\nR² = ", r2)
    } else {
      label_text <- "No overlap"
    }

    x_pos <- axis_min + 1
    y_pos <- axis_max - 1

    p <- ggplot(df_pair, aes(x = x, y = y)) +
      geom_point(color = "darkgreen", alpha = 0.6, size = 0.7) +
      geom_smooth(method = "lm", se = FALSE, color = "steelblue", linewidth = 0.4) +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50", linewidth = 0.3) +
      coord_fixed(xlim = c(axis_min, axis_max), ylim = c(axis_min, axis_max)) +
      annotate(
        "text",
        x = x_pos, y = y_pos,
        label = label_text,
        hjust = 0, vjust = 1,
        size = 2
      ) +
      labs(
        x = paste("Site", x_site),
        y = paste("Site", y_site)
      ) +
      theme_minimal(base_size = 7) +
      theme(
        axis.title = element_text(size = 6),
        axis.text = element_text(size = 5),
        panel.grid.minor = element_blank(),
        plot.margin = margin(2, 2, 2, 2)
      )

    plot_matrix[[index]] <- p
  }
}

grid_plot <- grid.arrange(
  grobs = plot_matrix,
  ncol = n_sites,
  nrow = n_sites,
  top = "EPA Optical Site-to-Site Matrix Plot (Hourly PM2.5)"
)

print(grid_plot)
```

# Correlation vs Distance
```{r}
epa_coords <- epa_data %>%
  select(site_number, latitude, longitude) %>%
  distinct() %>%
  mutate(site_number = as.character(site_number))

# Compute distance matrix
library(geosphere)

dist_matrix <- distm(
  epa_coords[, c("longitude", "latitude")],
  fun = distHaversine
)

rownames(dist_matrix) <- epa_coords$site_number
colnames(dist_matrix) <- epa_coords$site_number 

# Convert to long format
library(tidyr)

dist_long <- as.data.frame(dist_matrix) %>%
  mutate(site1 = rownames(dist_matrix)) %>%
  pivot_longer(-site1, names_to = "site2", values_to = "distance_m") %>%
  mutate(distance_km = distance_m / 1000)

epa_r2 <- list()

site_names <- sort(setdiff(names(epa_filter_wide), "date"))

for (i in seq_along(site_names)) {
  for (j in seq_along(site_names)) {

    site_i <- site_names[i]
    site_j <- site_names[j]

    df_pair <- epa_filter_wide %>%
      select(x = all_of(site_i), y = all_of(site_j)) %>%
      filter(!is.na(x) & !is.na(y))

    if (nrow(df_pair) > 5) {
      r2 <- summary(lm(y ~ x, data = df_pair))$r.squared
    } else {
      r2 <- NA
    }

    epa_r2[[length(epa_r2) + 1]] <- data.frame(
      site1 = site_i,
      site2 = site_j,
      r2 = r2
    )
  }
}

epa_r2_df <- bind_rows(epa_r2)

cor_dist <- epa_r2_df %>%
  left_join(dist_long, by = c("site1", "site2")) %>%
  filter(site1 != site2)

cor_dist_clean <- cor_dist %>%
  filter(site1 < site2)   # removes duplicates + diagonal

# Plot correlation vs distance
ggplot(cor_dist, aes(x = distance_km, y = r2)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Correlation vs Distance Between EPA Sites",
    x = "Distance (km)",
    y = "R² (Correlation Strength)"
  ) +
  theme_minimal()


```
# EPA daily tables 
```{r}
library(tidyr)

# FRM daily wide
epa_frm_wide <- epa_frm_daily %>%
  mutate(site = paste0("site_", site_number)) %>%
  select(date, site, frm_pm25) %>%
  pivot_wider(
    names_from = site,
    values_from = frm_pm25
  )

head(epa_frm_wide) 

# FEM daily wide, all days
library(dplyr)
library(tidyr)

# collapse duplicate FRM rows first
epa_frm_daily_clean <- epa_frm_daily %>%
  group_by(site_number, date) %>%
  summarise(
    frm_pm25 = mean(frm_pm25, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(site_number, date)

# now make the wide table
epa_frm_wide <- epa_frm_daily_clean %>%
  mutate(site = paste0("site_", site_number)) %>%
  select(date, site, frm_pm25) %>%
  pivot_wider(
    names_from = site,
    values_from = frm_pm25
  )

# check
str(epa_frm_wide)
head(epa_frm_wide)

# FEM daily wide, FRM days only
epa_fem_on_frm_days <- epa_fem_daily %>%
  inner_join(
    epa_frm_daily %>% select(site_number, date),
    by = c("site_number", "date")
  )

epa_fem_on_frm_wide <- epa_fem_on_frm_days %>%
  mutate(site = paste0("site_", site_number)) %>%
  select(date, site, fem_pm25_daily) %>%
  pivot_wider(
    names_from = site,
    values_from = fem_pm25_daily
  )

head(epa_fem_on_frm_wide)

# Create the EPA site location table
epa_locations <- epa_data %>%
  mutate(site = paste0("site_", site_number)) %>%
  select(site, latitude, longitude) %>%
  distinct() %>%
  filter(!is.na(latitude), !is.na(longitude)) %>%
  group_by(site) %>%
  summarise(
    lat = first(latitude),
    lon = first(longitude),
    .groups = "drop"
  ) %>%
  arrange(site)

epa_locations

# Build EPA distance table
library(geosphere)

epa_coords <- epa_locations %>%
  select(lon, lat) %>%
  as.matrix()

epa_dist_matrix <- distm(epa_coords, fun = distHaversine) / 1000

rownames(epa_dist_matrix) <- epa_locations$site
colnames(epa_dist_matrix) <- epa_locations$site

epa_dist_long <- as.data.frame(as.table(epa_dist_matrix))
colnames(epa_dist_long) <- c("site1", "site2", "distance_km")

head(epa_dist_long)


# EPA pairwise matrix plots
library(ggplot2)
library(gridExtra)

make_epa_pairwise_matrix <- function(wide_df, value_label, plot_title) {
  
  site_names <- names(wide_df)[names(wide_df) != "date"]
  plots <- list()
  
  for (i in seq_along(site_names)) {
    for (j in seq_along(site_names)) {
      
      site_i <- site_names[i]
      site_j <- site_names[j]
      
      df <- wide_df %>%
        select(date, all_of(site_i), all_of(site_j)) %>%
        rename(x = all_of(site_i), y = all_of(site_j)) %>%
        filter(!is.na(x), !is.na(y))
      
      if (nrow(df) > 2) {
        fit <- lm(y ~ x, data = df)
        r2 <- summary(fit)$r.squared
        intercept <- coef(fit)[1]
        slope <- coef(fit)[2]
        label_text <- sprintf("y = %.2fx %+ .2f\nR² = %.2f", slope, intercept, r2)
      } else {
        label_text <- "n < 3"
      }
      
      p <- ggplot(df, aes(x = x, y = y)) +
        geom_point(size = 1.6, alpha = 0.7) +
        geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
        geom_smooth(method = "lm", se = FALSE, color = "steelblue", linewidth = 0.9) +
        annotate(
          "text",
          x = min(df$x, na.rm = TRUE),
          y = max(df$y, na.rm = TRUE),
          label = label_text,
          hjust = 0, vjust = 1,
          size = 3.2
        ) +
        labs(
          title = paste0(sub("site_", "", site_i), " vs ", sub("site_", "", site_j)),
          x = paste0(sub("site_", "", site_i), " ", value_label),
          y = paste0(sub("site_", "", site_j), " ", value_label)
        ) +
        theme_minimal(base_size = 10) +
        theme(
          plot.title = element_text(face = "bold", hjust = 0.5, size = 10),
          panel.grid.minor = element_blank()
        )
      
      plots[[length(plots) + 1]] <- p
    }
  }
  
  marrangeGrob(plots, nrow = length(site_names), ncol = length(site_names), top = plot_title)
}

```


```{r fig_epa_frm_pairwise, fig.width=18, fig.height=18}
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)

# Clean FRM daily data
epa_24hr_clean <- epa_frm_daily %>%
  group_by(site_number, date) %>%
  summarise(
    pm25 = mean(frm_pm25, na.rm = TRUE),
    .groups = "drop"
  )

# Wide format
epa_wide <- epa_24hr_clean %>%
  select(date, site_number, pm25) %>%
  mutate(site_number = as.character(site_number)) %>%
  pivot_wider(
    names_from = site_number,
    values_from = pm25,
    values_fn = mean
  ) %>%
  filter(rowSums(is.na(across(-date))) < 3)

# Plot settings
site_names <- sort(setdiff(names(epa_wide), "date"))
n_sites <- length(site_names)
axis_min <- 0
axis_max <- 50

plot_matrix <- vector("list", n_sites * n_sites)

for (i in seq_len(n_sites)) {
  for (j in seq_len(n_sites)) {
    
    x_site <- site_names[j]
    y_site <- site_names[i]
    index <- (i - 1) * n_sites + j

    df_pair <- epa_wide %>%
      select(x = all_of(x_site), y = all_of(y_site)) %>%
      filter(!is.na(x) & !is.na(y))

    if (nrow(df_pair) > 2) {
      model <- lm(y ~ x, data = df_pair)
      r2 <- round(summary(model)$r.squared, 2)
      slope <- round(coef(model)[2], 2)
      intercept <- round(coef(model)[1], 2)
      label_text <- paste0("y = ", slope, "x + ", intercept, "\nR² = ", r2)
    } else {
      label_text <- "Insufficient data"
    }

    x_pos <- axis_min + 1
    y_pos <- axis_max - 1

    p <- ggplot(df_pair, aes(x = x, y = y)) +
      geom_point(color = "steelblue", alpha = 0.7, size = 1.8) +
      geom_smooth(method = "lm", se = FALSE, color = "red", linewidth = 1) +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
      coord_fixed(xlim = c(axis_min, axis_max), ylim = c(axis_min, axis_max)) +
      labs(
        x = paste("Site", x_site),
        y = paste("Site", y_site)
      ) +
      annotate(
        "text",
        x = x_pos, y = y_pos,
        label = label_text,
        hjust = 0, vjust = 1,
        size = 3.5
      ) +
      theme_minimal(base_size = 12) +
      theme(
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 9),
        plot.margin = margin(5, 5, 5, 5)
      )

    plot_matrix[[index]] <- p
  }
}

grid_plot <- grid.arrange(
  grobs = plot_matrix,
  ncol = n_sites,
  nrow = n_sites,
  top = "EPA FRM vs FRM Pairwise Comparison"
)

print(grid_plot)

```


```{r epa-6x6-fem-grid-all-days, fig.width=22, fig.height=22}


library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)

# Keep only FEM data on dates where FRM exists at the same site
epa_fem_on_frm_days <- epa_fem_daily %>%
  inner_join(
    epa_frm_daily %>% select(site_number, date),
    by = c("site_number", "date")
  )

# Clean FEM daily data on FRM days only
epa_fem_frm_clean <- epa_fem_on_frm_days %>%
  group_by(site_number, date) %>%
  summarise(
    pm25 = mean(fem_pm25_daily, na.rm = TRUE),
    .groups = "drop"
  )

# Wide format
epa_fem_frm_wide <- epa_fem_frm_clean %>%
  select(date, site_number, pm25) %>%
  mutate(site_number = as.character(site_number)) %>%
  pivot_wider(
    names_from = site_number,
    values_from = pm25,
    values_fn = mean
  ) %>%
  filter(rowSums(is.na(across(-date))) < 3)

# Plot settings
site_names <- sort(setdiff(names(epa_fem_frm_wide), "date"))
n_sites <- length(site_names)
axis_min <- 0
axis_max <- 50

plot_matrix <- vector("list", n_sites * n_sites)

for (i in seq_len(n_sites)) {
  for (j in seq_len(n_sites)) {
    
    x_site <- site_names[j]
    y_site <- site_names[i]
    index <- (i - 1) * n_sites + j

    df_pair <- epa_fem_frm_wide %>%
      select(x = all_of(x_site), y = all_of(y_site)) %>%
      filter(!is.na(x) & !is.na(y))

    if (nrow(df_pair) > 2) {
      model <- lm(y ~ x, data = df_pair)
      r2 <- round(summary(model)$r.squared, 2)
      slope <- round(coef(model)[2], 2)
      intercept <- round(coef(model)[1], 2)
      label_text <- paste0("y = ", slope, "x + ", intercept, "\nR² = ", r2)
    } else {
      label_text <- "Insufficient data"
    }

    x_pos <- axis_min + 1
    y_pos <- axis_max - 1

    p <- ggplot(df_pair, aes(x = x, y = y)) +
      geom_point(color = "steelblue", alpha = 0.7, size = 1.8) +
      geom_smooth(method = "lm", se = FALSE, color = "red", linewidth = 1) +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
      coord_fixed(xlim = c(axis_min, axis_max), ylim = c(axis_min, axis_max)) +
      labs(
        x = paste("Site", x_site),
        y = paste("Site", y_site)
      ) +
      annotate(
        "text",
        x = x_pos, y = y_pos,
        label = label_text,
        hjust = 0, vjust = 1,
        size = 3.5
      ) +
      theme_minimal(base_size = 12) +
      theme(
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 9),
        plot.margin = margin(5, 5, 5, 5)
      )

    plot_matrix[[index]] <- p
  }
}

grid_plot <- grid.arrange(
  grobs = plot_matrix,
  ncol = n_sites,
  nrow = n_sites,
  top = "EPA FEM vs FEM Pairwise Comparison (FRM Days Only)"
)

print(grid_plot)
```



```{r epa-6x6-fem-grid-all-days, fig.width=22, fig.height=22, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)

# Clean FEM daily data (all days)
epa_fem_clean <- epa_fem_daily %>%
  group_by(site_number, date) %>%
  summarise(
    pm25 = mean(fem_pm25_daily, na.rm = TRUE),
    .groups = "drop"
  )

# Wide format
epa_fem_wide <- epa_fem_clean %>%
  select(date, site_number, pm25) %>%
  mutate(site_number = as.character(site_number)) %>%
  pivot_wider(
    names_from = site_number,
    values_from = pm25,
    values_fn = mean
  ) %>%
  filter(rowSums(is.na(across(-date))) < 3)

# Plot settings
site_names <- sort(setdiff(names(epa_fem_wide), "date"))
n_sites <- length(site_names)
axis_min <- 0
axis_max <- 50

plot_matrix <- vector("list", n_sites * n_sites)

for (i in seq_len(n_sites)) {
  for (j in seq_len(n_sites)) {
    
    x_site <- site_names[j]
    y_site <- site_names[i]
    index <- (i - 1) * n_sites + j

    df_pair <- epa_fem_wide %>%
      select(x = all_of(x_site), y = all_of(y_site)) %>%
      filter(!is.na(x) & !is.na(y))

    if (nrow(df_pair) > 2) {
      model <- lm(y ~ x, data = df_pair)
      r2 <- round(summary(model)$r.squared, 2)
      slope <- round(coef(model)[2], 2)
      intercept <- round(coef(model)[1], 2)
      label_text <- paste0("y = ", slope, "x + ", intercept, "\nR² = ", r2)
    } else {
      label_text <- "Insufficient data"
    }

    x_pos <- axis_min + 1
    y_pos <- axis_max - 1

    p <- ggplot(df_pair, aes(x = x, y = y)) +
      geom_point(color = "steelblue", alpha = 0.7, size = 1.8) +
      geom_smooth(method = "lm", se = FALSE, color = "red", linewidth = 1) +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
      coord_fixed(xlim = c(axis_min, axis_max), ylim = c(axis_min, axis_max)) +
      labs(
        x = paste("Site", x_site),
        y = paste("Site", y_site)
      ) +
      annotate(
        "text",
        x = x_pos, y = y_pos,
        label = label_text,
        hjust = 0, vjust = 1,
        size = 3.5
      ) +
      theme_minimal(base_size = 12) +
      theme(
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 9),
        plot.margin = margin(5, 5, 5, 5)
      )

    plot_matrix[[index]] <- p
  }
}

grid_plot <- grid.arrange(
  grobs = plot_matrix,
  ncol = n_sites,
  nrow = n_sites,
  top = "EPA FEM vs FEM Pairwise Comparison (All Days)"
)

print(grid_plot)

```

# EPA site location table 
epa_locations <- epa_data %>%
  mutate(site = paste0("site_", site_number)) %>%
  select(site, latitude, longitude) %>%
  distinct() %>%
  filter(!is.na(latitude), !is.na(longitude)) %>%
  group_by(site) %>%
  summarise(
    lat = first(latitude),
    lon = first(longitude),
    .groups = "drop"
  ) %>%
  arrange(site)

epa_locations



```
# Convert EPA data to wide format

```{r}
library(tidyr)

epa_daily_wide <- epa_daily %>%
  pivot_wider(
    names_from = site_number,
    values_from = pm25_daily,
    names_prefix = "site_"
  )

head(epa_daily_wide)

# EPA correlation matrix
epa_cor_matrix <- cor(
  epa_daily_wide %>% select(-date),
  use = "pairwise.complete.obs"
)

round(epa_cor_matrix, 2)

# EPA correlation heatmap
library(reshape2)
library(ggplot2)

epa_cor_long <- melt(epa_cor_matrix)

ggplot(epa_cor_long, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(
    title = "EPA PM2.5 Correlation Between Sites",
    x = "Site",
    y = "Site",
    fill = "Correlation"
  ) +
  theme_minimal()

round(epa_cor_matrix, 2)

#
library(reshape2)

epa_cor_df <- melt(epa_cor_matrix) %>%
  filter(Var1 != Var2) %>%
  rename(
    site1 = Var1,
    site2 = Var2,
    correlation = value
  )
```

#
```{r}
# epa_locations 
epa_locations <- epa_data %>%
  mutate(site = paste0("site_", site_number)) %>%
  select(site, latitude, longitude) %>%
  distinct() %>%
  rename(
    lat = latitude,
    lon = longitude
  ) %>%
  arrange(site)

epa_locations

epa_locations %>%
  count(site)

epa_locations

# EPA distance table 
library(geosphere)

epa_coords <- epa_locations %>%
  select(lon, lat) %>%
  as.matrix()

epa_dist_matrix <- distm(epa_coords, fun = distHaversine) / 1000

rownames(epa_dist_matrix) <- epa_locations$site
colnames(epa_dist_matrix) <- epa_locations$site

epa_dist_long <- as.data.frame(as.table(epa_dist_matrix))
colnames(epa_dist_long) <- c("site1", "site2", "distance_km")

head(epa_dist_long)

library(ggplot2)
library(gridExtra)

make_epa_pairwise_matrix <- function(wide_df, value_label, plot_title) {
  
  site_names <- names(wide_df)[names(wide_df) != "date"]
  plots <- list()
  
  for (i in seq_along(site_names)) {
    for (j in seq_along(site_names)) {
      
      site_i <- site_names[i]
      site_j <- site_names[j]
      
      df <- wide_df %>%
        select(date, all_of(site_i), all_of(site_j)) %>%
        rename(x = all_of(site_i), y = all_of(site_j)) %>%
        filter(!is.na(x), !is.na(y))
      
      if (nrow(df) > 2) {
        fit <- lm(y ~ x, data = df)
        r2 <- summary(fit)$r.squared
        intercept <- coef(fit)[1]
        slope <- coef(fit)[2]
        label_text <- sprintf("y = %.2fx %+ .2f\nR² = %.2f", slope, intercept, r2)
      } else {
        label_text <- "n < 3"
      }
      
      p <- ggplot(df, aes(x = x, y = y)) +
        geom_point(size = 1.6, alpha = 0.7) +
        geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
        geom_smooth(method = "lm", se = FALSE, color = "steelblue", linewidth = 0.9) +
        annotate(
          "text",
          x = min(df$x, na.rm = TRUE),
          y = max(df$y, na.rm = TRUE),
          label = label_text,
          hjust = 0, vjust = 1,
          size = 3.2
        ) +
        labs(
          title = paste0(sub("site_", "", site_i), " vs ", sub("site_", "", site_j)),
          x = paste0(sub("site_", "", site_i), " ", value_label),
          y = paste0(sub("site_", "", site_j), " ", value_label)
        ) +
        theme_minimal(base_size = 10) +
        theme(
          plot.title = element_text(face = "bold", hjust = 0.5, size = 10),
          panel.grid.minor = element_blank()
        )
      
      plots[[length(plots) + 1]] <- p
    }
  }
  
  marrangeGrob(plots, nrow = length(site_names), ncol = length(site_names), top = plot_title)
}

# EPA pairwise matrix plots
library(ggplot2)
library(gridExtra)

make_epa_pairwise_matrix <- function(wide_df, value_label, plot_title) {
  
  site_names <- names(wide_df)[names(wide_df) != "date"]
  plots <- list()
  
  for (i in seq_along(site_names)) {
    for (j in seq_along(site_names)) {
      
      site_i <- site_names[i]
      site_j <- site_names[j]
      
      df <- wide_df %>%
        select(date, all_of(site_i), all_of(site_j)) %>%
        rename(x = all_of(site_i), y = all_of(site_j)) %>%
        filter(!is.na(x), !is.na(y))
      
      if (nrow(df) > 2) {
        fit <- lm(y ~ x, data = df)
        r2 <- summary(fit)$r.squared
        intercept <- coef(fit)[1]
        slope <- coef(fit)[2]
        label_text <- sprintf("y = %.2fx %+ .2f\nR² = %.2f", slope, intercept, r2)
      } else {
        label_text <- "n < 3"
      }
      
      p <- ggplot(df, aes(x = x, y = y)) +
        geom_point(size = 1.6, alpha = 0.7) +
        geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
        geom_smooth(method = "lm", se = FALSE, color = "steelblue", linewidth = 0.9) +
        annotate(
          "text",
          x = min(df$x, na.rm = TRUE),
          y = max(df$y, na.rm = TRUE),
          label = label_text,
          hjust = 0, vjust = 1,
          size = 3.2
        ) +
        labs(
          title = paste0(sub("site_", "", site_i), " vs ", sub("site_", "", site_j)),
          x = paste0(sub("site_", "", site_i), " ", value_label),
          y = paste0(sub("site_", "", site_j), " ", value_label)
        ) +
        theme_minimal(base_size = 10) +
        theme(
          plot.title = element_text(face = "bold", hjust = 0.5, size = 10),
          panel.grid.minor = element_blank()
        )
      
      plots[[length(plots) + 1]] <- p
    }
  }
  
  marrangeGrob(plots, nrow = length(site_names), ncol = length(site_names), top = plot_title)
}

```

```{r}



```
```

# EPA FRM CORRELATION vs. DISTANCE

```{r}
# distances between EPA sites
library(geosphere)

epa_pairs <- expand.grid(
  site1 = epa_locations$site,
  site2 = epa_locations$site
)

epa_pairs <- epa_pairs %>%
  left_join(epa_locations, by = c("site1" = "site")) %>%
  rename(lat1 = lat, lon1 = lon) %>%
  left_join(epa_locations, by = c("site2" = "site")) %>%
  rename(lat2 = lat, lon2 = lon) %>%
  mutate(
    distance_km = distHaversine(
      cbind(lon1, lat1),
      cbind(lon2, lat2)
    ) / 1000
  )

head(epa_pairs)

# merge with correlations
epa_compare <- epa_cor_df %>%
  left_join(
    epa_pairs,
    by = c("site1", "site2")
  )

# EPA correlation vs distance
epa_compare <- epa_compare %>%
  filter(!is.na(correlation), !is.na(distance_km))

ggplot(epa_compare, aes(distance_km, correlation)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", color = "blue") +
  labs(
    title = "EPA FRM PM2.5 Correlation vs Distance",
    x = "Distance between EPA sites (km)",
    y = "Correlation"
  ) +
  coord_cartesian(ylim = c(0.80, 1.00)) +
  theme_minimal()
```

# FRM filter data (24-hour reference measurements)

```{r}
epa_frm_24hr <- epa_data %>%
  filter(
    method_type == "FRM",
    sample_duration == "24 HOUR"
  )

head(epa_frm_24hr)
```

# FEM optical daily averages

```{r}
epa_fem_24hr <- epa_data %>%
  filter(
    method_type == "FEM",
    sample_duration == "24 HOUR"
  )

epa_data %>%
  count(method_type, sample_duration)
```

```{r}

# EPA hourly dataset
epa_fem_hourly <- epa_data %>%
  filter(method_type == "FEM", sample_duration == "1 HOUR")

# daily averages from hourly FEM
epa_fem_daily <- epa_fem_hourly %>%
  mutate(date = as.Date(date_local)) %>%
  group_by(site_number, date) %>%
  summarise(
    fem_pm25_daily = mean(sample_measurement, na.rm = TRUE),
    n_obs = n(),
    .groups = "drop"
  )

# FRM filter dataset
epa_frm_daily <- epa_data %>%
  filter(method_type == "FRM", sample_duration == "24 HOUR") %>%
  mutate(date = as.Date(date_local)) %>%
  select(site_number, date, sample_measurement) %>%
  rename(pm25_filter = sample_measurement)

# EPA optical vs filter (Merge them by site and date)
epa_compare <- inner_join(
  epa_frm_daily,
  epa_fem_daily,
  by = c("site_number", "date")
)

head(epa_compare)

ggplot(epa_compare,
       aes(pm25_filter, fem_pm25_daily)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue") +
  labs(
    title = "EPA Optical vs Filter PM2.5",
    x = "Filter (FRM) PM2.5",
    y = "Optical (FEM) PM2.5"
  ) +
  theme_minimal()
```

```{r figA-plot, fig.width=12, fig.height=9}
library(dplyr)
library(ggplot2)
library(scales)

# Build plotting table
figA_table <- epa_compare %>%
  filter(!is.na(pm25_filter), !is.na(pm25_daily)) %>%
  mutate(Site = site_number) %>%
  select(
    Site,
    PM25_FRM = pm25_filter,
    PM25_FEM = pm25_daily
  )

# Fixed panel limits to match template style
panel_min <- 0
panel_max <- 50
lims <- c(panel_min, panel_max)

# Per-site regression labels
lab_tbl <- figA_table %>%
  group_by(Site) %>%
  summarise(
    n = n(),
    slope = if (n() > 2) coef(lm(PM25_FEM ~ PM25_FRM))[2] else NA_real_,
    intercept = if (n() > 2) coef(lm(PM25_FEM ~ PM25_FRM))[1] else NA_real_,
    r2 = if (n() > 2) summary(lm(PM25_FEM ~ PM25_FRM))$r.squared else NA_real_,
    .groups = "drop"
  ) %>%
  mutate(
    label = ifelse(
      is.finite(r2),
      sprintf("y = %.2fx %+ .2f\nR² = %.2f", slope, intercept, r2),
      "n < 3"
    ),
    x = panel_min + (panel_max - panel_min) * 0.04,
    y = panel_max - (panel_max - panel_min) * 0.07
  )

ggplot(figA_table, aes(PM25_FRM, PM25_FEM)) +
  geom_point(alpha = 0.9, size = 2) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.9, color = "steelblue") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
  coord_equal(xlim = lims, ylim = lims, expand = FALSE) +
  facet_wrap(~Site, ncol = 3) +
  geom_text(
    data = lab_tbl,
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust = 0,
    vjust = 1,
    size = 4
  ) +
  labs(
    title = "EPA Continuous (FEM daily avg) vs EPA FRM (24hr) — by site",
    x = "EPA FRM PM2.5 (µg/m³)",
    y = "EPA Continuous PM2.5 (µg/m³)"
  ) +
  scale_x_continuous(breaks = seq(0, 50, 5)) +
  scale_y_continuous(breaks = seq(0, 50, 5)) +
  theme_minimal(base_size = 14) +
  theme(
    panel.spacing = unit(1.5, "lines"),
    axis.title = element_text(face = "bold"),
    plot.title = element_text(face = "bold", hjust = 0.5),
    strip.text = element_text(face = "bold", size = 14),
    panel.grid.minor = element_blank()
  )

```

# Compare CinC vs EPA (Filter / FRM)

```{r cinc_epa_frm_comparison, fig.width=10, fig.height=7, echo=FALSE}
# Identify the nearest CinC sensor to each EPA site
library(dplyr)
library(geosphere)

# EPA coordinates
epa_coords <- epa_locations %>%
  select(site, lat, lon)

# CinC coordinates
cinc_coords <- sensor_locations_clean %>%
  select(device_number, latitude, longitude)

# Find nearest CinC sensor for each EPA site
epa_cinc_pairs <- epa_coords %>%
  rowwise() %>%
  mutate(
    nearest_device = cinc_coords$device_number[
      which.min(
        distHaversine(
          c(lon, lat),
          as.matrix(cinc_coords[, c("longitude", "latitude")])
        )
      )
    ]
  ) %>%
  ungroup()

epa_cinc_pairs

# EPA daily filter data
epa_filter_daily <- epa_compare %>%
  select(date, site_number, pm25_filter) %>%
  distinct()

# CinC daily averages
cinc_daily <- cinc_pm25 %>%
  mutate(date = as.Date(time)) %>%
  group_by(device_number, date) %>%
  summarise(
    pm25_cinc = mean(pm_2.5, na.rm = TRUE),
    .groups = "drop"
  )

# EPA–CinC pairing table 
# Rename since epa_cinc_pairs currently uses site, while epa_filter_daily uses site_number
library(dplyr)

epa_cinc_pairs2 <- epa_cinc_pairs2 %>%
  mutate(site_number = sub("^site_", "", site_number))

epa_filter_daily <- epa_filter_daily %>%
  mutate(site_number = as.character(site_number))

cinc_daily <- cinc_daily %>%
  mutate(device_number = as.character(device_number))

# EPA filter data with nearest CinC daily data
epa_cinc_compare <- epa_cinc_pairs2 %>%
  left_join(epa_filter_daily, by = "site_number") %>%
  left_join(
    cinc_daily,
    by = c("nearest_device" = "device_number", "date" = "date")
  )

epa_cinc_compare_plot <- epa_cinc_compare %>%
  filter(!is.na(pm25_filter), !is.na(pm25_cinc))

#Check
head(epa_cinc_compare)

# Plot CinC vs EPA filter by site
library(dplyr)
library(ggplot2)
library(patchwork)

# Build plotting table
figB_table <- epa_cinc_compare %>%
  filter(!is.na(pm25_filter), !is.na(pm25_cinc)) %>%
  mutate(Site = site_number) %>%
  select(
    Site,
    EPA_FRM = pm25_filter,
    CINC_PM25 = pm25_cinc
  ) %>%
  left_join(
    epa_cinc_pairs2 %>%
      mutate(site_number = sub("^site_", "", site_number)),
    by = c("Site" = "site_number")
  )

# Common axis limits
panel_min <- 0
panel_max <- 45
lims <- c(panel_min, panel_max)

# Function to make one panel exactly in template style
make_panel <- function(site_id) {
  
  df <- figB_table %>%
    filter(Site == site_id)
  
  dev_id <- unique(df$nearest_device)[1]
  
  fit <- lm(CINC_PM25 ~ EPA_FRM, data = df)
  slope <- coef(fit)[2]
  intercept <- coef(fit)[1]
  r2 <- summary(fit)$r.squared
  
  eq_label <- sprintf("y = %.2fx %+ .2f\nR² = %.2f", slope, intercept, r2)
  
  ggplot(df, aes(x = EPA_FRM, y = CINC_PM25)) +
    geom_point(alpha = 0.8, size = 1.9) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 1.0, color = "steelblue") +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
    annotate(
      "text",
      x = panel_min + (panel_max - panel_min) * 0.04,
      y = panel_max - (panel_max - panel_min) * 0.07,
      label = eq_label,
      hjust = 0,
      vjust = 1,
      size = 4
    ) +
    coord_equal(xlim = lims, ylim = lims, expand = FALSE) +
    labs(
      title = paste0(site_id),
      x = paste0("Site ", site_id, " (µg/m³)"),
      y = paste0("Device ", dev_id, " (µg/m³)")
    ) +
    scale_x_continuous(breaks = seq(0, 45, 10)) +
    scale_y_continuous(breaks = seq(0, 45, 10)) +
    theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5, size = 13),
      axis.title = element_text(size = 11),
      axis.text = element_text(size = 10, color = "black"),
      panel.grid.minor = element_blank()
    )
}

# Build all six panels
p1 <- make_panel("0034")
p2 <- make_panel("0038")
p3 <- make_panel("0045")
p4 <- make_panel("0060")
p5 <- make_panel("0065")
p6 <- make_panel("0073")

# Combine like the template
(p1 + p2 + p3) / (p4 + p5 + p6) +
  plot_annotation(
    title = "CLEANinCLE vs EPA FRM PM2.5 - by Site",
    theme = theme(
      plot.title = element_text(face = "bold", hjust = 0.5, size = 16)
    )
  )
```

# CLEANinCLE vs EPA optical (daily averages)

```{r fig_cinc_vs_epa_optical, fig.width=11, fig.height=7}
# EPA optical daily data
epa_optical_daily <- epa_fem_daily %>%
  select(date, site_number, pm25_daily) %>%
  rename(pm25_epa_optical = pm25_daily) %>%
  mutate(site_number = as.character(site_number))

epa_cinc_pairs2 <- epa_cinc_pairs2 %>%
  mutate(site_number = sub("^site_", "", site_number))

# EPA optical with CinC daily
epa_cinc_optical_compare <- epa_cinc_pairs2 %>%
  left_join(epa_optical_daily, by = "site_number") %>%
  left_join(
    cinc_daily %>% mutate(device_number = as.character(device_number)),
    by = c("nearest_device" = "device_number", "date" = "date")
  )

# Filter matched rows
epa_cinc_optical_plot <- epa_cinc_optical_compare %>%
  filter(!is.na(pm25_epa_optical), !is.na(pm25_cinc))

# Plot
figC_table <- epa_cinc_optical_compare %>%
  filter(!is.na(pm25_epa_optical), !is.na(pm25_cinc)) %>%
  mutate(Site = site_number) %>%
  select(
    Site,
    EPA_OPTICAL = pm25_epa_optical,
    CINC_PM25 = pm25_cinc
  ) %>%
  left_join(
    epa_cinc_pairs2 %>%
      mutate(site_number = sub("^site_", "", site_number)),
    by = c("Site" = "site_number")
  )

# same style as your FRM plot
make_panel_optical <- function(site_id) {

  df <- figC_table %>%
    filter(Site == site_id)

  dev_id <- unique(df$nearest_device)[1]

  fit <- lm(CINC_PM25 ~ EPA_OPTICAL, data = df)

  slope <- coef(fit)[2]
  intercept <- coef(fit)[1]
  r2 <- summary(fit)$r.squared

  eq_label <- sprintf("y = %.2fx %+ .2f\nR² = %.2f", slope, intercept, r2)

  ggplot(df, aes(EPA_OPTICAL, CINC_PM25)) +
    geom_point(alpha = 0.8, size = 1.9) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 1.0, color = "steelblue") +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
    annotate(
      "text",
      x = min(df$EPA_OPTICAL) + 0.5,
      y = max(df$CINC_PM25) * 0.95,
      label = eq_label,
      hjust = 0,
      vjust = 1,
      size = 4
    ) +
    labs(
      title = site_id,
      x = paste0("Site ", site_id, " (µg/m³)"),
      y = paste0("Device ", dev_id, " (µg/m³)")
    ) +
    scale_x_continuous(breaks = scales::pretty_breaks(5)) +
    scale_y_continuous(breaks = scales::pretty_breaks(5)) +
    theme_minimal(base_size = 14) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5),
      panel.grid.minor = element_blank()
    )
}
p1 <- make_panel_optical("0034")
p2 <- make_panel_optical("0038")
p3 <- make_panel_optical("0045")
p4 <- make_panel_optical("0060")
p5 <- make_panel_optical("0065")
p6 <- make_panel_optical("0073")

(p1 + p2 + p3) / (p4 + p5 + p6) +
  plot_annotation(
    title = "CLEANinCLE vs EPA Optical PM2.5 - by Site",
    theme = theme(
      plot.title = element_text(face = "bold", hjust = 0.5, size = 16)
    )
  )
```

```{r fig_epa_optical_vs_frm, fig.width=11, fig.height=7}
figD_table <- epa_compare %>%
  filter(!is.na(pm25_filter), !is.na(pm25_daily)) %>%
  mutate(Site = site_number) %>%
  select(
    Site,
    EPA_FRM = pm25_filter,
    EPA_OPTICAL = pm25_daily
  )

make_panel_epa <- function(site_id) {

  df <- figD_table %>%
    filter(Site == site_id)

  fit <- lm(EPA_OPTICAL ~ EPA_FRM, data = df)

  slope <- coef(fit)[2]
  intercept <- coef(fit)[1]
  r2 <- summary(fit)$r.squared

  eq_label <- sprintf("y = %.2fx %+ .2f\nR² = %.2f", slope, intercept, r2)

  ggplot(df, aes(EPA_FRM, EPA_OPTICAL)) +
    geom_point(alpha = 0.8, size = 1.9) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 1.0, color = "steelblue") +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
    annotate(
      "text",
      x = min(df$EPA_FRM) + 0.5,
      y = max(df$EPA_OPTICAL) * 0.95,
      label = eq_label,
      hjust = 0,
      vjust = 1,
      size = 4
    ) +
    labs(
      title = site_id,
      x = paste0("EPA FRM — Site ", site_id, " (µg/m³)"),
      y = paste0("EPA Optical — Site ", site_id, " (µg/m³)")
    ) +
    scale_x_continuous(breaks = scales::pretty_breaks(5)) +
    scale_y_continuous(breaks = scales::pretty_breaks(5)) +
    theme_minimal(base_size = 14) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5),
      panel.grid.minor = element_blank()
    )
}

(p1 + p2 + p3) / (p4 + p5 + p6) +
  plot_annotation(
    title = "EPA Optical vs EPA FRM PM2.5 — by Site",
    theme = theme(
      plot.title = element_text(face = "bold", hjust = 0.5, size = 16)
    )
  )

```

```{r}
# EPA optical daily table
epa_optical_daily_wide <- epa_fem_daily %>%
  mutate(site = paste0("site_", site_number)) %>%
  select(date, site, pm25_daily) %>%
  pivot_wider(
    names_from = site,
    values_from = pm25_daily
  )

#Correlation matrix
epa_optical_cor_matrix <- cor(
  epa_optical_daily_wide %>% select(-date),
  use = "pairwise.complete.obs"
)

library(reshape2)
library(ggplot2)

epa_optical_cor_long <- melt(epa_optical_cor_matrix)

ggplot(epa_optical_cor_long, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_viridis_c(name = "Correlation") +
  labs(
    title = "Correlation Between EPA Optical Sites (Daily PM2.5)",
    x = "EPA Optical Site",
    y = "EPA Optical Site"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  )

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
